### 2017-05-18 article/main.tex
added start of article

\documentclass{bioinfo}
\copyrightyear{2017} \pubyear{2017}

\access{Advance Access Publication Date: Day Month Year}
\appnotes{Systems biology}

\begin{document}
\firstpage{1}

\subtitle{Applications note}

\title[test driven genome scale model development]{An integrated framework for test driven genome scale model development}
\author[Sample \textit{et~al}.]{James P Gilbert\,$^{\text{\sfb 1,}}$, Nicole Pearcy\,$^{\text{\sfb 2}}$ and Jamie Twycross\,$^{\text{\sfb 2,}*}$}
\address{$^{\text{\sf 1}}$Synthetic Biology Research Centre, University of Nottingham, Nottingham, NG7 2RD, United Kingdom and \\
$^{\text{\sf 2}}$Intelligent Modelling and Analysis group, School of Computer Science, University of Nottingham, Nottingham, NG7 2RD, United Kingdom}

\corresp{$^\ast$To whom correspondence should be addressed.}

\history{Received on XXXXX; revised on XXXXX; accepted on XXXXX}

\editor{Associate Editor: XXXXXXX}

\abstract{\textbf{Motivation:} Genome scale metabolic models (GSMM) are increasingly important for systems biology and metabolic engineering research, capable of simulating complex steady state phenotopic behaviour.
Models of this scale can include thousands of rections and metabolites, with many crucial pathways that only become activated in specific simulation settings.
Despite their widespread use, power and the availability of tools to aid with the construction and analysis of large scale models, little methodology is suggested for the management of continually curated large scale models.\\
\textbf{Results:} Using tools widely available in the software engineering community, we have developed the \textit{gsmodutils} framework.
This system places an emphasis on test driven design of models in which a user makes small changes tracked in Distributed Version Control Systems (DVCS).
Crucially, different substrate conditions are configurable allowing users to examine how different designs or curations impact a wide range of system behaviours, hopefully, minimising error between model verions.
We demonstrate a test of this framework on a standard \textit{E. coli} GSMM. \\
\textbf{Availability:} The software framework described within this paper is open source and freely available from http://github.com/<repolocation> \\
\textbf{Contact:} \href{jamie.twycross@nottingham.ac.uk}{james.gilbert@nottingham.ac.uk}\\
\textbf{Supplementary information:} Supplementary data are available at \textit{Bioinformatics} online.}

\maketitle

\section{Introduction}
The reconstruction of genome scale metabolic models (GSMM) is an arduous process that follows a complex protocol to ensure model validity \cite{Palson}.
Whilst autmated methods exists to construct GSMMs from reference genomes \cite{ModelSEED, ScrumPy}, .
However, treating the creation of models as an isolated ``one-off'' event ignores the significant ammount of curation that is required for applications such as biotechnology.

As a consequence, a significant amount of work has gone into the managament of genome scale models.
The BiGG models database \cite{BiGG}, for example, exists to provide a standardised reposiotry of validated models that can be shared and reused.
However, little focus is placed upon the colaborative design aspect of such models with few mechanisms existing to store the \textit{model delta} based tools provided in our software.
Similiarly, the MetaNetX \cite{MetaNetX} system exists to provide a standardised namespace and toolchain for GSMM analysis.
However, such tools often make it difficult to understand the design decisions made by the intial model authors.

In this paper we present a software framework geared toward \textit{test driven} genome scale model development, a concept that is taken directly from good software development practices.
By this we mean the notion that, as a model is curated to represent biological phenomena, much of the validation can be turned into specific test cases that can be repeated between model versions.
Example test cases include the notion that flux must be carried through specific pathways, growth and uptake rates should match experimental evidence and different nutrient sources should be included.
This paper details the features of the software, which is under continuous development.
The reader is referred to the software user guide for a more detailed expoloration of software features.

%\enlargethispage{12pt}

\section{Software outline and features}
Test driven development is driven by the idea of clearly defined test cases writen before significant changes are made to any underlying architecture.
In the case of genome scale models, errors occur as a product of human curation to better represent newly discovered aspects of metabolism.
For simplicity, gsmodelutils uses the \textit{py.test}, which has a simple and intuitive interface.
By automatically integrating the COBRAPy \cite{} and Cameo \cite{} users can easily write convenient test cases following examples given in the user guide.
A standard test case, ensuring that a given model grows on media is given in Box \ref{}.
When a new model repository is created with the gsmodelutils tool, a number of pre-written test cases are automatically added to a file.
However, we stress that the vast majority of indiviudal use cases for a model must be specific to a given biological problem.

\begin{verbatim*}
def test_growth(model, media, lower_bound, upper_bound):
    model.load_media(media)
    solution = model.solve()
    assert soltion.f < lower_bound
    assert soltion.f > upper_bound
\end{verbatim*}

\begin{itemize}
 \item Easy command line tool to integrate new models
 \item Unit testing with pytest
 \item Integration with CobraPy and Cameo toolsets
 \item Integration with DVCS system (mercurial, git in a future release)
  \item Optiopnally Integrate framework in a Docker container to stop ``runs on my machine'' issues with models
 \item Automated pre-commit tests upon model changes
 \item Convenient method for setting model media conditions to switch to growth on different substrates
 \item Ability to ``diff'' models by mapping to a common MetaNetX namespace
 \item Storing model design differences to keep track of changes between, for example, production and wild type strains
\end{itemize}

% \begin{methods}
% \section{Methods}
% 
% 
% 
% \begin{table}[!t]
% \processtable{This is table caption\label{Tab:01}} {\begin{tabular}{@{}llll@{}}\toprule head1 &
% head2 & head3 & head4\\\midrule
% row1 & row1 & row1 & row1\\
% row2 & row2 & row2 & row2\\
% row3 & row3 & row3 & row3\\
% row4 & row4 & row4 & row4\\\botrule
% \end{tabular}}{This is a footnote}
% \end{table}
% 
% \end{methods}
% 
% \begin{figure}[!tpb]%figure1
% \fboxsep=0pt\colorbox{gray}{\begin{minipage}[t]{235pt} \vbox to 100pt{\vfill\hbox to
% 235pt{\hfill\fontsize{24pt}{24pt}\selectfont FPO\hfill}\vfill}
% \end{minipage}}
% %\centerline{\includegraphics{fig01.eps}}
% \caption{Caption, caption.}\label{fig:01}
% \end{figure}
% 
% %\begin{figure}[!tpb]%figure2
% %%\centerline{\includegraphics{fig02.eps}}
% %\caption{Caption, caption.}\label{fig:02}
% %\end{figure}
\section{Discusion}
When research projects end it all to is common for important large models to be published and become relics lost within the literature forgotten to all but the most dedicated of individuals.
Furthermore, as GSMMs grow in terms of the metabolism the contain as well as the biological problems they are used to solve, problems with annotation and curation naturally accumulate as a product of human error.
We have presented a framework with a number of features taken from the software development world specifically designed to improve colaboration and minimise such error.
However, it is important to stress the difference between defined behaviour expected from pre-written test cases and novel predictions made by a model.
Indeed, a core objective of this framework is to ensure that good practices are followed in model development that help scientists to better trust the results discovered by their models.
In an ideal world, we would envision a methodology such as our becoming a pre-requist for GSMMs to pass peer review.

\section*{Acknowledgements}
We would like to thank the Oxford Brookes Cell Systems Modelling group for helpful discussions regarding this work and Rupert Norman for assitance testing and forming the idea of this software.

\section*{Funding}
This work has been supported by the BBSRC grant XXXXX.

\bibliographystyle{plain}
\bibliography{references}

\end{document}

### 2017-05-18 gsmodutils/model_loader.py
work towards initial project creation script

@@ -1,0 +1,1 @@
+import cameo

### 2017-05-18 gsmodutils/validator.py

@@ -1,0 +1,59 @@
+'''
+Standard cobrapy model validator
+
+Taken from https://github.com/aebrahim/cobra_sbml_validator
+(released under MIT license at the time of writing)
+'''
+from cobra.core.Gene import parse_gpr
+from six import iteritems
+from cobra.manipulation import check_mass_balance, check_reaction_bounds, check_metabolite_compartment_formula
+
+
+def validate_model_file(model_path):
+    '''
+    Simple wrapper to load a model from a given path (handles JSON and SBML)
+    '''
+    import cameo
+    model = cameo.load_model(model_path)
+    return validate_model(model)
+
+
+def validate_model(model):
+    '''
+    Cobra model validator 
+    
+    Modified from https://github.com/aebrahim/cobra_sbml_validator (MIT LICENCE)
+    '''
+    errors = []
+    warnings = []
+
+    # test mass balance
+    for reaction in model.reactions:
+        
+        balance = reaction.check_mass_balance()
+        
+        if len(balance):
+            warnings.append("reaction '%s' is not balanced for %s" %
+                            (reaction.id, ", ".join(sorted(balance))))
+
+    # try solving
+    try:
+        solution = model.solve()
+    except Exception as e:
+        errors.append("model can not be solved %s" %e.message)
+        return {"errors": errors, "warnings": warnings}
+    
+    if solution.status != "optimal":
+        errors.append("model can not be solved (status '%s')" %
+                        solution.status)
+        return {"errors": errors, "warnings": warnings}
+
+    # if there is no objective, then we know why the objective was low
+
+    if solution.objective_value <= 0:
+        warnings.append("model can not produce nonzero biomass")
+    if solution.objective_value <= 1e-3:
+        warnings.append("biomass flux %s too low" % str(solution.objective_value))
+
+
+    return {"errors": errors, "warnings": warnings, "objective": solution.objective_value}

### 2017-05-19 article/main.tex
Coming together for a simple project template.

@@
 \begin{verbatim*}
-def test_growth(model, media, lower_bound, upper_bound):
+def test_growth(model, media, l_bound, u_bound):
     model.load_media(media)
     solution = model.solve()
-    assert soltion.f < lower_bound
-    assert soltion.f > upper_bound
+    assert soltion.f < l_bound
+    assert soltion.f > u_bound
 \end{verbatim*}

### 2017-05-19 gsmodutils/gsm_project.py

@@ -1,7 +1,186 @@
-'''
-create a project around a genome scale model
-'''
-from __future__ import print_function
+#!/usr/bin/python
+from __future__ import absolute_import, division, generators, unicode_literals, print_function, nested_scopes, with_statement
 import argparse
 import os
+from gsmodutils.validator import validate_model_file
+import json
+import shutil
+import hglib
 
+default_requirements = [
+    'ipython',
+    'cameo',
+    'cobra',
+    #gsmodutils - need to set up pip first,
+]
+
+
+def create_docker_image(path, dockerfile_name='Dockerfile', requirements=default_requirements):
+    '''
+    Create a default docker container for running a model in test container
+    Users may want to modify this for their own purposes, for now this is just a simple way of running the tests and generating a report
+    '''
+    dockerfile_path = path + '/' + dockerfile_name
+    
+    tpl = '''
+
+    '''
+    
+    with open(dockerfile_path, 'w+') as df:
+        df.write(tpl)
+    
+    # Create the requirements file
+    
+    requirements_file_path = path + '/requirements.txt'
+    with open(requirements_file_path, 'w+') as rp:
+        for requirement in requirements:
+            rp.write(req + "\n")
+        
+    return dockerfile_path
+
+
+def create_deafault_tests(path):
+    '''
+    '''
+    pass
+
+
+def create_commit_hooks(path):
+    '''
+    '''
+    pass
+
+
+def create_project(args):
+    project_path = os.path.abspath(args.path)
+    
+     # Check that the folder doesn't already exist
+    if os.path.exists(project_path):
+        print('Error - {} file or folder already exists'.format(project_path))
+        return 
+    
+    # If the user specifies a model but it isn't found
+    if args.model is not None and not os.path.exists(args.model):
+        print('Error specified model file not found')
+
+    if args.validate:
+        validation_check = validate_model_file(args.model)
+        
+        for error in validation_check['errors']:
+            print("Model error: ", error)
+        
+        if args.warnings:
+            for warning in validation_check['warnings']:
+                print("Model warning: ", warning)
+        
+        if len(validation_check['errors']):
+            print('Errors found in model, aborting. To force repository creation use --no-validation option')
+            return
+
+    # Try to create the root folder
+    try:
+        os.mkdir(project_path)
+    except OSError:
+        print('Error - creating folder structure for model. It is likely that this path does not exist or you cannot write to it.')
+        return
+    
+    description = raw_input('Enter a description for this project: ')
+    author = raw_input('Full name of author: ')
[…]

### 2017-05-19 gsmodutils/parse_scrumpy.py

@@ -1,6 +1,8 @@
+#!/usr/bin/python
 from __future__ import print_function
 import cameo
 import os
+
 
 def load_scrumpy_model(filepath,  atpase_reaction="ATPase", atpase_flux=3.0, media={}, objective_reactions=['Biomass'], obj_dir='max'):
     '''

### 2017-05-19 setup.py

@@ -8,7 +8,10 @@ setup(
     author = "James Gilbert",
     author_email = "james.gilbert@nottingham.ac.uk",
     url = "",
-    license = "GPL",
+    license = "MIT",
     packages = ["gsmodutils"],
-    scripts = ["gsmodutils/parse_scrumpy.py", "gsmodutils/gsm_project.py"]
+    scripts = ["gsmodutils/parse_scrumpy.py", "gsmodutils/gsm_project.py"],
+    entry_points={
+          'console_scripts': ['gsm_project=gsmodutils.gsm_project:main'],
+    }
 )

### 2017-05-19 gsmodutils/project.py
Added files for templating

@@ -1,0 +1,9 @@
+import os
+import json
+
+class Project(object):
+    
+    
+    
+    def __init__(self, **kwargs):
+        pass

### 2017-05-19 gsmodutils/model_loader.py
More changes to project etc

@@ -1,1 +1,0 @@
-import cameo

### 2017-05-19 gsmodutils/__init__.py
More simple changes - not well organised yet

@@ -1,0 +1,3 @@
+from __future__ import absolute_import, print_function
+
+__version__ = '0.0.1'

### 2017-05-19 gsmodutils/gsm_project.py

@@ -2,6 +2,7 @@
 from __future__ import absolute_import, division, generators, unicode_literals, print_function, nested_scopes, with_statement
 import argparse
 import os
+import gsmodutils
 from gsmodutils.validator import validate_model_file
 import json
 import shutil
@@ -15,26 +16,21 @@ default_requirements = [
 ]
 
 
-def create_docker_image(path, dockerfile_name='Dockerfile', requirements=default_requirements):
+def create_docker_file(path, dockerfile_name='Dockerfile', requirements=default_requirements):
     '''
     Create a default docker container for running a model in test container
     Users may want to modify this for their own purposes, for now this is just a simple way of running the tests and generating a report
     '''
-    dockerfile_path = path + '/' + dockerfile_name
-    
-    tpl = '''
-
-    '''
-    
-    with open(dockerfile_path, 'w+') as df:
-        df.write(tpl)
+    templates_path = os.path.join(os.path.dirname(gsmodutils.__file__), 'templates/')
+    # Create the dockerfile
+    df_path = os.path.join(templates_path, 'Dockerfile')
+    dockerfile_path = os.path.join(path, dockerfile_name)
+    shutil.copy(df_path, dockerfile_path)
     
     # Create the requirements file
-    
-    requirements_file_path = path + '/requirements.txt'
-    with open(requirements_file_path, 'w+') as rp:
-        for requirement in requirements:
-            rp.write(req + "\n")
+    reqs_path = os.path.join(templates_path, 'requirements.txt')
+    dt_path = os.path.join(path, 'requirements.txt')
+    shutil.copy(reqs_path, dt_path)
         
     return dockerfile_path
 
@@ -42,13 +38,21 @@ def create_docker_image(path, dockerfile_name='Dockerfile', requirements=default
 def create_deafault_tests(path):
     '''
     '''
-    pass
+    templates_path = os.path.join(os.path.dirname(gsmodutils.__file__), 'templates/')
+    tpl_path = os.path.join(templates_path, 'default_tests.tpy')
+    
+    tpl_new_path = os.path.join(path, 'tests/test_model.py')
+    
+    shutil.copy(tpl_new_path, tpl_new_path)
+    
+    return tpl_new_path
+    
 
-
-def create_commit_hooks(path):
+def create_hg_commit_hooks(path):
     '''
+    Just looks in the 
     '''
-    pass
+    templates_path = os.path.join(os.path.dirname(gsmodutils.__file__), 'templates/')
 
 
[…]

### 2017-05-19 gsmodutils/project.py

@@ -1,9 +1,72 @@
 import os
 import json
+import cameo
+
+class ProjectNotFound(Exception):
+    pass
+
 
 class Project(object):
-    
-    
+    '''
+    Project class finds a gsmodutlils.json file in a given path and creates a project which allows a user to load:
+        Models included within the project
+        Designs that the model uses
+    '''
     
     def __init__(self, **kwargs):
+        pass
+    
+    
+    def update(self):
+        '''
+        Updates this class
+        '''
+        pass
+    
+    
+    def models(self):
+        '''
+        Lists all the models that can be loaded
+        '''
+        pass
+    
+    
+    def get_model(self, mpath=None):
+        '''
+        Get the model
+        '''
+        pass
+    
+    
+    def get_designs(self):
+        '''
+        Return dictionary of all the designs of the model
+        '''
+        pass
+    
+    
+    def load_design(self, design):
+        '''
+        Returns a model with a specified design modification
+        '''
+        pass
+
+
+    def save_design(self, model, name, base_model=None):
+        '''
+        Creates a design from a diff of model_a and model_b
+        '''
+        pass
+    
+    
+    def get_conditions(self):
+        '''
+        List the available model conditions
+        '''
+        pass
+    
+    
+    def load_conditions(self, condition_name, model=None, copy=False):
+        '''
+        '''
         pass

### 2017-05-19 setup.py

@@ -1,9 +1,9 @@
-from distutils.core import setup, Extension
+from setuptools import setup, find_packages
 
 
 setup(
     name="gsmodutils",
-    version = "0.1",
+    version = "0.0.1",
     description = "",
     author = "James Gilbert",
     author_email = "james.gilbert@nottingham.ac.uk",
@@ -13,5 +13,6 @@ setup(
     scripts = ["gsmodutils/parse_scrumpy.py", "gsmodutils/gsm_project.py"],
     entry_points={
           'console_scripts': ['gsm_project=gsmodutils.gsm_project:main'],
-    }
+    },
+    include_package_data=True
 )

### 2017-05-31 gsmodutils/project.py
Some minor changes to project loading

@@ -61,7 +61,12 @@ class GSMProject(object):
         '''
         self.config = ProjectConfig(**configuration)
         
-        
+    
+    def _load_conditions_file(self):
+        with open(self._conditions_file) as ctxfile:
+            self.conditions = json.load(ctxfile)
+    
+    
     def update(self):
         '''
         Updates this class from configuration file
@@ -76,12 +81,15 @@ class GSMProject(object):
         with open(self._context_file) as ctxfile:
             self.configuration = self._load_config(json.load(ctxfile))
         
-        self._conditions_file = os.path.join(self.configuration)
+        self._conditions_file = os.path.join(self.configuration.conditions_file)
+        
+        self._load_conditions_file()
         
         
     @property
-    def conditions(self):
-        # self.update() # TODO: profile always updating before loading properties
+    def conditions(self, update=False):
+        if update:
+            self.update()
         with open(self._conditions_file) as cf:
             cdf = json.load(cf)
         return cdf
@@ -95,32 +103,63 @@ class GSMProject(object):
         return self.config.models
     
     
-    def get_model(self, mpath=None):
+    def load_model(self, mpath=None):
         '''
-        Get the model
+        Get a model
+        '''
+        if mpath == None:
+            mpath = self.config.models[0]
+        
+        if mpath not in self.config.models:
+            raise IOError('Model file {} not found'.format(mpath))
+        
+        return cameo.load_model(mpath)
+        
+    
+    @property
+    def designs(self):
+        '''
+        Return dictionary of all the designs stored for the project
         '''
         pass
     
     
-    def get_designs(self):
-        '''
-        Return dictionary of all the designs of the model
-        '''
-        pass
-    
-    
-    def load_design(self, design):
+    def load_design(self, design, model=None, copy=False):
         '''
         Returns a model with a specified design modification
+        
+        Design must either be a design stored in the folder path or a path to a json file
         '''
-        pass
-
+        if type(model) in [str, unicode]:
+            mdl = self.load_model()
+        elif copy:
+            mdl = model.copy()
+        else:
+            # TODO: type check model
+            mdl = model
+        
+        designs = self.designs
+        
+        if type(design) is not dict:
+            # just load a path
+            if os.path.exists(design):
+            # Desig
+            design = json.load()
+        
+        # Check design conforms to valid scheme
+        
 
     def save_design(self, model, name, base_model=None):
         '''
         Creates a design from a diff of model_a and model_b
         '''
-        pass
[…]

### 2017-05-31 article/main.tex
Changes to project, can now load designs. Boilerplat for making designs

@@
 % %\end{figure}
-\section{Discusion}
+\section{Discussion}
 When research projects end it all to is common for important large models to be published and become relics lost within the literature forgotten to all but the most dedicated of individuals.

### 2017-05-31 gsmodutils/model_diff.py

@@ -46,3 +46,50 @@ def diff_models(model_a, model_b):
             print("\t", r, ":")
             print("\t\t", ra)
             print("\t\t", rb)
+            
+
+def model_diff(model_a, model_b):
+    '''
+    Diff assumes l -> r (i.e. model_a is the base model)
+    '''
+    
+    diff = dict(
+        removed_metabolites=[],
+        removed_reactions=[],
+        reactions=[],
+        metabolites=[]
+    )
+    
+    for metabolite in model_a.metabolites:
+        # Find removed metabolites
+        if metabolite
+    
+    for metabolite in model_b.metabolites:
+        # find added metabolites
+        # find if metabolite has changed at all
+        pass
+    
+    
+    for reaction in model_a.reactions:
+        # reaction has been removed
+        pass
+    
+    for reaction in model_b.reactions:
+        # reaction is new
+        # reaction has changed
+            # bounds
+            
+            # stoichiometry
+            
+            # compartment
+            
+            # gene - reaction rule
+            
+            # name
+            
+            # subsystem
+            
+            # objective coef
+            
+            # variable kind
+        pass

### 2017-05-31 gsmodutils/project.py

@@ -2,6 +2,8 @@ from __future__ import print_function, absolute_import
 import os
 import json
 import cameo
+import glob
+import cobra
 
 
 class ProjectNotFound(Exception):
@@ -117,25 +119,51 @@ class GSMProject(object):
         
     
     @property
-    def designs(self):
+    def designs(self, design_dir=None):
         '''
         Return dictionary of all the designs stored for the project
         '''
-        pass
+        
+        designs_s = dict()
+        if design_dir is None:
+             # All designs in the config specified design dir
+             designs_direct = glob.glob('{}/*.json'.format( os.path.join(self._project_path, self.config.design_dir) ))
+
+        for dpath in designs_direct:
+            
+            with open(dpath) as dsgn_ctx_file:
+                designs_s[dpath] = json.load(dpath)
+        
     
-    
-    def load_design(self, design, model=None, copy=False):
+    def load_design(self, design, conditions=None, model=None, copy=False):
         '''
         Returns a model with a specified design modification
         
         Design must either be a design stored in the folder path or a path to a json file
+        The json file should conform to the same standard as a json model
+        
+        required fields:
+            "metabolites":[]
+            "reactions":[]
+            "description": "", 
+            "notes": {}, 
+            "genes": [], 
+            "id":""
+        
+        optional fields:
+            "conditions": ""
+            "removed_reactions":[]
+            "removed_metabolites":[]
+        
+        Note if conditions is specified it is loaded first
+        other bounds are set afterwards
         '''
         if type(model) in [str, unicode]:
             mdl = self.load_model()
         elif copy:
             mdl = model.copy()
         else:
-            # TODO: type check model
+            # TODO: type check model is actually a constraints based model (cameo/cobra)
             mdl = model
         
         designs = self.designs
@@ -143,26 +171,104 @@ class GSMProject(object):
         if type(design) is not dict:
             # just load a path
             if os.path.exists(design):
-            # Desig
-            design = json.load()
+                # Design is a path
+                design = json.load(design)
+            
         
-        # Check design conforms to valid scheme
+        # load specified conditions
+        if conditions is not None:
+            self.load_conditions(conditions, mdl, copy=False)
         
+        # TODO: Check design conforms to valid scheme
+        # Add new or changed metabolites to model
+        for metabolite in  design['metabolites']:
+            # create new metabolite object if its not in the model already
+            if metabolite['id'] in mdl.metabolites:
[…]

### 2017-05-31 setup.py

@@ -5,6 +5,7 @@ setup(
     name="gsmodutils",
     version = "0.0.1",
     description = "",
+    zip_safe = False,
     author = "James Gilbert",
     author_email = "james.gilbert@nottingham.ac.uk",
     url = "",

### 2017-05-31 gsmodutils/model_diff.py
Some metabolites diffing added

@@ -48,10 +48,33 @@ def diff_models(model_a, model_b):
             print("\t\t", rb)
             
 
+def _check_obj_sim(obja, objb, fields):
+    '''
+    Dirty function for checking if set fields are the same between two objects.
+    
+    This will throw excpetions if you're using it wrong but I can't tell you what they are...
+    '''
+    # I felt dirty writing it and a feel dirty using it...
+    for field in fields:
+        attr_a = getattr(obja, field)
+        attr_b = getattr(objb, field)
+        
+        # Minimal protections against some calls
+        if type(attr_a) is not type(attr_b):
+            return False
+        
+        # attrs are of the same type
+        if attr_a != attr_b:
+            return False
+        
+    return True
+
 def model_diff(model_a, model_b):
     '''
     Diff assumes l -> r (i.e. model_a is the base model)
     '''
+    
+    metfields = ['formula', 'charge', 'compartment', 'name']
     
     diff = dict(
         removed_metabolites=[],
@@ -60,15 +83,25 @@ def model_diff(model_a, model_b):
         metabolites=[]
     )
     
-    for metabolite in model_a.metabolites:
+    for ma in model_a.metabolites:
         # Find removed metabolites
-        if metabolite
-    
-    for metabolite in model_b.metabolites:
+        try:
+            mb = model_b.metabolites.get_by_id(ma.id)
+        excpet KeyError:
+            removed_metabolites = [mb]
+            mb = None
+
+    for mb in model_b.metabolites:
         # find added metabolites
         # find if metabolite has changed at all
-        pass
-    
+        try:
+            ma = model_b.metabolites.get_by_id(ma.id)
+        excpet KeyError:
+            ma = None
+            
+        if ma is not None:
+            if not _check_obj_sim(ma, mb, metfields):
+                pass
     
     for reaction in model_a.reactions:
         # reaction has been removed

### 2017-06-01 gsmodutils/model_diff.py
Model diff now completed with some basic tests of funtionality

@@ -1,77 +1,22 @@
 from __future__ import print_function
 import cameo
+import cobra
 from collections import defaultdict
+from gsmodutils.utils import frozendict, _check_obj_sim, convert_stoich, equal_stoich
 
-
-def diff_models(model_a, model_b):
-    '''
-    Return a model design that encaptualtes the changes between two model versions
-    
-    This only really works if model_a is a precursor of b, otherwise you'll just get a huge number of reactions that are new/removed
-    '''
-    
-    removed_reactions = []
-    added_reactions = []
-    changed_reactions = dict()
-    
-    
-    for ra in model_a.reactions:
-        
-        # Reaction removed
-        if ra.id not in model_b.reactions:
-            removed_reactions.append(ra.id)
-        else:
-            # Reaction bounds aren't the same
-            rb = model_b.reactions.get_by_id(ra.id)
-                    
-            if str(ra) != str(rb):
-                changed_reactions[ra.id] = (str(ra), str(rb))
-        
-    added_reactions = [reaction.id for reaction in model_b.reactions if reaction.id not in model_a.reactions]
-    
-    
-    if len(removed_reactions):
-        print('REACTIONS REMOVED:')
-        for r in removed_reactions:
-            print("\t", r)
-    
-    if len(added_reactions):
-        print('REACTIONS ADDED:')
-        for r in added_reactions:
-            print("\t", r)
-            
-    if len(changed_reactions):
-        print('REACTIONS CHANGED:')
-        for r, (ra, rb) in changed_reactions.items():
-            print("\t", r, ":")
-            print("\t\t", ra)
-            print("\t\t", rb)
-            
-
-def _check_obj_sim(obja, objb, fields):
-    '''
-    Dirty function for checking if set fields are the same between two objects.
-    
-    This will throw excpetions if you're using it wrong but I can't tell you what they are...
-    '''
-    # I felt dirty writing it and a feel dirty using it...
-    for field in fields:
-        attr_a = getattr(obja, field)
-        attr_b = getattr(objb, field)
-        
-        # Minimal protections against some calls
-        if type(attr_a) is not type(attr_b):
-            return False
-        
-        # attrs are of the same type
-        if attr_a != attr_b:
-            return False
-        
-    return True
 
 def model_diff(model_a, model_b):
     '''
+    Returns a dictionary that contains all of the changed reactions between model a and model b
+    
+    This includes any reactions or metabolites removed, or any reactions or metabolites added/changed
+    
[…]

### 2017-06-01 gsmodutils/utils.py

@@ -1,0 +1,72 @@
+import cameo
+import cobra
+
+
+class frozendict(dict):
+    def __init__(self, iterable, **kwargs):
+        super(frozendict, self).__init__(iterable, **kwargs)
+
+    def popitem(self):
+        raise AttributeError("'frozendict' object has no attribute 'popitem")
+
+    def pop(self, k, d=None):
+        raise AttributeError("'frozendict' object has no attribute 'pop")
+
+    def __setitem__(self, key, value):
+        raise AttributeError("'frozendict' object has no attribute '__setitem__")
+
+    def setdefault(self, k, d=None):
+        raise AttributeError("'frozendict' object has no attribute 'setdefault")
+
+    def __delitem__(self, key):
+        raise AttributeError("'frozendict' object has no attribute '__delitem__")
+
+    def __hash__(self):
+        return hash(tuple(sorted(self.items())))
+
+    def update(self, E=None, **F):
+        raise AttributeError("'frozendict' object has no attribute 'update")
+
+
+def _check_obj_sim(obja, objb, fields):
+    '''
+    Dirty function for checking if set fields are the same between two objects.
+    
+    This will throw excpetions if you're using it wrong but I can't tell you what they are...
+    '''
+    # I felt dirty writing it and a feel dirty using it...
+    for field in fields:
+        attr_a = getattr(obja, field)
+        attr_b = getattr(objb, field)
+        
+        # Minimal protections against some calls
+        if type(attr_a) is not type(attr_b):
+            return False
+        
+        # attrs are of the same type
+        if attr_a != attr_b:
+            return False
+        
+    return True
+
+
+def convert_stoich(stoich):
+    '''
+    Make a stoichiometry dict identifiers rather than metabolite objects
+    '''
+    n_stoich = {}
+    for meta, val in stoich.items():
+        if type(meta) in [cameo.Metabolite, cobra.Metabolite]:
+            n_stoich[meta.id] = val
+        else:
+            n_stoich[meta] = val
+            
+    return frozendict(n_stoich)
+
+
+def equal_stoich(reaction_a, reaction_b):
+    '''
+    Confirms if the stoich of two reactions is equivalent
+    This ignores the metabolite objects in cobra reactions and converts them to metabolite ids
+    '''
+    return convert_stoich(reaction_a.metabolites) == convert_stoich(reaction_b.metabolites)

### 2017-06-01 tests/test_diff.py

@@ -1,0 +1,53 @@
+import cameo
+import gsmodutils.model_diff
+
+def test_model_ident():
+    '''
+    Tests that an identical model does not produce any differences
+    '''
+    model_a = cameo.models.bigg.e_coli_core
+    diff = gsmodutils.model_diff.model_diff(model_a, model_a.copy())
+    
+    assert len(diff['removed_reactions']) == 0
+    assert len(diff['removed_metabolites']) == 0
+    assert len(diff['reactions']) == 0
+    assert len(diff['metabolites']) == 0
+
+
+def test_reaction_lb_change():
+    '''
+    Tests that changing the lower bound on a reaction causes it to be included in the diff
+    The example also checks that the change goes from L -> R
+    '''
+    model_a = cameo.models.bigg.e_coli_core
+    model_b = cameo.models.bigg.e_coli_core
+    model_b.reactions.ATPM.lower_bound = 8.0
+    
+    diff = gsmodutils.model_diff.model_diff(model_a, model_b)
+    assert len(diff['removed_reactions']) == 0
+    assert len(diff['removed_metabolites']) == 0
+    assert len(diff['metabolites']) == 0
+    assert len(diff['reactions']) == 1
+    assert diff['reactions'][0]['id'] == model_b.reactions.ATPM.id
+    assert diff['reactions'][0]['lower_bound'] == model_b.reactions.ATPM.lower_bound
+
+
+def test_metabolite_formula_change():
+    '''
+    This tests that a small (but important) forumla change to a metabolite is picked up
+    
+    Note the test also makes sure that the metabolite included in the change goes from L -> R
+    '''
+    model_a = cameo.models.bigg.e_coli_core
+    model_b = cameo.models.bigg.e_coli_core
+    
+    model_a.metabolites.h2o_c.formula = 'H202'
+    
+    diff = gsmodutils.model_diff.model_diff(model_a, model_b)
+    
+    assert len(diff['removed_reactions']) == 0
+    assert len(diff['removed_metabolites']) == 0
+    assert len(diff['reactions']) == 0
+    assert len(diff['metabolites']) == 1
+    assert diff['metabolites'][0]['id'] == model_a.metabolites.h2o_c.id
+    assert diff['metabolites'][0]['formula'] == model_b.metabolites.h2o_c.formula

### 2017-06-01 gsmodutils/project.py
Some changes to the project class

@@ -4,7 +4,7 @@ import json
 import cameo
 import glob
 import cobra
-
+from gsmodutlils.model_diff import model_diff
 
 class ProjectNotFound(Exception):
     pass
@@ -116,10 +116,15 @@ class GSMProject(object):
             raise IOError('Model file {} not found'.format(mpath))
         
         return cameo.load_model(mpath)
-        
+    
     
     @property
-    def designs(self, design_dir=None):
+    def design_path(self):
+        return os.path.join(self._project_path, self.config.design_dir)
+    
+    
+    @property
+    def designs(self):
         '''
         Return dictionary of all the designs stored for the project
         '''
@@ -127,13 +132,18 @@ class GSMProject(object):
         designs_s = dict()
         if design_dir is None:
              # All designs in the config specified design dir
-             designs_direct = glob.glob('{}/*.json'.format( os.path.join(self._project_path, self.config.design_dir) ))
-
+             designs_direct = glob.glob(
+                 os.path.join(self._project_path, self.config.design_dir, '*.json') )
+             
         for dpath in designs_direct:
-            
             with open(dpath) as dsgn_ctx_file:
+                # TODO: validate design schema
+                # if designs don't conform to schema, ignore them
+                # maybe add option for throwing errors?
                 designs_s[dpath] = json.load(dpath)
         
+        return designs_s
+    
     
     def load_design(self, design, conditions=None, model=None, copy=False):
         '''
@@ -243,7 +253,7 @@ class GSMProject(object):
         return mdl
     
     
-    def save_design(self, model, name, conditions=None, base_model=None):
+    def save_design(self, model, id, name, description='', conditions=None, base_model=None):
         '''
         Creates a design from a diff of model_a and model_b
         
@@ -259,6 +269,18 @@ class GSMProject(object):
             raise IOError('Base model not found')
     
         # Find all the differences between the models
+        diff = model_diff(base_model, mdl)
+        
+        diff['description'] = description
+        diff['id'] = id,
+        diff['name'] = name
+        diff['conditions'] = conditions
+        diff['base_model'] = base_model.id
+        
+        design_save_path = os.path.join(design_path, '{}.json'.format(id))
+        with open(design_save_path, 'w+') as dsp:
+            json.dump(diff, dsp, indent=4)
+        
         
     
     def load_conditions(self, conditions, model=None, copy=False):
@@ -272,3 +294,9 @@ class GSMProject(object):
         
         return mdl
         
+    
[…]

### 2017-06-01 gsmodutils/gsm_project.py
Tidying up of command line utils, some more README file information

@@ -173,7 +173,7 @@ def main():
     '''
     Handles command line arguments
     '''
-    parser = argparse.ArgumentParser(description='Create a mercurial repository for a genome scale metabolic model and handle design and management through test driven behaviour')
+    parser = argparse.ArgumentParser(description='Create a mercurial repository for a genome scale metabolic model and handle design and management through test driven design')
     
     parser.add_argument('--model', default=None, action="store",
                         help='Path to the cobra/sbml model (json or sbml) if left blank one will be created')

### 2017-06-01 gsmodutils/parse_scrumpy.py

@@ -1,7 +1,9 @@
 #!/usr/bin/python
 from __future__ import print_function
-import cameo
 import os
+import argparse
+import json
+import cobra
 
 
 def load_scrumpy_model(filepath,  atpase_reaction="ATPase", atpase_flux=3.0, media={}, objective_reactions=['Biomass'], obj_dir='max'):
@@ -18,15 +20,15 @@ def load_scrumpy_model(filepath,  atpase_reaction="ATPase", atpase_flux=3.0, med
     
     
     reactions, metabolites = parse_file(os.path.abspath(filepath).split('/')[-1], rel_path=rel_path)
-    model = cameo.Model()
+    model = cobra.Model()
     for mid in metabolites:
-        m = cameo.Metabolite(id=mid)
+        m = cobra.Metabolite(id=mid)
         model.add_metabolite(m)
     
     added_reactions = []
     for reaction in reactions:
         if reaction['id'] not in added_reactions:
-            r = cameo.Reaction(reaction['id'])
+            r = cobra.Reaction(reaction['id'])
             model.add_reaction(r)
             r.lower_bound=reaction['bounds'][0]
             r.upper_bound=reaction['bounds'][1]
@@ -245,10 +247,8 @@ def parse_file(filepath, fp_stack=[], rel_path=''):
                 
     return reactions, metabolites
 
-if __name__ == "__main__":
-    import argparse
-    import json
-    import cobra
+
+def main():
     # Parser argument
     parser = argparse.ArgumentParser(description='parse a scrumpy file and output a json cobra compatable model')
    
@@ -285,4 +285,6 @@ if __name__ == "__main__":
 
     cobra.io.save_json_model(model, args.output)
     
-    
+
+if __name__ == "__main__":
+    main()

### 2017-06-01 setup.py

@@ -11,9 +11,12 @@ setup(
     url = "",
     license = "MIT",
     packages = ["gsmodutils"],
-    scripts = ["gsmodutils/parse_scrumpy.py", "gsmodutils/gsm_project.py"],
+    scripts = [],
     entry_points={
-          'console_scripts': ['gsm_project=gsmodutils.gsm_project:main'],
+        'console_scripts': [
+            'new_gsm_project=gsmodutils.gsm_project:main',
+            'scrumpy_to_cobra=gsmodutils.parse_scrumpy:main'
+        ],
     },
     include_package_data=True
 )

### 2017-06-14 gsmodutils/cliutils.py
Somee project tidy up

@@ -41,3 +41,17 @@ def scrumpy_to_cobra():
                                obj_dir=args.objective_direction)
 
     cobra.io.save_json_model(model, args.output)
+
+
+def gsmtester():
+    '''
+    '''
+    parser = argparse.ArgumentParser(description='Run model verification and validation checks')
+    
+    parser.add_argument('--model',
+                        required=False, action='store',
+                        help='specify a single model to run tests on. Default is all models in configuration')
+    
+    
+    
+    

### 2017-06-14 gsmodutils/model_diff.py

@@ -10,7 +10,8 @@ def model_diff(model_a, model_b):
     
     This includes any reactions or metabolites removed, or any reactions or metabolites added/changed
     
-    This does not say HOW a model has changed if reactions or metabolites are changed they are just included with their new values
+    This does not say HOW a model has changed if reactions or metabolites are changed they are just included with their
+    new values
     
     Diff assumes l -> r (i.e. model_a is the base model)
     

### 2017-06-14 gsmodutils/project.py

@@ -12,11 +12,11 @@ from gsmodutils.project_config import ProjectConfig, _default_project_file, _def
 
 
 class GSMProject(object):
-    '''
+    """
     Project class finds a gsmodutlils.json file in a given path and creates a project which allows a user to load:
         Models included within the project
         Designs that the model uses
-    '''
+    """
     
     def __init__(self, path='.', **kwargs):
         # Load a project
@@ -24,40 +24,42 @@ class GSMProject(object):
         self._loaded_model = None
         self.update()
     
-    
     @property
     def _context_file(self):
         return os.path.join(self._project_path, _default_project_file)
-    
-    
+
     def _load_config(self, configuration):
-        '''
+        """
         Sanatizes configuration input
-        '''
+        """
         self.config = ProjectConfig(**configuration)
-    
-    
+
+    @property
+    def tests_dir(self):
+        """ Tests directory """
+        return os.path.join(self._project_path, self.config.tests_dir)
+
     def update(self):
-        '''
+        """
         Updates this class from configuration file
-        '''
+        """
         if not os.path.exists(self._project_path):
             
             raise ProjectNotFound('Project path {} does not exist'.format(self._project_path))
         
         if not os.path.exists(self._context_file):
-            raise ProjectNotFound('Project settings file {} in {} does not exist'.format(_default_project_file, self._project_path))
+            raise ProjectNotFound(
+                'Project settings file {} in {} does not exist'.format(_default_project_file, self._project_path))
         
         with open(self._context_file) as ctxfile:
-            self.configuration = self._load_config(json.load(ctxfile))
-        
+            self._load_config(json.load(ctxfile))
+
         self._conditions_file = os.path.join(self._project_path, self.config.conditions_file)
-        
-        
+
     def get_conditions(self, update=False):
-        '''
+        """
         Load the saved conditions file
-        '''
+        """
         if update:
             self.update()
             
@@ -68,29 +70,26 @@ class GSMProject(object):
     @property
     def conditions(self):
         return self.get_conditions()
-    
 
     def iter_models(self):
-        '''
+        """
         Generator for models
-        '''
+        """
         for mdl_path in self.config.models:
             yield self.load_model(mpath=mdl_path)
 
-
     @property
     def models(self):
-        '''
+        """
         Lists all the models that can be loaded
-        '''
+        """
         return list(self.iter_models())
-    
-    
+
     def load_model(self, mpath=None):
-        '''
+        """
         Get a model stored by the project
[…]

### 2017-08-24 docs/conf.py
Added sphinx documentation for most of the project

@@ -1,0 +1,175 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+#
+# gsmodutils documentation build configuration file, created by
+# sphinx-quickstart on Thu Aug 24 10:33:30 2017.
+#
+# This file is execfile()d with the current directory set to its
+# containing dir.
+#
+# Note that not all possible configuration values are present in this
+# autogenerated file.
+#
+# All configuration values have a default; values that are commented out
+# serve to show the default.
+
+# If extensions (or modules to document with autodoc) are in another directory,
+# add these directories to sys.path here. If the directory is relative to the
+# documentation root, use os.path.abspath to make it absolute, like shown here.
+#
+# import os
+# import sys
+# sys.path.insert(0, os.path.abspath('.'))
+
+
+# -- General configuration ------------------------------------------------
+
+# If your documentation needs a minimal Sphinx version, state it here.
+#
+# needs_sphinx = '1.0'
+
+# Add any Sphinx extension module names here, as strings. They can be
+# extensions coming with Sphinx (named 'sphinx.ext.*') or your custom
+# ones.
+extensions = ['sphinx.ext.autodoc',
+              'sphinx.ext.doctest',
+              'sphinx.ext.intersphinx',
+              'sphinx.ext.todo',
+              'sphinx.ext.coverage',
+              'sphinx.ext.mathjax',
+              'sphinx.ext.ifconfig',
+              'sphinx.ext.viewcode',
+              'sphinx.ext.githubpages']
+
+# Add any paths that contain templates here, relative to this directory.
+templates_path = ['_templates']
+
+# The suffix(es) of source filenames.
+# You can specify multiple suffix as a list of string:
+#
+# source_suffix = ['.rst', '.md']
+source_suffix = '.rst'
+
+# The master toctree document.
+master_doc = 'index'
+
+# General information about the project.
+project = 'gsmodutils'
+copyright = '2017, James Gilbert'
+author = 'James Gilbert'
+
+# The version info for the project you're documenting, acts as replacement for
+# |version| and |release|, also used in various other places throughout the
+# built documents.
+#
+# The short X.Y version.
+version = '0.0.1'
+# The full version, including alpha/beta/rc tags.
+release = '0.0.1'
+
+# The language for content autogenerated by Sphinx. Refer to documentation
+# for a list of supported languages.
+#
+# This is also used if you do content translation via gettext catalogs.
[…]

### 2017-08-24 gsmodutils/metacyc.py
Sphinx docs started for some of the utility software. Fixed code reformatting

@@ -1,9 +1,11 @@
-from cobra import Reaction, Metabolite
 import os
 from collections import defaultdict, Counter
 
+from cobra import Reaction, Metabolite
+
 
 def parse_db(db_path):
+    """ Parse metacyc dat files to build dict containing entries """
     compounds_path = os.path.join(db_path, 'compounds.dat')
     enzymes_path = os.path.join(db_path, 'enzrxns.dat')
     reactions_path = os.path.join(db_path, 'reactions.dat')
@@ -17,6 +19,12 @@ def parse_db(db_path):
 
 
 def parse_metacyc_file(fpath, unique_fields):
+    """
+    Parses a dat file
+    :str fpath: path to dat file
+    :list unique_fields: list of fields that there should only be a single item of in each entry
+    :return:
+    """
     db = dict()
 
     with open(fpath) as datfile:
@@ -61,8 +69,15 @@ def parse_metacyc_file(fpath, unique_fields):
 
 
 def add_pathway(model, enzyme_ids=None, reaction_ids=None, db_path=None, copy=False):
-    """For a given model add enzymes"""
-
+    """
+    For a given model add enzymes from metacyc database
+    :param model: cobra model object
+    :param enzyme_ids: list of EC entires in format ["EC-x.x.x.x", ...]
+    :param reaction_ids: set of reaction ids to add to model
+    :param db_path: path to the metacyc dat files on disk
+    :param copy:
+    :return:
+    """
     if copy:
         model = model.copy()
 
@@ -106,7 +121,12 @@ def add_pathway(model, enzyme_ids=None, reaction_ids=None, db_path=None, copy=Fa
 
 
 def get_enzyme_reactions(eid, db):
-    """For a given ec number return associated reaction ids"""
+    """
+    For a given ec number return associated reaction ids
+    :param eid: enzyme id format "EC-x.x.x.x"
+    :param db: database dict
+    :return:
+    """
     erids = []
 
     for rid, react in db['reactions'].items():
@@ -117,7 +137,13 @@ def get_enzyme_reactions(eid, db):
 
 
 def add_reaction(model, reaction_id, db):
-
+    """
+    Add a metacyc reaction id to a cobrapy model
+    :param model: cobrapy model instance
+    :param reaction_id: reaction identifier in metacyc
+    :param db: dictionary db object
+    :return: tuple(reaction, added_metabolites) cobrapy Reaction and Metabolite instances
+    """
     metabolites = dict()
     reaction = Reaction(reaction_id)
 

### 2017-08-24 gsmodutils/scrumpy.py

@@ -1,7 +1,9 @@
 from __future__ import print_function
-import os
+
 import argparse
 import json
+import os
+
 import cobra
 from cameo.core.utils import load_medium
 
@@ -11,8 +13,16 @@ def load_scrumpy_model(filepath, atpase_reaction="ATPase", atpase_flux=3.0,
     """
     Specify a base scrumpy structural model file and returns a cameo model.
     This hasn't be thoroughly tested so expect there to be bugs
-    
+
     To get a solution from the returned object you need to specify nice stuff like the atpase reaction and media
+
+    :param filepath:
+    :param atpase_reaction:
+    :param atpase_flux:
+    :param media:
+    :param objective_reactions:
+    :param obj_dir:
+    :return:
     """
 
     if objective_reactions is None:
@@ -85,10 +95,13 @@ def load_scrumpy_model(filepath, atpase_reaction="ATPase", atpase_flux=3.0,
 def get_tokens(line_dt):
     """
     Goes through each charachter in scrumpy file attempting to find tokens
-    
-    BUG: if there is a numeric after a direction token this fails
+
+    FIXME: if there is a numeric after a direction token this fails
     e.g. '->2 "PROTON"' fails but '-> 2 "PROTON"' works
+    :param line_dt:
+    :return:
     """
+
     tokens = []
     
     quoted = False
@@ -129,9 +142,13 @@ def get_tokens(line_dt):
     
 def parse_file(filepath, fp_stack=None, rel_path=''):
     """
-    Recursive function - takes in a scrumpy spy file and parses it, returning a set of reactions
-    
-    This code is not that stable. Expect bugs.
+     Recursive function - takes in a scrumpy spy file and parses it, returning a set of reactions
+
+    Note this code is not fully tested. Expect some bugs.
+    :param filepath:
+    :param fp_stack:
+    :param rel_path:
+    :return:
     """
     if fp_stack is None:
         fp_stack = []
@@ -245,7 +262,12 @@ def parse_file(filepath, fp_stack=None, rel_path=''):
 
 
 def scrumpy_to_cobra():
-    
+    """
+    Command line utility for parsing scrumpy files and creating cobrapy models
+    TODO: update to use click rather than argparse
+    TODO: update for better handling of import constraints
+    TODO: setting minimisation of sum of fluxes as an objective rather than biomass maximisation
+    """
     # Parser argument
     parser = argparse.ArgumentParser(description='parse a scrumpy file and output a json cobra compatable model')
    

### 2017-08-25 gsmodutils/__init__.py
Refactor of code to make it easier to understand. Still need to update docs.

@@ -1,5 +1,5 @@
 from __future__ import absolute_import, print_function
-from gsmodutils.project import GSMProject
-from gsmodutils.project_config import ProjectConfig
+from gsmodutils.project.project import GSMProject
+from gsmodutils.project.project_config import ProjectConfig
 
 __version__ = '0.0.1'

### 2017-08-25 gsmodutils/cli.py

@@ -14,7 +14,7 @@ from gsmodutils.model_diff import model_diff
 
 
 def load_project(project_path):
-    from gsmodutils.project import GSMProject
+    from gsmodutils import GSMProject
     project = None
     try:
         project = GSMProject(project_path)
@@ -206,7 +206,7 @@ def test(project_path, test_id, display_only, skip_default, verbose, log_path):
 @click.argument('model_path', type=str, default=None)
 def create_project(project_path, model_path):
     """Create a new gsmodutils project"""
-    from gsmodutils.project_config import ProjectConfig
+    from gsmodutils.project.project_config import ProjectConfig
     click.echo('Project creation {}'.format(project_path))
 
     click.echo('Using mode {}'.format(project_path))

### 2017-08-25 gsmodutils/project/project.py

@@ -9,11 +9,11 @@ from cobra.exceptions import Infeasible
 from lockfile import LockFile
 from six import string_types
 
-from gsmodutils.design import StrainDesign
 from gsmodutils.exceptions import ProjectNotFound, DesignError
 from gsmodutils.model_diff import model_diff
-from gsmodutils.project_config import ProjectConfig, default_project_file
-from gsmodutils.tester import GSMTester
+from gsmodutils.project.design import StrainDesign
+from gsmodutils.project.project_config import ProjectConfig, default_project_file
+from gsmodutils.test.tester import GSMTester
 
 
 class GSMProject(object):

### 2017-08-25 gsmodutils/project/project_config.py

@@ -8,7 +8,7 @@ import cobra
 
 import gsmodutils
 from gsmodutils.exceptions import ProjectConfigurationError
-from gsmodutils.validator import validate_model_file
+from gsmodutils.utils.validator import validate_model_file
 
 default_project_file = '.gsmod_project.json'
 default_model_conditionsfp = 'model_conditions.json'

### 2017-08-25 gsmodutils/test/tester.py

@@ -8,10 +8,10 @@ from io import StringIO
 
 from cobra.core import get_solution
 from cobra.exceptions import Infeasible
+from six import exec_
 
 import gsmodutils
-from gsmodutils.testutils import TestRecord
-from six import  exec_
+from gsmodutils.test.utils import TestRecord
 
 
 @contextlib.contextmanager
@@ -35,7 +35,7 @@ class GSMTester(object):
     def __init__(self, project):
         """Creates the storage locations for logs"""
         
-        if not isinstance(project, gsmodutils.project.GSMProject):
+        if not isinstance(project, gsmodutils.GSMProject):
             raise TypeError('Requires valid gsmodutils project')
             
         self.project = project

### 2017-08-25 setup.py

@@ -12,8 +12,7 @@ setup(
     packages=["gsmodutils"],
     entry_points={
         'console_scripts': [
-            'gsm_new_project=gsmodutils.gsm_project:main',
-            'scrumpy_to_cobra=gsmodutils.scrumpy:scrumpy_to_cobra',
+            'scrumpy_to_cobra=gsmodutils.utils.scrumpy:scrumpy_to_cobra',
             'gsmodutils=gsmodutils.cliutils:cli',
         ],
     },

### 2017-08-25 tests/helpers/tutils.py

@@ -1,11 +1,12 @@
-import cameo
-import cobra
-import tempfile
-from gsmodutils.project_config import ProjectConfig
-from gsmodutils.project_config import default_model_conditionsfp
-from gsmodutils.project import GSMProject
 import os
 import shutil
+import tempfile
+
+import cameo
+import cobra
+
+from gsmodutils import GSMProject
+from gsmodutils.project.project_config import ProjectConfig, default_model_conditionsfp
 
 
 class FakeProjectContext(object):

### 2017-08-25 tests/test_project.py

@@ -13,7 +13,7 @@ import cobra
 import pytest
 from tutils import FakeProjectContext
 
-from gsmodutils.project import GSMProject
+from gsmodutils import GSMProject
 
 
 def test_load_project():

### 2017-08-25 tests/test_project_config.py

@@ -4,17 +4,19 @@ Tests for GSMProject class
 TODO: tests for command line interface
 TODO: update use the fake project (written after these tests)
 """
-import pytest
-from gsmodutils.exceptions import ProjectConfigurationError
-from gsmodutils.project import GSMProject
-from gsmodutils.project_config import ProjectConfig
-from gsmodutils.project_config import default_model_conditionsfp, default_project_file
-import cameo
-import cobra
 import os
 import shutil
 import tempfile
+
+import cameo
+import cobra
+import pytest
 from tutils import FakeProjectContext
+
+from gsmodutils.exceptions import ProjectConfigurationError
+from gsmodutils import GSMProject
+from gsmodutils.project.project_config import ProjectConfig
+from gsmodutils.project.project_config import default_model_conditionsfp, default_project_file
 
 
 def project_creator(test_path, model=True, model_path=None):

### 2017-08-25 tests/test_tests.py

@@ -3,7 +3,7 @@ Test cases for the GSMTests class and associated files
 """
 
 from __future__ import print_function, absolute_import, division
-from gsmodutils.project import GSMProject
+from gsmodutils import GSMProject
 import json
 from tutils import FakeProjectContext
 import os
@@ -50,7 +50,7 @@ def test_py_tests():
     """
     
     code_str = """
-from gsmodutils.testutils import ModelTestSelector
+from gsmodutils.test.testutils import ModelTestSelector
 
 @ModelTestSelector(models=[], conditions=[], designs=[])
 def test_func(model, project, log):

### 2017-08-25 gsmodutils/__init__.py
Refactor of docs.

@@ -1,5 +1,5 @@
 from __future__ import absolute_import, print_function
-from gsmodutils.project.project import GSMProject
+from gsmodutils.project.interface import GSMProject
 from gsmodutils.project.project_config import ProjectConfig
 
 __version__ = '0.0.1'

### 2017-08-25 setup.py

@@ -13,7 +13,7 @@ setup(
     entry_points={
         'console_scripts': [
             'scrumpy_to_cobra=gsmodutils.utils.scrumpy:scrumpy_to_cobra',
-            'gsmodutils=gsmodutils.cliutils:cli',
+            'gsmodutils=gsmodutils.cli:cli',
         ],
     },
     setup_requires=['pytest-runner'],

### 2017-08-25 article/main.tex
Somee project tidy up
Some changes to the article

@@
 \title[test driven genome scale model development]{An integrated framework for test driven genome scale model development}
-\author[Sample \textit{et~al}.]{James P Gilbert\,$^{\text{\sfb 1,}}$, Nicole Pearcy\,$^{\text{\sfb 2}}$ and Jamie Twycross\,$^{\text{\sfb 2,}*}$}
+\author[Sample \textit{et~al}.]{James P Gilbert\,$^{\text{\sfb 1,}}$, Nicole Pearcy\,$^{\text{\sfb 1}}$ and Jamie Twycross\,$^{\text{\sfb 2,}*}$}
 \address{$^{\text{\sf 1}}$Synthetic Biology Research Centre, University of Nottingham, Nottingham, NG7 2RD, United Kingdom and \\
@@
 
-\abstract{\textbf{Motivation:} Genome scale metabolic models (GSMM) are increasingly important for systems biology and metabolic engineering research, capable of simulating complex steady state phenotopic behaviour.
+\abstract{\textbf{Motivation:} Genome scale metabolic models are increasingly important for systems biology and metabolic engineering research, capable of simulating complex steady state phenotopic behaviour.
 Models of this scale can include thousands of rections and metabolites, with many crucial pathways that only become activated in specific simulation settings.
 Despite their widespread use, power and the availability of tools to aid with the construction and analysis of large scale models, little methodology is suggested for the management of continually curated large scale models.\\
-\textbf{Results:} Using tools widely available in the software engineering community, we have developed the \textit{gsmodutils} framework.
-This system places an emphasis on test driven design of models in which a user makes small changes tracked in Distributed Version Control Systems (DVCS).
+\textbf{Results:} Based on python and cobrapy integration, we have developed the \textit{gsmodutils} modelling framework.
+This system places an emphasis on test driven design of models through easy to write unit testing.
 Crucially, different substrate conditions are configurable allowing users to examine how different designs or curations impact a wide range of system behaviours, hopefully, minimising error between model verions.
-We demonstrate a test of this framework on a standard \textit{E. coli} GSMM. \\
+A core feature of the software is to allow users to easily create sharable, self contained, docker images that can be run in cross platform envrionments.
 \textbf{Availability:} The software framework described within this paper is open source and freely available from http://github.com/<repolocation> \\
-\textbf{Contact:} \href{jamie.twycross@nottingham.ac.uk}{james.gilbert@nottingham.ac.uk}\\
+\textbf{Contact:} \href{jamie.twycross@nottingham.ac.uk}\\
 \textbf{Supplementary information:} Supplementary data are available at \textit{Bioinformatics} online.}
@@
 \section{Introduction}
-The reconstruction of genome scale metabolic models (GSMM) is an arduous process that follows a complex protocol to ensure model validity \cite{Palson}.
-Whilst autmated methods exists to construct GSMMs from reference genomes \cite{ModelSEED, ScrumPy}, .
+Constraints based modelling for biological systems has become a mainstay of systems biology for several decades.
+More recently, the structural approach has become a popular method for modelling steady state behaviour of large, metabolic networks.
+Such large scale reconstructions are often reffered to as genome scale models (GSMs), as the processes is significantly aided throught the advent of genome sequencing.
+
+The reconstruction of GSMs is an arduous process that follows a complex protocol to ensure model validity \cite{Palson}.
+Whilst many popular autmated methods exists to construct GSMs from reference genomes \cite{ModelSEED, ScrumPy}, there is a still a significant amount of manual curation.
 However, treating the creation of models as an isolated ``one-off'' event ignores the significant ammount of curation that is required for applications such as biotechnology.
@@
 
+Furthermore, the proliferation of tools available to users is significant.
+This covers huge variety of software platforms including the COBRA toolbox in Matlab \cite{},
+ScrumPy and CobraPy in Python \cite{ScrumPy, cobrapy} with additional tools and libraries such as cameo \cite{}, OptFlux in Java \cite{}, SurreyFBA \cite{}.
+Whilst most of these tools are Open Source and follow standards, such as SBML \cite{}, there are often few options to make analysis easy to replicate.
+
 In this paper we present a software framework geared toward \textit{test driven} genome scale model development, a concept that is taken directly from good software development practices.
@@
 Example test cases include the notion that flux must be carried through specific pathways, growth and uptake rates should match experimental evidence and different nutrient sources should be included.
-This paper details the features of the software, which is under continuous development.
-The reader is referred to the software user guide for a more detailed expoloration of software features.
+This article aims to summarise the main objectives of the gsmodutils software and we refer the reader to the software user guide for a more detailed expoloration features.
 
@@
 For simplicity, gsmodelutils uses the \textit{py.test}, which has a simple and intuitive interface.
-By automatically integrating the COBRAPy \cite{} and Cameo \cite{} users can easily write convenient test cases following examples given in the user guide.
+By automatically integrating the COBRAPy \cite{} users can easily write convenient test cases following examples given in the user guide.
 A standard test case, ensuring that a given model grows on media is given in Box \ref{}.
@@
  \item Integration with CobraPy and Cameo toolsets
- \item Integration with DVCS system (mercurial, git in a future release)
-  \item Optiopnally Integrate framework in a Docker container to stop ``runs on my machine'' issues with models
+ \item Use of flat files for ease of integration with DVCS systems such as git and mercurial
+ \item Easilly, integrate framework in a Docker container to stop ``runs on my machine'' issues with models
  \item Automated pre-commit tests upon model changes
  \item Convenient method for setting model media conditions to switch to growth on different substrates
- \item Ability to ``diff'' models by mapping to a common MetaNetX namespace
+ \item Ability to ``diff'' model versions to review what reactions have changed
  \item Storing model design differences to keep track of changes between, for example, production and wild type strains
@@
 % %\end{figure}
+
+\section{Related work}
+
+\textbf{cameo:} Cameo \cite{} is a set of strain design utilities, such as genetic algorithms for finding optimal gene knock-out
+strategies, as well as other flux balanace analysis utilities. Written in python and part of the cobrapy suite of applications,
+gsmodutils integrates easily with cameo.
+Indeed, making a set of design changes can easily be integrated into gsmoduitls through the python API.
+
+\textbf{KBase:} Designed to be a fully integrated utility for designing and applying genome scale models, KBase is
+a web based utility.
+The aims of this software are distinct from gsmodutils
+
 \section{Discussion}

### 2017-09-14 article/main.tex
article changes

@@
 
-\abstract{\textbf{Motivation:} Genome scale metabolic models are increasingly important for systems biology and metabolic engineering research, capable of simulating complex steady state phenotopic behaviour.
-Models of this scale can include thousands of rections and metabolites, with many crucial pathways that only become activated in specific simulation settings.
-Despite their widespread use, power and the availability of tools to aid with the construction and analysis of large scale models, little methodology is suggested for the management of continually curated large scale models.\\
-\textbf{Results:} Based on python and cobrapy integration, we have developed the \textit{gsmodutils} modelling framework.
-This system places an emphasis on test driven design of models through easy to write unit testing.
-Crucially, different substrate conditions are configurable allowing users to examine how different designs or curations impact a wide range of system behaviours, hopefully, minimising error between model verions.
-A core feature of the software is to allow users to easily create sharable, self contained, docker images that can be run in cross platform envrionments.
+\abstract{\textbf{Motivation:} Genome scale metabolic models are increasingly important for systems biology and metabolic engineering research, capable of simulating complex steady state phenotypic behaviour.
+Models of this scale can include thousands of reactions and metabolites, with many crucial pathways that only become activated in specific simulation settings.
+However, despite their widespread use, power and the availability of tools to aid with the construction and analysis of large scale models, little methodology is suggested for the management of continually curated large scale models.\\
+\textbf{Results:} Taking advantage of python and cobrapy integration, we have developed the \textit{gsmodutils} modelling framework.
+As a fully documented python API and series of command line utilities, gsmodutils places an emphasis on test driven design of models through easy to write test cases in json or python.
+Crucially, different substrate conditions are configurable allowing users to examine how different designs or curation impact a wide range of system behaviours, hopefully, minimising error between model versions.
+\\
+Furthermore, a core feature of the software is to allow users to easily create shareable, self contained, docker images that can be run in cross platform environments.\\
 \textbf{Availability:} The software framework described within this paper is open source and freely available from http://github.com/<repolocation> \\
-\textbf{Contact:} \href{jamie.twycross@nottingham.ac.uk}\\
+\textbf{Contact:} \href{mailto:jamie.twycross@nottingham.ac.uk}{jamie.twycross@nottingham.ac.uk}\\
 \textbf{Supplementary information:} Supplementary data are available at \textit{Bioinformatics} online.}
@@
 More recently, the structural approach has become a popular method for modelling steady state behaviour of large, metabolic networks.
-Such large scale reconstructions are often reffered to as genome scale models (GSMs), as the processes is significantly aided throught the advent of genome sequencing.
+Such large scale reconstructions are often refereed to as genome scale models (GSMs), as the processes is significantly aided throught the advent of genome sequencing.
 
-The reconstruction of GSMs is an arduous process that follows a complex protocol to ensure model validity \cite{Palson}.
-Whilst many popular autmated methods exists to construct GSMs from reference genomes \cite{ModelSEED, ScrumPy}, there is a still a significant amount of manual curation.
-However, treating the creation of models as an isolated ``one-off'' event ignores the significant ammount of curation that is required for applications such as biotechnology.
+The reconstruction of GSMs is an arduous process that follows a complex protocol to ensure model validity \cite{thiele2010protocol}.
+Whilst many popular automated methods exists to construct GSMs from reference genomes \cite{henry2010high, ScrumPy}, there is a still a significant amount of manual curation.
+However, treating the creation of models as an isolated ``one-off'' event ignores the significant amount of curation that is required for applications such as biotechnology.
 
-As a consequence, a significant amount of work has gone into the managament of genome scale models.
-The BiGG models database \cite{BiGG}, for example, exists to provide a standardised reposiotry of validated models that can be shared and reused.
-However, little focus is placed upon the colaborative design aspect of such models with few mechanisms existing to store the \textit{model delta} based tools provided in our software.
-Similiarly, the MetaNetX \cite{MetaNetX} system exists to provide a standardised namespace and toolchain for GSMM analysis.
-However, such tools often make it difficult to understand the design decisions made by the intial model authors.
+As a consequence, a significant amount of work has gone into the management of genome scale models.
+The BiGG models database \cite{king2015bigg}, for example, exists to provide a standardised repository of validated models that can be shared and reused.
+However, little focus is placed upon the collaborative design aspect of such models with few mechanisms existing to store the \textit{model delta} based tools provided in our software.
+Similarly, the MetaNetX \cite{moretti2016metanetx} system exists to provide a standardised namespace and toolchain for GSMM analysis.
+However, such tools often make it difficult to understand the design decisions made by the initial model authors.
 
 Furthermore, the proliferation of tools available to users is significant.
-This covers huge variety of software platforms including the COBRA toolbox in Matlab \cite{},
-ScrumPy and CobraPy in Python \cite{ScrumPy, cobrapy} with additional tools and libraries such as cameo \cite{}, OptFlux in Java \cite{}, SurreyFBA \cite{}.
-Whilst most of these tools are Open Source and follow standards, such as SBML \cite{}, there are often few options to make analysis easy to replicate.
+This covers huge variety of software platforms including the COBRA toolbox in Matlab \cite{schellenberger2011quantitative},
+ScrumPy and CobraPy in Python \cite{ScrumPy, ebrahim2013cobrapy} with additional tools and libraries such as cameo \cite{cardoso2017cameo}, OptFlux in Java \cite{rocha2010optflux}, SurreyFBA \cite{gevorgyan2010surreyfba}.
+Whilst most of these tools are Open Source and follow standards, such as SBML \cite{finney2003systems}, there are often few options to make analysis easy to replicate.
 
@@
 Example test cases include the notion that flux must be carried through specific pathways, growth and uptake rates should match experimental evidence and different nutrient sources should be included.
-This article aims to summarise the main objectives of the gsmodutils software and we refer the reader to the software user guide for a more detailed expoloration features.
+This article aims to summarise the main objectives of the gsmodutils software and we refer the reader to the software user guide for a more detailed exploration features.
 
@@
 \section{Software outline and features}
-Test driven development is driven by the idea of clearly defined test cases writen before significant changes are made to any underlying architecture.
+Test driven development is driven by the idea of clearly defined test cases written before significant changes are made to any underlying architecture.
 In the case of genome scale models, errors occur as a product of human curation to better represent newly discovered aspects of metabolism.
-For simplicity, gsmodelutils uses the \textit{py.test}, which has a simple and intuitive interface.
-By automatically integrating the COBRAPy \cite{} users can easily write convenient test cases following examples given in the user guide.
+For simplicity, gsmodutils uses the \textit{py.test}, which has a simple and intuitive interface.
+By automatically integrating the COBRAPy \cite{ebrahim2013cobrapy} users can easily write convenient test cases following examples given in the user guide.
 A standard test case, ensuring that a given model grows on media is given in Box \ref{}.
-When a new model repository is created with the gsmodelutils tool, a number of pre-written test cases are automatically added to a file.
-However, we stress that the vast majority of indiviudal use cases for a model must be specific to a given biological problem.
+When a new model repository is created with the gsmodutils tool, a number of pre-written test cases are automatically added to a file.
+However, we stress that the vast majority of individual use cases for a model must be specific to a given biological problem.
 
+Example python test case.
 \begin{verbatim*}
-def test_growth(model, media, l_bound, u_bound):
-    model.load_media(media)
-    solution = model.solve()
-    assert soltion.f < l_bound
-    assert soltion.f > u_bound
+def test_func(model, project, log):
+    status = model.solver.optimize()
+    log.assertion(status == 'optimal',
+                  "Works", # Success message
+                  "Does not work", # Failure message
+                  "Test description")
 \end{verbatim*}
@@
  \item Easy command line tool to integrate new models
- \item Unit testing with pytest
+ \item Simple testing with command line tool
  \item Integration with CobraPy and Cameo toolsets
  \item Use of flat files for ease of integration with DVCS systems such as git and mercurial
- \item Easilly, integrate framework in a Docker container to stop ``runs on my machine'' issues with models
+ \item Easily, integrate framework in a Docker container to stop ``runs on my machine'' issues with models
  \item Automated pre-commit tests upon model changes
@@
 \end{itemize}
+
+\subsection{Strain designs}
+A core aspect behind the implementation of gsmodutils is the concept of a \textit{design}, this encompasses a simple set of changes to a ``wild-type`` model that are required for analysis.
+For example a mutant strain that has undergone several gene deletions is not sufficiently different from the original wild-type to merit having its own genome scale model.
+However, it is often the case that such deletions are of scientific or industrial interest and, as such, the strain will be used in future work.
+Consequently, such designs are hereditary in nature.
+By taking the delta between the constraints applied to an initial model and subsequent modifications, gsmodutils allows users to easily reuse and export models with this design.
+As designs inherit from a base model, future curation to a wild type model will automatically be included in the designs.
+Similarly, designs are self contained and will not interfere with one another allowing a form of project management.
+
+\subsection{Development workflow}
+In this section we propose a method for the development of genome scale models that integrates gsmodutils with version control systems.
+TODO: a figure that demonstrates the design, test and fix cycle suggested for developers of genome scale models.
 
@@
 
-\textbf{cameo:} Cameo \cite{} is a set of strain design utilities, such as genetic algorithms for finding optimal gene knock-out
-strategies, as well as other flux balanace analysis utilities. Written in python and part of the cobrapy suite of applications,
+\textbf{cameo:} Cameo \cite{cardoso2017cameo} is a set of strain design utilities, such as genetic algorithms for finding optimal gene knock-out
+strategies, as well as other flux balance analysis utilities. Written in python and part of the cobrapy suite of applications,
 gsmodutils integrates easily with cameo.
-Indeed, making a set of design changes can easily be integrated into gsmoduitls through the python API.
+Indeed, making a set of design changes can easily be integrated into gsmodutils through the python API.
 
-\textbf{KBase:} Designed to be a fully integrated utility for designing and applying genome scale models, KBase is
-a web based utility.
-The aims of this software are distinct from gsmodutils
+...
 
@@
 When research projects end it all to is common for important large models to be published and become relics lost within the literature forgotten to all but the most dedicated of individuals.
-Furthermore, as GSMMs grow in terms of the metabolism the contain as well as the biological problems they are used to solve, problems with annotation and curation naturally accumulate as a product of human error.
-We have presented a framework with a number of features taken from the software development world specifically designed to improve colaboration and minimise such error.
+Furthermore, as GSMs grow in terms of the metabolism the contain as well as the biological problems they are used to solve, problems with annotation and curation naturally accumulate as a product of human error.
+We have presented a framework with a number of features taken from the software development world specifically designed to improve collaboration and minimise such error.
 However, it is important to stress the difference between defined behaviour expected from pre-written test cases and novel predictions made by a model.
 Indeed, a core objective of this framework is to ensure that good practices are followed in model development that help scientists to better trust the results discovered by their models.
-In an ideal world, we would envision a methodology such as our becoming a pre-requist for GSMMs to pass peer review.
+In an ideal world, we would envision a methodology such as our becoming a pre-requisite for GSMs to pass peer review.
+
+As with most software development projects, gsmodutils will see expanded features.
+Initially this will include tighter integration with version control systems such as git and mercurial.
+The ease of which one can create docker images from models also allows the potential for a
 
 \section*{Acknowledgements}
-We would like to thank the Oxford Brookes Cell Systems Modelling group for helpful discussions regarding this work and Rupert Norman for assitance testing and forming the idea of this software.
+We would like to thank the Oxford Brookes Cell Systems Modelling group for helpful discussions regarding this work and Rupert Norman for assistance testing and forming the idea of this software.
 

### 2017-10-09 article/article.tex
Article updated

--- 2017-09-14 article/main.tex
@@
-\documentclass{bioinfo}
-\copyrightyear{2017} \pubyear{2017}
+\documentclass[a4paper,11pt]{article}
+\usepackage[utf8]{inputenc}
+\usepackage{fullpage}
+\usepackage{color}
+\usepackage{authblk}
+\usepackage{listings}
+\usepackage{graphicx}
+\usepackage{subfigure}
 
-\access{Advance Access Publication Date: Day Month Year}
-\appnotes{Systems biology}
+\definecolor{codegreen}{rgb}{0,0.6,0}
+\definecolor{codegray}{rgb}{0.5,0.5,0.5}
+\definecolor{codepurple}{rgb}{0.58,0,0.82}
+\definecolor{backcolour}{rgb}{0.95,0.95,0.92}
+
+%opening
+% \address{$^{\text{\sf 1}}$Synthetic Biology Research Centre, University of Nottingham, Nottingham, NG7 2RD, United Kingdom and \\
+% $^{\text{\sf 2}}$Intelligent Modelling and Analysis group, School of Computer Science, University of Nottingham, Nottingham, NG7 2RD, United Kingdom}
+
+\title{Gsmodutils: A python based framework for test driven genome scale metabolic model development}
+\author[1]{James Gilbert}
+\author[1]{Nicole Pearcy}
+\author[2]{Jamie Twycross}
+\affil[1]{Synthetic Biology Research Centre, University of Nottingham, Nottingham, NG7 2RD, United Kingdom}
+\affil[2]{Intelligent Modelling and Analysis group, School of Computer Science, University of Nottingham, Nottingham, NG7 2RD, United Kingdom}
 
 \begin{document}
-\firstpage{1}
 
-\subtitle{Applications note}
+\maketitle
 
-\title[test driven genome scale model development]{An integrated framework for test driven genome scale model development}
-\author[Sample \textit{et~al}.]{James P Gilbert\,$^{\text{\sfb 1,}}$, Nicole Pearcy\,$^{\text{\sfb 1}}$ and Jamie Twycross\,$^{\text{\sfb 2,}*}$}
-\address{$^{\text{\sf 1}}$Synthetic Biology Research Centre, University of Nottingham, Nottingham, NG7 2RD, United Kingdom and \\
-$^{\text{\sf 2}}$Intelligent Modelling and Analysis group, School of Computer Science, University of Nottingham, Nottingham, NG7 2RD, United Kingdom}
-
-\corresp{$^\ast$To whom correspondence should be addressed.}
-
-\history{Received on XXXXX; revised on XXXXX; accepted on XXXXX}
-
-\editor{Associate Editor: XXXXXXX}
-
-\abstract{\textbf{Motivation:} Genome scale metabolic models are increasingly important for systems biology and metabolic engineering research, capable of simulating complex steady state phenotypic behaviour.
+\begin{abstract}
+\textbf{Motivation:} Genome scale metabolic models are increasingly important for systems biology and metabolic engineering research, capable of simulating complex steady state phenotypic behaviour.
 Models of this scale can include thousands of reactions and metabolites, with many crucial pathways that only become activated in specific simulation settings.
@@
 Furthermore, a core feature of the software is to allow users to easily create shareable, self contained, docker images that can be run in cross platform environments.\\
-\textbf{Availability:} The software framework described within this paper is open source and freely available from http://github.com/<repolocation> \\
-\textbf{Contact:} \href{mailto:jamie.twycross@nottingham.ac.uk}{jamie.twycross@nottingham.ac.uk}\\
-\textbf{Supplementary information:} Supplementary data are available at \textit{Bioinformatics} online.}
-
-\maketitle
+\textbf{Availability:} The software framework described within this paper is open source and freely available from http://github.com/gsmodutils \\
+\textbf{Contact:}{jamie.twycross@nottingham.ac.uk}
+\end{abstract}
 
@@
 The reconstruction of GSMs is an arduous process that follows a complex protocol to ensure model validity \cite{thiele2010protocol}.
-Whilst many popular automated methods exists to construct GSMs from reference genomes \cite{henry2010high, ScrumPy}, there is a still a significant amount of manual curation.
+Whilst many popular automated methods exists to construct GSMs from reference genomes \cite{henry2010high, poolman2006scrumpy}, there is a still a significant amount of manual curation.
 However, treating the creation of models as an isolated ``one-off'' event ignores the significant amount of curation that is required for applications such as biotechnology.
@@
 This covers huge variety of software platforms including the COBRA toolbox in Matlab \cite{schellenberger2011quantitative},
-ScrumPy and CobraPy in Python \cite{ScrumPy, ebrahim2013cobrapy} with additional tools and libraries such as cameo \cite{cardoso2017cameo}, OptFlux in Java \cite{rocha2010optflux}, SurreyFBA \cite{gevorgyan2010surreyfba}.
+ScrumPy and CobraPy in Python \cite{poolman2006scrumpy, ebrahim2013cobrapy} with additional tools and libraries such as cameo \cite{cardoso2017cameo}, OptFlux in Java \cite{rocha2010optflux}, SurreyFBA \cite{gevorgyan2010surreyfba}.
 Whilst most of these tools are Open Source and follow standards, such as SBML \cite{finney2003systems}, there are often few options to make analysis easy to replicate.
@@
 
-\section{Software outline and features}
+\section{Software}
+
+\subsection{Outline and features}
 Test driven development is driven by the idea of clearly defined test cases written before significant changes are made to any underlying architecture.
 In the case of genome scale models, errors occur as a product of human curation to better represent newly discovered aspects of metabolism.
-For simplicity, gsmodutils uses the \textit{py.test}, which has a simple and intuitive interface.
+
 By automatically integrating the COBRAPy \cite{ebrahim2013cobrapy} users can easily write convenient test cases following examples given in the user guide.
-A standard test case, ensuring that a given model grows on media is given in Box \ref{}.
+A standard test case, ensuring that a given model grows on media is given in Box \ref{fig:test_case}.
 When a new model repository is created with the gsmodutils tool, a number of pre-written test cases are automatically added to a file.
@@
 
-Example python test case.
-\begin{verbatim*}
+The software provides a number of features such as import an export of models in different formats and the generation of test reports through use of the command line. 
+The use of flat files for easy integration with version control software such as git and mercurial.
+
+In addition, projects are easy to export using portable standardised docker images \cite{merkel2014docker}. 
+The idea being to allow users to share models as quickly and easily as possible without concern for custom system configurations.
+
+\begin{figure}
+
+\lstset{language=Python}   
+\begin{lstlisting}
 def test_func(model, project, log):
@@
     log.assertion(status == 'optimal',
-                  "Works", # Success message
-                  "Does not work", # Failure message
-                  "Test description")
-\end{verbatim*}
-
-\begin{itemize}
- \item Easy command line tool to integrate new models
- \item Simple testing with command line tool
- \item Integration with CobraPy and Cameo toolsets
- \item Use of flat files for ease of integration with DVCS systems such as git and mercurial
- \item Easily, integrate framework in a Docker container to stop ``runs on my machine'' issues with models
- \item Automated pre-commit tests upon model changes
- \item Convenient method for setting model media conditions to switch to growth on different substrates
- \item Ability to ``diff'' model versions to review what reactions have changed
- \item Storing model design differences to keep track of changes between, for example, production and wild type strains
-\end{itemize}
+                  ``Works'', # Success message
+		  ``Does not work'', # Failure message
+                  ``Test description'')
+\end{lstlisting}
+\label{fig:test_case} 
+\caption{An example gsmodutils test case written in python}
+\end{figure}
 
 \subsection{Strain designs}
-A core aspect behind the implementation of gsmodutils is the concept of a \textit{design}, this encompasses a simple set of changes to a ``wild-type`` model that are required for analysis.
+A core aspect behind the implementation of gsmodutils is the concept of a \textit{design}, this encompasses a simple set of changes to a ``wild-type'' model that are required for analysis.
 For example a mutant strain that has undergone several gene deletions is not sufficiently different from the original wild-type to merit having its own genome scale model.
@@
 Similarly, designs are self contained and will not interfere with one another allowing a form of project management.
+Figure \ref{fig:strain_hered}
+
+\begin{figure}
+  \includegraphics[width=\textwidth]{inheritence.png}
+  \label{fig:strain_hered}
+  \caption{Examples of gsmodutils hierarchy for reuse and future compatibility of designs}
+\end{figure}
 
@@
 In this section we propose a method for the development of genome scale models that integrates gsmodutils with version control systems.
-TODO: a figure that demonstrates the design, test and fix cycle suggested for developers of genome scale models.
-
-% \begin{methods}
-% \section{Methods}
-% 
-% 
-% 
-% \begin{table}[!t]
-% \processtable{This is table caption\label{Tab:01}} {\begin{tabular}{@{}llll@{}}\toprule head1 &
-% head2 & head3 & head4\\\midrule
-% row1 & row1 & row1 & row1\\
-% row2 & row2 & row2 & row2\\
-% row3 & row3 & row3 & row3\\
-% row4 & row4 & row4 & row4\\\botrule
-% \end{tabular}}{This is a footnote}
-% \end{table}
-% 
-% \end{methods}
-% 
-% \begin{figure}[!tpb]%figure1
-% \fboxsep=0pt\colorbox{gray}{\begin{minipage}[t]{235pt} \vbox to 100pt{\vfill\hbox to
-% 235pt{\hfill\fontsize{24pt}{24pt}\selectfont FPO\hfill}\vfill}
-% \end{minipage}}
-% %\centerline{\includegraphics{fig01.eps}}
-% \caption{Caption, caption.}\label{fig:01}
-% \end{figure}
-% 
-% %\begin{figure}[!tpb]%figure2
-% %%\centerline{\includegraphics{fig02.eps}}
-% %\caption{Caption, caption.}\label{fig:02}
-% %\end{figure}
+Figure \ref{fig:test_case} highlights the notion of test cases, taken from test driven development.
+The outlined workflow is that the user writes a formal test case for some modelling goal, perhaps driven by captured experimental data, that fits a specific form of valdiation criteria.
+We note that, in principle, this test case should be written \textit{a priori} of changes to a model.
 
 \section{Related work}
-
 \textbf{cameo:} Cameo \cite{cardoso2017cameo} is a set of strain design utilities, such as genetic algorithms for finding optimal gene knock-out
@@
 Indeed, making a set of design changes can easily be integrated into gsmodutils through the python API.
-
-...
+\\
+\textbf{Model repositories:} models are frequently shared, at the time of publication through services such as BiGG \cite{king2015bigg} and Biomodels \cite{chelliah2013biomodels}. 
+Whilst these repositories encourage the reuse of models and the reproducibility of \textit{in silico} predictions they are not designed to improved collaboration.
+The software presented here is designed with the notion that genome scale models are never finished, \textit{per se}, but under continuous development.
+The cornerstone of this is the use of test cases, which formalise modelling validation criteria.
 
@@
 Initially this will include tighter integration with version control systems such as git and mercurial.
-The ease of which one can create docker images from models also allows the potential for a
+Furthermore, the objective of the project is to cultivate collaboration by simplifying the process of distributing large models to different users.
 
@@
 
-\bibliographystyle{plain}
+\bibliographystyle{unsrt}
 \bibliography{references}

### 2017-11-28 article/article.tex
Article major updates

@@
-\documentclass[a4paper,11pt]{article}
+\documentclass[a4paper,10pt]{article}
 \usepackage[utf8]{inputenc}
@@
 
-%opening
-% \address{$^{\text{\sf 1}}$Synthetic Biology Research Centre, University of Nottingham, Nottingham, NG7 2RD, United Kingdom and \\
-% $^{\text{\sf 2}}$Intelligent Modelling and Analysis group, School of Computer Science, University of Nottingham, Nottingham, NG7 2RD, United Kingdom}
-
 \title{Gsmodutils: A python based framework for test driven genome scale metabolic model development}
@@
 \begin{abstract}
-\textbf{Motivation:} Genome scale metabolic models are increasingly important for systems biology and metabolic engineering research, capable of simulating complex steady state phenotypic behaviour.
+\textbf{Motivation:} Genome scale metabolic models are increasingly important for systems biology and metabolic engineering research as they are capable of simulating complex steady state phenotypic behaviour.
 Models of this scale can include thousands of reactions and metabolites, with many crucial pathways that only become activated in specific simulation settings.
-However, despite their widespread use, power and the availability of tools to aid with the construction and analysis of large scale models, little methodology is suggested for the management of continually curated large scale models.\\
+However, despite their widespread use, power and the availability of tools to aid with the construction and analysis of large scale models, little methodology is suggested for the management of continually curated large scale models.
+This is quickly becoming an issue for large scale industrial biotechnology projects, which require good quality reusable models integral to the design, build and test cycle.
+\\
 \textbf{Results:} Taking advantage of python and cobrapy integration, we have developed the \textit{gsmodutils} modelling framework.
@@
 Crucially, different substrate conditions are configurable allowing users to examine how different designs or curation impact a wide range of system behaviours, hopefully, minimising error between model versions.
-\\
 Furthermore, a core feature of the software is to allow users to easily create shareable, self contained, docker images that can be run in cross platform environments.\\
 \textbf{Availability:} The software framework described within this paper is open source and freely available from http://github.com/gsmodutils \\
-\textbf{Contact:}{jamie.twycross@nottingham.ac.uk}
+\textbf{Contact:}{james.gilbert@nottingham.ac.uk}
 \end{abstract}
@@
 
+The archetype design, build and test cycle of synthetic biology heavily relies on the usage of bioinformatics software and modelling to improve the production of natural products \cite{carbonell2016bioinformatics}.
+In order to speed up the use of bioinformatics tools to produce high value platform chemicals, genome scale models are often used to discover methods for process optimisation \textit{in silico}.
+
+For example, many tools such as RetroPath \cite{carbonell2013retropath}, XTMS \cite{carbonell2014xtms} and GEM-Path \cite{campodonico2014generation} suggest thousands of potential hetrologous pathways as well as providing predictions for flux distributions.
+Many of these tools significantly increase the value of genome scale models, for example by coupling commodity production to an organism's growth \cite{feist2010model}.
+However, significant changes to wild-type strains must be tracked and compared to allow models to remain relevant.
+
+Similarly, many conventional applications of genome scale models in systems biology have often suffered from unnecessary replication of work due to a lack of standards.
+For example, there are now three independently developed models of \textit{Clostridium acetobutylicum} \cite{mcanulty2012genome, dash2014capturing, yoo2015quantitative}, an organism used in the production of solvents for over a century.
+These models all exist to solve similar biological problems, however there is disagreement over fundamental biochemical properties of these anaerobic organisms particularly with the focus on redox balancing \cite{dash2014capturing}.
+Indeed, such models also include updates based on improved genome annotations and the inclusion of fluxomic, transcriptomic and metabolomic characterisations \cite{yoo2015quantitative}.
+
+% Perhaps more worryingly, from an industrial perspective, is that the use of models in bioproduction has not escaped regulatory notice \cite{}.
+Furthermore, as models are often used to explain issues with online production and facilitate the optimisation of bioprocessing, it stands to reason that model correctness should attract oversight.
+Consequently, methods are needed to ensure that the models themselves are tightly version controlled.
+In this paper, we argue for the use of a form of \textit{test driven} model of development, a conceptual approach borrowed from software engineering.
+
 In this paper we present a software framework geared toward \textit{test driven} genome scale model development, a concept that is taken directly from good software development practices.
 By this we mean the notion that, as a model is curated to represent biological phenomena, much of the validation can be turned into specific test cases that can be repeated between model versions.
-Example test cases include the notion that flux must be carried through specific pathways, growth and uptake rates should match experimental evidence and different nutrient sources should be included.
+We provide an example test case for \textit{C.
+acetobutylicum} an organism that has had considerable focus in terms of genome scale models and how a working methodology using the software presented here can reduce repetition of work and improve the reproducibility of results.
 This article aims to summarise the main objectives of the gsmodutils software and we refer the reader to the software user guide for a more detailed exploration features.
 
-%\enlargethispage{12pt}
+\section{Improving the design phase of industrial biotechnology}
+
+The design phase of industrial synthetic biotechnology, due to historical and regulatory convention, often follows a conventional waterfall model for development of genome scale models.
+This is borne out of a number significant constraints in the development of genome scale models.
+The waterfall approach for model is characterised in Figure \ref{fig:methodology_approaches} where requirements for a model are captured at an early stage and are hard to refine during the production process.
+
+In contrast, an \textit{agile} methodology for the development of models places the focus on adapting work to new requirements \cite{martin2002agile}.
+Such an approach best fits genome scale models because they are rarely created to investigate individual processes and, instead, capture the complexity of large systems.
+This modelling formation costs an in depth understanding of dynamic behaviour. 
+However, capturing steady state phenomena still provides a good understanding of systemic properties \cite{o2015using}.
+
+Constraints based models are intrinsically related to available genome annotations.
+Such annotations rely heavily on automated matches to related species, with the characterisations of individual genes often being left to a few of critical interest \cite{seemann2014prokka}.
+
+Such approaches often leave models with missing reactions \cite{}, incorrect gene-reaction rules \cite{} or with pathways based on gap filling methods that add reactions that may not be present within the organism \cite{benedict2014likelihood}.
+Genome annotation is frequently updated and models are often corrected ``on the fly'' when attempting to understand specific natural phenomena.
+
+\begin{figure}[ht]
+\includegraphics[width=\textwidth]{Waterfall.png}
+\caption{Conventional waterfall methodology for model development.
+Each process is considered to be an isolated aspect of model development.}
+\label{fig:methodology_approaches}
+\end{figure}
+
+Consequently, models undergo significant manual annotation which can lead to errors.
+In this work we advocate a test driven approach to model development highlighted in Figure \ref{fig:test_driven}.
+Here, the model is changed to achieve core research goals that are dynamic in response to the changes of a project.
+In order to achieve this goal, validation criteria for a model such as growth conditions or the impact of gene knock outs, should be formally set.
+When a model is changed, all such validation criteria should be retested to ensure that models do not regress to previous states.
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=0.5\textwidth]{Testdriven.png}
+\caption{Workflow for model development proposed for gsmodutils.
+The objective of this approach is to simultaneously capture research questions, model validation criteria and minimise the impact of changes on previously completed models.
+}
+\label{fig:test_driven}
+\end{figure}
+
+We feel that while many of the current software tools for genome scale models do an excellent job of facilitating answers to crucial research and design questions, their is a significant gap in terms of the reliability and re-usability of models.
+The following sections provide an overview of the gsmodutils software framework.
+Gsmodutils aims to provide a basis for test driven, version controlled agile model development.
+All software is open source and is designed to be interoperable with platforms widely used in the domain of constraints based modelling.
 
@@
 
-\begin{figure}
+\begin{figure}[ht]
 
@@
 Figure \ref{fig:test_case} highlights the notion of test cases, taken from test driven development.
-The outlined workflow is that the user writes a formal test case for some modelling goal, perhaps driven by captured experimental data, that fits a specific form of valdiation criteria.
+The outlined workflow is that the user writes a formal test case for some modelling goal, perhaps driven by captured experimental data, that fits a specific form of validation criteria.
 We note that, in principle, this test case should be written \textit{a priori} of changes to a model.
+
+\section{Example usage: Clostridium acetobutylicum models}
+
+% Scientific background
+
+% Prodblems models solve
+
+% Created designs in gsmodutils - start with original model, make changes
+
+% Diffs between existing models
+
+% Conclusions
 
@@
 Furthermore, the objective of the project is to cultivate collaboration by simplifying the process of distributing large models to different users.
+As a consequence, methods for sharing docker containers will be investigated.
 
 \section*{Acknowledgements}
-We would like to thank the Oxford Brookes Cell Systems Modelling group for helpful discussions regarding this work and Rupert Norman for assistance testing and forming the idea of this software.
+We would like to thank the Oxford Brookes Cell Systems Modelling group for helpful discussions regarding this work, Rupert Norman for assistance testing and forming the idea of this software and Thomas Millat for crucial scientific advice.
 

### 2017-11-29 article/article.tex
Added references and updated figure

@@
 \usepackage{subfigure}
+
+%PREPRINT/MANUSCRIPT OPTIONS
+\usepackage{setspace}
+\usepackage{lineno}
+\linenumbers
+\doublespacing
 
@@
 \textbf{Motivation:} Genome scale metabolic models are increasingly important for systems biology and metabolic engineering research as they are capable of simulating complex steady state phenotypic behaviour.
-Models of this scale can include thousands of reactions and metabolites, with many crucial pathways that only become activated in specific simulation settings.
-However, despite their widespread use, power and the availability of tools to aid with the construction and analysis of large scale models, little methodology is suggested for the management of continually curated large scale models.
-This is quickly becoming an issue for large scale industrial biotechnology projects, which require good quality reusable models integral to the design, build and test cycle.
+Constraints based models of this form can include thousands of reactions and metabolites, with many crucial pathways that only become activated in specific simulation settings.
+However, despite their widespread use, power and the availability of tools to aid with the construction and analysis of large scale models little methodology is suggested for the management of continually curated large scale models.
+This is quickly becoming an issue for industrial systems and synthetic biotechnology applications, which require good quality reusable models integral to the design, build and test cycle.
 \\
@@
 \section{Introduction}
-Constraints based modelling for biological systems has become a mainstay of systems biology for several decades.
-More recently, the structural approach has become a popular method for modelling steady state behaviour of large, metabolic networks.
-Such large scale reconstructions are often refereed to as genome scale models (GSMs), as the processes is significantly aided throught the advent of genome sequencing.
+Stoichiometric constraints based modelling for biological systems has been a mainstay of systems biology for several decades \cite{fell1986fat, varma1994stoichiometric}.
+Given its flexibility, low barrier to entry and requirement only on minimal knowledge regarding the stoichiometry of metabolic networks this structural approach has become an extremely popular method for modelling steady state behaviour of large, biochemical networks \cite{kauffman2003advances}.
+Such large scale reconstructions are often refereed to as genome scale models (GSMs), as the processes is significantly aided through the advent of relatively inexpensive genome sequencing \cite{o2015using, land2015insights}.
 
@@
 However, little focus is placed upon the collaborative design aspect of such models with few mechanisms existing to store the \textit{model delta} based tools provided in our software.
-Similarly, the MetaNetX \cite{moretti2016metanetx} system exists to provide a standardised namespace and toolchain for GSMM analysis.
+Similarly, the MetaNetX \cite{moretti2016metanetx} system exists to provide a standardised namespace and toolchain for GSM analysis.
 However, such tools often make it difficult to understand the design decisions made by the initial model authors.
 
-Furthermore, the proliferation of tools available to users is significant.
-This covers huge variety of software platforms including the COBRA toolbox in Matlab \cite{schellenberger2011quantitative},
-ScrumPy and CobraPy in Python \cite{poolman2006scrumpy, ebrahim2013cobrapy} with additional tools and libraries such as cameo \cite{cardoso2017cameo}, OptFlux in Java \cite{rocha2010optflux}, SurreyFBA \cite{gevorgyan2010surreyfba}.
-Whilst most of these tools are Open Source and follow standards, such as SBML \cite{finney2003systems}, there are often few options to make analysis easy to replicate.
+Furthermore, as with many areas of bioinformatic study the proliferation of available computational tools has become somewhat vast.
+This covers huge variety of software platforms including the COBRA toolbox in Matlab \cite{schellenberger2011quantitative}, ScrumPy and CobraPy in Python \cite{poolman2006scrumpy, ebrahim2013cobrapy} with additional tools and libraries such as cameo \cite{cardoso2017cameo}, OptFlux in Java \cite{rocha2010optflux}, SurreyFBA \cite{gevorgyan2010surreyfba}.
+Whilst most of these tools are Open Source and follow standards, such as SBML \cite{finney2003systems}, there are often few options that make analysis easy to replicate.
 
@@
 Many of these tools significantly increase the value of genome scale models, for example by coupling commodity production to an organism's growth \cite{feist2010model}.
-However, significant changes to wild-type strains must be tracked and compared to allow models to remain relevant.
+These tools all suggest major changes to wild-type strains must be tracked and compared to allow models to remain relevant.
+In effect, mechanisms are required to related modified models to modified test and production strains.
 
 Similarly, many conventional applications of genome scale models in systems biology have often suffered from unnecessary replication of work due to a lack of standards.
-For example, there are now three independently developed models of \textit{Clostridium acetobutylicum} \cite{mcanulty2012genome, dash2014capturing, yoo2015quantitative}, an organism used in the production of solvents for over a century.
-These models all exist to solve similar biological problems, however there is disagreement over fundamental biochemical properties of these anaerobic organisms particularly with the focus on redox balancing \cite{dash2014capturing}.
+For example, there are now many independently developed models of \textit{Clostridium acetobutylicum} \cite{lee2008genome, senger2008genome, mcanulty2012genome, dash2014capturing, yoo2015quantitative}, an organism used in the production of solvents for over a century.
+These models all exist to solve similar biological problems some being updates to the inital base models. 
+However, there has been disagreement over fundamental biochemical properties of these anaerobic organisms, notably with the focus on redox balancing \cite{dash2014capturing}.
 Indeed, such models also include updates based on improved genome annotations and the inclusion of fluxomic, transcriptomic and metabolomic characterisations \cite{yoo2015quantitative}.
+
+Many of the results presented in original sources are difficult to compare due to a lack of standardisation of reference identifiers for reaction names.
+Where models are shared they are often in non-standard spreadsheet formats, rather than SBML models.
+Indeed, even in the case of valid SBML models being made available at the time of publication few details are given as to how to run such models.
 
@@
 In this paper, we argue for the use of a form of \textit{test driven} model of development, a conceptual approach borrowed from software engineering.
+Given the availability of widespread open source virtualisation tools that allow the standardisation of platforms \cite{merkel2014docker}, modellers have the opportunity to share highly reproducible research \cite{chamberlain2014using}.
 
@@
 
-The design phase of industrial synthetic biotechnology, due to historical and regulatory convention, often follows a conventional waterfall model for development of genome scale models.
+The design phase of industrial synthetic biotechnology, due to historical and regulatory convention, often follows a waterfall model for development of genome scale models.
 This is borne out of a number significant constraints in the development of genome scale models.
@@
 Such an approach best fits genome scale models because they are rarely created to investigate individual processes and, instead, capture the complexity of large systems.
+Constraints based models are intrinsically related to available genome annotations.
+Such annotations rely heavily on automated matches to related species, with the characterisations of individual genes often being left to a few of critical interest \cite{seemann2014prokka}.
 This modelling formation costs an in depth understanding of dynamic behaviour. 
@@
 
-Constraints based models are intrinsically related to available genome annotations.
-Such annotations rely heavily on automated matches to related species, with the characterisations of individual genes often being left to a few of critical interest \cite{seemann2014prokka}.
-
-Such approaches often leave models with missing reactions \cite{}, incorrect gene-reaction rules \cite{} or with pathways based on gap filling methods that add reactions that may not be present within the organism \cite{benedict2014likelihood}.
-Genome annotation is frequently updated and models are often corrected ``on the fly'' when attempting to understand specific natural phenomena.
+As such approaches often leave models with missing reactions, incorrect gene-reaction rules \cite{thiele2010protocol} or with pathways based on gap filling methods that add reactions that may not be present within the organism \cite{benedict2014likelihood}.
+When attempting to understand specific natural phenomena, genome annotations are frequently updated and models are often corrected in an \textit{ad hoc} manner.
 
@@
 
-Consequently, models undergo significant manual annotation which can lead to errors.
+Therefore, models undergo significant manual annotation - a process which naturally leads to errors.
 In this work we advocate a test driven approach to model development highlighted in Figure \ref{fig:test_driven}.
 Here, the model is changed to achieve core research goals that are dynamic in response to the changes of a project.
-In order to achieve this goal, validation criteria for a model such as growth conditions or the impact of gene knock outs, should be formally set.
+In order to meet this objective, validation criteria for a model such as growth conditions or the impact of gene knock outs, should be formally set.
 When a model is changed, all such validation criteria should be retested to ensure that models do not regress to previous states.
@@
 
-We feel that while many of the current software tools for genome scale models do an excellent job of facilitating answers to crucial research and design questions, their is a significant gap in terms of the reliability and re-usability of models.
+We feel that while many of the current software tools for genome scale models do an excellent job of facilitating answers to crucial research and design questions.
+However, their is a major gap in terms of the reliability and re-usability of models due to a lack of standardisation and software tools to aid such processes.
 The following sections provide an overview of the gsmodutils software framework.
@@
   \label{fig:strain_hered}
-  \caption{Examples of gsmodutils hierarchy for reuse and future compatibility of designs}
+  \caption{Examples of gsmodutils design inheritance.
+  Each design stores the delta between the wild-type base model, any parents and the changes to constraints the design contains.
+  In the example presented above a hetrologous production pathway is combined with a reusable set of knock-outs.
+  In practice, ideally, these designs should relate to real constructs and strains evaluated by wet lab biologists.
+  }
 \end{figure}
@@
 \section{Discussion}
-When research projects end it all to is common for important large models to be published and become relics lost within the literature forgotten to all but the most dedicated of individuals.
-Furthermore, as GSMs grow in terms of the metabolism the contain as well as the biological problems they are used to solve, problems with annotation and curation naturally accumulate as a product of human error.
+In order to facilitate the sharing and dissemination of research good standards and software are required \cite{}.
+Naturally a great deal of effort has gone in to producing high quality systems and synthetic biology standards \cite{}.
+Furthermore, when research projects end it all to is common for important large models to be published and become relics lost within the literature forgotten to all but the most dedicated of individuals.
+As GSMs grow in terms of the metabolism the contain as well as the biological problems they are used to solve, problems with annotation and curation naturally accumulate as a product of human error.
+Software that facilitates actively improving how researchers develop and apply models to new phenomena is required.
+
 We have presented a framework with a number of features taken from the software development world specifically designed to improve collaboration and minimise such error.

### 2017-11-29 gsmodutils/cli.py
Updates to docs

@@ -272,7 +272,11 @@ def diff(model_path, base_model, project_path, parent, output, names):
 @click.option('--project_path', default='.', help='gsmodutils project path')
 @click.option('--validate/--no-validate', default=True, help='Chose to validate the model before it is added.')
 def add_model(path, project_path, validate):
-    """Add a model to a specified gsm project"""
+    """
+    Add a model to a specified gsm project.
+    If validation is selected, where the model fails to conform to cobra standards, the model will not be added to the
+    project.
+    """
     project = _load_project(project_path)
     try:
         project.add_model(path, validate=validate)

### 2017-11-29 gsmodutils/utils/metacyc.py
updated metacyc parser to use correct encoding fixes python 3, uses new context manager for python2 and 3 compatability.

@@ -3,6 +3,7 @@ from collections import defaultdict, Counter
 from gsmodutils.utils.io import load_model
 from cobra import Reaction, Metabolite, Model
 from cobra.io import save_json_model
+import sys
 
 
 def parse_db(db_path):
@@ -19,6 +20,30 @@ def parse_db(db_path):
     return database
 
 
+class FileEncodingCtx(object):
+
+    def __init__(self, filename, encoding='latin-1', **kwargs):
+        """
+        Python 2 and 3 context manager for opening latin-1 encoded files
+        Parsing Metacyc .dat files breaks opening files in python 3
+        """
+        self.filename = filename
+        self.encoding = encoding
+        self.kwargs = kwargs
+
+    def __enter__(self):
+        if sys.version_info[0] < 3:
+            self.open_file = open(self.filename, **self.kwargs)
+        else:
+            self.open_file = open(self.filename, encoding=self.encoding, **self.kwargs)
+
+        return self.open_file
+
+    def __exit__(self, exec_type, exec_value, traceback):
+        self.open_file.close()
+        return self.open_file
+
+
 def parse_metacyc_file(fpath, unique_fields):
     """
     Parses a dat file
@@ -28,7 +53,7 @@ def parse_metacyc_file(fpath, unique_fields):
     """
     db = dict()
 
-    with open(fpath) as datfile:
+    with FileEncodingCtx(fpath) as datfile:
         entry = defaultdict(list)
         lprev = None
         for line in datfile:

### 2017-11-29 article/article.tex
Expanded article to cover c aceto examples

@@
 
-\section{Example usage: Clostridium acetobutylicum models}
+\section{Example usage Clostridium acetobutylicum models}
+\subsection{Scientific background}
+\textit{C. acetobutylicum} ATCC 824 is a widely researched organism that has found use in the production of solvents for decades.
+More recently, \textit{C. acetobutylicum} has been applied for the production of alcohols such as ethanol and n-butanols.
+The industrial relevance and strong background of experimental data has led to the application of many genome scale models \cite{}.
+Whilst increasingly informative when combined with experimental data, the history of the management of models for \textit{C. acetobutylicum} tells a different story.
 
-% Scientific background
+Initially, two rival models were published containing a similar number of genes \cite{lee2008genome, senger2008genome}.
+% \textbf{TODO: describe in more detail}.
+%Discuss limitations of rival models
+This led to the further development of subsequent models, iCaC 490 \cite{mcanulty2012genome} and an automatically generated model from  the Model SEED  pipeline \cite{}.
 
-% Prodblems models solve
+In this example we focus on the more recent contributions of iCac802 \cite{dash2014capturing} and iCac967 \cite{yoo2015quantitative}.
+iCac802 combines a significantly updated set of annotations from previously curated model versions, an improved TCA cycle, full proton balancing in reactions and reactions for more complete coverage of hexadecanoyl-acp, hexadecanoyl-CoA, crotonyl-CoA pathways as well as pathways for the uptake and biosynthesis of more compounds.
+Crucially, the model itself received significant validation against fluxonomic data from 13C-metabolic flux analysis ($13$C-MFA).
+This additional validation set considerable flux bounds upon pathways, leading to improved curation.
 
-% Created designs in gsmodutils - start with original model, make changes
+Whilst iCac802 was informed by previous models, critically, it was started as a new curation project, a significant undertaking of work.
+Such an approach may be necessary to uncover errors or oversights in construction processes.
+However, this naturally adds a significant time burden for model curators and may be due to a lack of adaptability from previous models.
 
-% Diffs between existing models
+The iCac 967 model presents a smaller set of changes to the annotation of the \textit{C. acetobutylicum} model.
+While the number of genes increased from 802 to 967 manual annotation actually reduced the number of reactions and metabolites from 1462 to 1137 and 1231 to 1058, respectively.
+The annotation ane experimental evidence from iCac802 \cite{dash2014capturing} to iCac967 \cite{yoo2015quantitative} is greatly expanded.
+However, it is not entirely clear, from the initial publication, the extent to which the expanded iCaC907 model preserves crucial experimental validation.
+The following sections aim to check the later model against previous validation criteria.
+
+\subsection{Model integration}
+\textbf{Project construction:}
+An initial project of the \textit{C. acetobutylicum} model was created in gsmodutils.
+This is based around the iCac802 base model.
+Included are also knockout strains $\Delta adc, \Delta ack, \Delta ptb$ and $\Delta hbd$ (TODO: describe knockouts).
+
+In addition, a design listing the constraints changed in model version iCaC907 
+\footnote{We would like to note that such a significant change to a model, in many practical cases, would not be considered a design but, rather, a separate model.
+Separate models within the same project is a feature supported by gsmodutils.}.
+\\
+\textbf{Model comparison: TODO Diffs between existing models - use (and further develop) gsmodutils diff for model stats}.
+
+\subsection{Evaluation of model validation criteria}
+Several crucial validation criteria were identified from the experimental and theoretical analysis of the models found in the original publications.
+We note that this is not an exhaustive set of criteria.
+This section is designed to serve as a use case for the gsmodutils framework, not to comprehensively review the initial publications.
+\\
+\textbf{TCA cycle}\\
+\textbf{Proton balanced results}\\
+\textbf{Predictions of knockouts}\\
+\textbf{Flux validation against 13C data}\\
+\textbf{Under acidogenic stress (H inhibited conditions) } \\
+\textbf{Solvent formation Cell recycling in solventogenic phase}\\
 
 % Conclusions
+\subsection{Remarks} 
+\textbf{To fill in when validation is complete.}
 
@@
 
+% Furthermore, several methods for the automated construction of gsms exist.
+% However, to our knowledge no methods include a fully automated pipeline for updating existing manually created models once constructed.
+% We would like to note that the collection of data for the validation of models and the construction of tests could be used to allow automated updates to take place.
+% For example, as GPR are curated models could automatically be updated with the inclusion of new information whilst being validated against existing criteria.
+
 As with most software development projects, gsmodutils will see expanded features.

### 2017-12-11 article/article.tex
minor changes to article

@@
 The annotation ane experimental evidence from iCac802 \cite{dash2014capturing} to iCac967 \cite{yoo2015quantitative} is greatly expanded.
-However, it is not entirely clear, from the initial publication, the extent to which the expanded iCaC907 model preserves crucial experimental validation.
+However, it is not entirely clear, from the initial publication, the extent to which the expanded iCaC967 model preserves crucial experimental validation.
 The following sections aim to check the later model against previous validation criteria.
@@
 
-In addition, a design listing the constraints changed in model version iCaC907 
+In addition, a design listing the constraints changed in model version iCaC967 
 \footnote{We would like to note that such a significant change to a model, in many practical cases, would not be considered a design but, rather, a separate model.
@@
 \section*{Acknowledgements}
-We would like to thank the Oxford Brookes Cell Systems Modelling group for helpful discussions regarding this work, Rupert Norman for assistance testing and forming the idea of this software and Thomas Millat for crucial scientific advice.
+We would like to thank the Oxford Brookes Cell Systems Modelling group for helpful discussions regarding this work, Rupert Norman for assistance testing and forming the idea of this software, Thomas Millat for crucial scientific advice and Phillipe Soucaille for assistance reconstructing the iCac967
 

### 2017-12-19 gsmodutils/cli.py
Changes to docker command to user dockerpy to build project

@@ -427,7 +427,9 @@ Tests directory - {tests_dir}
 @click.command()
 @click.option('--project_path', default='.', help='gsmodutils project path')
 @click.option('--overwrite/--no-overwrite', default=False, help='overwrite existing dockerfile')
-def docker(project_path, overwrite):
+@click.option('--build/--no-build', default=False, help='build docker container')
+@click.option('--tag', default=None, help='tag name for docker container (appended to project name).')
+def docker(project_path, overwrite, build, tag):
     """ Create a dockerfile for the project """
     project = _load_project(project_path)
     if os.path.exists(os.path.join(project_path, 'Dockerfile')) and not overwrite:
@@ -436,10 +438,40 @@ def docker(project_path, overwrite):
         project.config.create_docker_file(project_path)
         click.echo('**** Created Dockerfile **** \n')
 
-    click.echo('''Use "docker build -t=\'container_name\'" to build a container.
-    and use "docker save container_name -o container_name" to create a shareable image''')
+    # Check to see if docker is installed on the machine
+    import docker
+    client = docker.from_env()
+    try:
+        client.info()
+    except ConnectionError:
+        click.echo(
+            click.style('Error: Docker is either not installed or not configured on your system'
+                        'please consult the documentation at docs.docker.com', fg='red')
+        )
+        exit(-1)
 
-    click.echo('\n\nIf docker is not configured on your system please consult the documentation at docs.docker.com')
+    if not build:
+        click.echo('Build option not specified use "gsmodutils docker --build"'
+                   'or "docker build -t=\'project_name\'" to build a container.')
+    else:
+        tag = project.config.name
+        if tag is not None:
+            tag += "-" + tag
+        click.echo('Running docker build, this may take some time...')
+        try:
+            client.image.build(path=project.project_path, tag=tag)
+            click.echo('Image built')
+            click.echo('Use docker save container_name -o container_name" to create a shareable image')
+
+        except docker.errors.BuildError as e:
+            click.echo(
+                click.style('Error building project:\n{}'.format(e), fg='red')
+            )
+
+        except docker.errors.APIError as e:
+            click.echo(
+                click.style('Docker returned an error:\n{}'.format(e), fg='red')
+            )
 
 
 cli.add_command(test)

### 2017-12-20 article/article.tex
Article additions

@@
 \\
+
 \textbf{TCA cycle}\\
@@
 \textbf{Predictions of knockouts}\\
+The inital publication for iCAC802 included validation of four gene knockout mutant strains.
+$\Delta ack$ acetate kinase.\\
+$\Delta ptb$ phosphotransbutrylase.\\
+$\Delta adc$ acetoacetate decarboxylase. \textbf{This reaction is present and the knock out works as expected but not annotated correctly.}\\
+$\Delta hdb$ hydroxybutyryl-CoA dehydrogenase. \\
+
 \textbf{Flux validation against 13C data}\\
@@
 
-% Conclusions
+% Discuss these results
 \subsection{Remarks} 
-\textbf{To fill in when validation is complete.}
+Reproducing the results of the initial publication of the iCAC802 model \cite{dash2014capturing} was possible.
+However, it is notworthy to point out that there were a number of inconsistencies between the annotation in the published model and the publication.
+Notably, the gene ``adc'', which codes for acetoacetate decarboxylase was missing from the initial annotation.
+We would stress that these did not significantly impact the results of the model or the claims made in the publication.
 

### 2017-12-20 gsmodutils/cli.py
Docker now allows the saving of images, updated requirements template

@@ -428,8 +428,10 @@ Tests directory - {tests_dir}
 @click.option('--project_path', default='.', help='gsmodutils project path')
 @click.option('--overwrite/--no-overwrite', default=False, help='overwrite existing dockerfile')
 @click.option('--build/--no-build', default=False, help='build docker container')
+@click.option('--save/--no-save', default=False, help='save docker image of tagged container')
 @click.option('--tag', default=None, help='tag name for docker container (appended to project name).')
-def docker(project_path, overwrite, build, tag):
+@click.option('--save_path', default=None, help='Save path for shared docker image')
+def docker(project_path, overwrite, build, save, tag, save_path):
     """ Create a dockerfile for the project """
     project = _load_project(project_path)
     if os.path.exists(os.path.join(project_path, 'Dockerfile')) and not overwrite:
@@ -450,18 +452,18 @@ def docker(project_path, overwrite, build, tag):
         )
         exit(-1)
 
+    tag = project.config.name
+    if tag is not None:
+        tag += ":" + tag
+
     if not build:
         click.echo('Build option not specified use "gsmodutils docker --build"'
-                   'or "docker build -t=\'project_name\'" to build a container.')
+                   'or "docker build -t=\'{}\'" to build a container.'.format(project.config.name))
     else:
-        tag = project.config.name
-        if tag is not None:
-            tag += "-" + tag
         click.echo('Running docker build, this may take some time...')
         try:
-            client.image.build(path=project.project_path, tag=tag)
+            client.images.build(path=project.project_path, tag=tag)
             click.echo('Image built')
-            click.echo('Use docker save container_name -o container_name" to create a shareable image')
 
         except docker.errors.BuildError as e:
             click.echo(
@@ -472,6 +474,39 @@ def docker(project_path, overwrite, build, tag):
             click.echo(
                 click.style('Docker returned an error:\n{}'.format(e), fg='red')
             )
+            exit(-1)
+
+    if save:
+        # ensure that build has been completed first
+        try:
+            image = client.images.get(tag)
+        except docker.errors.ImageNotFound:
+            click.echo(
+                click.style('Image not found try building image first', fg='red')
+            )
+            exit(-1)
+        except docker.errors.APIError as e:
+            click.echo(
+                click.style('Docker returned an error:\n{}'.format(e), fg='red')
+            )
+            exit(-1)
+
+        if save_path is None:
[…]

### 2017-12-20 gsmodutils/cli.py
Forces lowercase tag name

@@ -456,6 +456,8 @@ def docker(project_path, overwrite, build, save, tag, save_path):
     if tag is not None:
         tag += ":" + tag
 
+    tag = tag.lower()
+
     if not build:
         click.echo('Build option not specified use "gsmodutils docker --build"'
                    'or "docker build -t=\'{}\'" to build a container.'.format(project.config.name))

### 2018-02-26 gsmodutils/project/interface.py
Small changes to fix a few bugs

@@ -312,16 +312,17 @@ class GSMProject(object):
         :param copy: return copy of model or modify inplace
         :return:
         """
+
+        conditions_store = self.get_conditions(update=True)
+        cx = conditions_store['growth_conditions'][conditions_id]
+
         if model is None or isinstance(model, string_types):
             mdl = self.load_model(model)
         elif copy:
             mdl = model.copy()
         else:
             mdl = model
-        
-        conditions_store = self.get_conditions(update=True)
 
-        cx = conditions_store['growth_conditions'][conditions_id]
         load_medium(mdl, cx['media'])
         if "carbon_source" in cx and cx["carbon_source"] is not None:
             # Will throw error if invalid transporter
@@ -425,7 +426,7 @@ class GSMProject(object):
         :param model: cobrapy or cameo model
         :param conditions_id: identifier for the conditions should be unique
         :param carbon_source: name of carbon source in the media that has a fixed uptake rate
-        :param apply_to: iterable of models that this set of conditions applies to
+        :param apply_to: iterable of models that this set of designs applies to
         :param observe_growth: bool or list.
         :return:
         """
@@ -467,7 +468,7 @@ class GSMProject(object):
 
         conditions_store['growth_conditions'][conditions_id] = dict(
             media=media,
-            models=apply_to,
+            designs=apply_to,
             observe_growth=observe_growth,
             carbon_source=carbon_source,
         )

### 2018-02-26 gsmodutils/utils/metacyc.py

@@ -155,6 +155,9 @@ def get_enzyme_reactions(eid, db):
     """
     erids = []
 
+    if eid[:3] != "EC-":
+        eid = "EC-{}".format(eid)
+
     for rid, react in db['reactions'].items():
         if 'EC-NUMBER' in react and eid in react['EC-NUMBER']:
             erids.append(rid)
@@ -186,9 +189,15 @@ def add_reaction(model, reaction_id, db):
     added_metabolites = []
     for mid in metabolites:
         if mid not in model.metabolites:
+            try:
+                cpd = db['compounds'][mid]
+            except KeyError:
+                # Handles missing metabolites
+                cpd = {"COMMON-NAME": mid}
+
             m = Metabolite(id=mid)
-            m.name = db['compounds'][mid]['COMMON-NAME']
-            m.annotation = dict(metacyc_data=db['compounds'][mid])
+            m.name =cpd['COMMON-NAME']
+            m.annotation = dict(metacyc_data=cpd)
             model.add_metabolites([m])
             added_metabolites.append(mid)
 

### 2018-02-26 article/article.tex
article names
changes to article

@@
 \author[2]{Jamie Twycross}
+\author[3]{John King}
+\author[1]{Nigel Minton}
 \affil[1]{Synthetic Biology Research Centre, University of Nottingham, Nottingham, NG7 2RD, United Kingdom}
 \affil[2]{Intelligent Modelling and Analysis group, School of Computer Science, University of Nottingham, Nottingham, NG7 2RD, United Kingdom}
+\affil[3]{School of Mathematical Sciences, University of Nottingham, Nottingham, NG7 2RD, United Kingdom}
 
@@
 \\
-\textbf{Results:} Taking advantage of python and cobrapy integration, we have developed the \textit{gsmodutils} modelling framework.
-As a fully documented python API and series of command line utilities, gsmodutils places an emphasis on test driven design of models through easy to write test cases in json or python.
-Crucially, different substrate conditions are configurable allowing users to examine how different designs or curation impact a wide range of system behaviours, hopefully, minimising error between model versions.
-Furthermore, a core feature of the software is to allow users to easily create shareable, self contained, docker images that can be run in cross platform environments.\\
-\textbf{Availability:} The software framework described within this paper is open source and freely available from http://github.com/gsmodutils \\
+\textbf{Results:} 
+As part of an ongoing effort to improve genome scale metabolic analysis, we have developed a test driven development methodology for the continuous integration of validation data from different sources.
+Contributing to the open source technology based around cobrapy, we have developed the \textit{gsmodutils} modelling framework placing an emphasis on test driven design of models through defined test cases.
+Crucially, different substrate conditions are configurable allowing users to examine how different designs or curation impact a wide range of system behaviours, minimising error between model versions.\\
+\textbf{Availability:} The software framework described within this paper is open source and freely available from http://github.com/nottingham\_sbrc/gsmodutils \\
 \textbf{Contact:}{james.gilbert@nottingham.ac.uk}
@@
 \end{document}
+\grid

### 2018-05-10 article/article.tex
Model diff updates

@@
 \begin{abstract}
-\textbf{Motivation:} Genome scale metabolic models are increasingly important for systems biology and metabolic engineering research as they are capable of simulating complex steady state phenotypic behaviour.
+\textbf{Motivation:} Genome scale metabolic models (GSMMs) are increasingly important for systems biology and metabolic engineering research as they are capable of simulating complex steady state phenotypic behaviour.
 Constraints based models of this form can include thousands of reactions and metabolites, with many crucial pathways that only become activated in specific simulation settings.
@@
 Given its flexibility, low barrier to entry and requirement only on minimal knowledge regarding the stoichiometry of metabolic networks this structural approach has become an extremely popular method for modelling steady state behaviour of large, biochemical networks \cite{kauffman2003advances}.
-Such large scale reconstructions are often refereed to as genome scale models (GSMs), as the processes is significantly aided through the advent of relatively inexpensive genome sequencing \cite{o2015using, land2015insights}.
+Such large scale reconstructions are often refereed to as genome scale metabolic models (GSMMs), as the processes is significantly aided through the advent of relatively inexpensive genome sequencing \cite{o2015using, land2015insights}.
+Indeed, owing to their ability to model complex aspects of metabolism, GSMMs have been widely adopted as a standard to elucidate and optimise of industrial biotechnology processes \cite{kim2017current}.
 
-The reconstruction of GSMs is an arduous process that follows a complex protocol to ensure model validity \cite{thiele2010protocol}.
-Whilst many popular automated methods exists to construct GSMs from reference genomes \cite{henry2010high, poolman2006scrumpy}, there is a still a significant amount of manual curation.
+The reconstruction of GSMMs is an arduous process that follows a complex protocol to ensure model validity \cite{thiele2010protocol}.
+Whilst many popular automated methods exists to construct GSMMs from reference genomes \cite{henry2010high, poolman2006scrumpy}, there is a still a significant amount of manual curation.
 However, treating the creation of models as an isolated ``one-off'' event ignores the significant amount of curation that is required for applications such as biotechnology.
@@
 Furthermore, when research projects end it all to is common for important large models to be published and become relics lost within the literature forgotten to all but the most dedicated of individuals.
-As GSMs grow in terms of the metabolism the contain as well as the biological problems they are used to solve, problems with annotation and curation naturally accumulate as a product of human error.
+As GSMMs grow in terms of the metabolism the contain as well as the biological problems they are used to solve, problems with annotation and curation naturally accumulate as a product of human error.
 Software that facilitates actively improving how researchers develop and apply models to new phenomena is required.
@@
 Indeed, a core objective of this framework is to ensure that good practices are followed in model development that help scientists to better trust the results discovered by their models.
-In an ideal world, we would envision a methodology such as our becoming a pre-requisite for GSMs to pass peer review.
+In an ideal world, we would envision a methodology such as our becoming a pre-requisite for GSMMs to pass peer review.
 

### 2018-05-10 gsmodutils/model_diff.py

@@ -75,9 +75,9 @@ class ModelDiff(dict):
     def _reaction_table(self):
         info_str = ""
         if len(self["reactions"]):
+            info_str += "<strong> Added/Changed reactions: </strong>"
             for reaction in self["reactions"]:
                 info_str += """
-                <strong> Added or changed reactions: </strong>
                 <table>
                 <th> Reaction id</th>
                 <th>name</th>
@@ -86,7 +86,6 @@ class ModelDiff(dict):
                 <th>upper_bound</th>
                 <th>lower_bound</th>
                 <th>objective_coefficient</th>
-                <th>rstr</th>
                 <tr>
                     <td>{id}</td>
                     <td>{name}</td>
@@ -95,18 +94,10 @@ class ModelDiff(dict):
                     <td>{upper_bound}</td>
                     <td>{lower_bound}</td>
                     <td>{objective_coefficient}</td>
-                    <td>{rstr}</td>
                 </tr>
-                <tr>
-                    <td>
-                        <table>
-                            <th> Metabolite </th>
-                            <th> Coef </th>
+                <tr> <td colspan=7 > {rstr} </td> </tr>
+                <tr> <td colspan=7> {rstr_names} </td> </tr>
                 """.format(**reaction)
-
-                for s, v in reaction["metabolites"]:
-                    info_str += "<tr> <td> {} </td> <td> {} </td> </tr>".format(s, v)
-                info_str += "</table></td> </tr>"
 
             info_str += "</table>"
         return info_str
@@ -293,7 +284,8 @@ class ModelDiff(dict):
                         name=rb.name,
                         variable_kind=rb.variable_kind,
                         metabolites=dict(convert_stoich(rb.metabolites)),
-                        rstr=repr(rb),
+                        rstr=rb.build_reaction_string(use_metabolite_names=False),
+                        rstr_names=rb.build_reaction_string(use_metabolite_names=True)
                     )
                 )
         # Gene reaction rules are stored in reactions, however models also contains metadata for genes

### 2018-05-16 article/article.tex
Article changes (moving to use c auto)

@@
 \author[1]{Nicole Pearcy}
+\author[1]{Rupert Norman}
 \author[2]{Jamie Twycross}
@@
 
-\section{Example usage Clostridium acetobutylicum models}
+% \section{Example usage Clostridium acetobutylicum models}
+% \subsection{Scientific background}
+% \textit{C. acetobutylicum} ATCC 824 is a widely researched organism that has found use in the production of solvents for decades.
+% More recently, \textit{C. acetobutylicum} has been applied for the production of alcohols such as ethanol and n-butanols.
+% The industrial relevance and strong background of experimental data has led to the application of many genome scale models \cite{}.
+% Whilst increasingly informative when combined with experimental data, the history of the management of models for \textit{C. acetobutylicum} tells a different story.
+% 
+% Initially, two rival models were published containing a similar number of genes \cite{lee2008genome, senger2008genome}.
+% % \textbf{TODO: describe in more detail}.
+% %Discuss limitations of rival models
+% This led to the further development of subsequent models, iCaC 490 \cite{mcanulty2012genome} and an automatically generated model from  the Model SEED  pipeline \cite{}.
+% 
+% In this example we focus on the more recent contributions of iCac802 \cite{dash2014capturing} and iCac967 \cite{yoo2015quantitative}.
+% iCac802 combines a significantly updated set of annotations from previously curated model versions, an improved TCA cycle, full proton balancing in reactions and reactions for more complete coverage of hexadecanoyl-acp, hexadecanoyl-CoA, crotonyl-CoA pathways as well as pathways for the uptake and biosynthesis of more compounds.
+% Crucially, the model itself received significant validation against fluxonomic data from 13C-metabolic flux analysis ($13$C-MFA).
+% This additional validation set considerable flux bounds upon pathways, leading to improved curation.
+% 
+% Whilst iCac802 was informed by previous models, critically, it was started as a new curation project, a significant undertaking of work.
+% Such an approach may be necessary to uncover errors or oversights in construction processes.
+% However, this naturally adds a significant time burden for model curators and may be due to a lack of adaptability from previous models.
+% 
+% The iCac 967 model presents a smaller set of changes to the annotation of the \textit{C. acetobutylicum} model.
+% While the number of genes increased from 802 to 967 manual annotation actually reduced the number of reactions and metabolites from 1462 to 1137 and 1231 to 1058, respectively.
+% The annotation ane experimental evidence from iCac802 \cite{dash2014capturing} to iCac967 \cite{yoo2015quantitative} is greatly expanded.
+% However, it is not entirely clear, from the initial publication, the extent to which the expanded iCaC967 model preserves crucial experimental validation.
+% The following sections aim to check the later model against previous validation criteria.
+
+% \subsection{Model integration}
+% \textbf{Project construction:}
+% An initial project of the \textit{C. acetobutylicum} model was created in gsmodutils.
+% This is based around the iCac802 base model.
+% Included are also knockout strains $\Delta adc, \Delta ack, \Delta ptb$ and $\Delta hbd$ (TODO: describe knockouts).
+% 
+% In addition, a design listing the constraints changed in model version iCaC967 
+% \footnote{We would like to note that such a significant change to a model, in many practical cases, would not be considered a design but, rather, a separate model.
+% Separate models within the same project is a feature supported by gsmodutils.}.
+% \\
+% \textbf{Model comparison: TODO Diffs between existing models - use (and further develop) gsmodutils diff for model stats}.
+% 
+% \subsection{Evaluation of model validation criteria}
+% Several crucial validation criteria were identified from the experimental and theoretical analysis of the models found in the original publications.
+% We note that this is not an exhaustive set of criteria.
+% This section is designed to serve as a use case for the gsmodutils framework, not to comprehensively review the initial publications.
+% \\
+% 
+% \textbf{TCA cycle}\\
+% \textbf{Proton balanced results}\\
+% \textbf{Predictions of knockouts}\\
+% The inital publication for iCAC802 included validation of four gene knockout mutant strains.
+% $\Delta ack$ acetate kinase.\\
+% $\Delta ptb$ phosphotransbutrylase.\\
+% $\Delta adc$ acetoacetate decarboxylase. \textbf{This reaction is present and the knock out works as expected but not annotated correctly.}\\
+% $\Delta hdb$ hydroxybutyryl-CoA dehydrogenase. \\
+% 
+% \textbf{Flux validation against 13C data}\\
+% \textbf{Under acidogenic stress (H inhibited conditions) } \\
+% \textbf{Solvent formation Cell recycling in solventogenic phase}\\
+
+\section{Example usage Clostridium autoethanogenum}
+\textit{Clostridium autoethanogenum} is an anaerobic, gram positive spore forming acetogenic bacteria that is of industrial relevance \cite{}.
+In this section we describe how gsmodutils has been utilised to ensure that future versions of a genome scale model CLAU--- will remain functionally relevant, from the perspective of industrial biotechnology.
+
 \subsection{Scientific background}
-\textit{C. acetobutylicum} ATCC 824 is a widely researched organism that has found use in the production of solvents for decades.
-More recently, \textit{C. acetobutylicum} has been applied for the production of alcohols such as ethanol and n-butanols.
-The industrial relevance and strong background of experimental data has led to the application of many genome scale models \cite{}.
-Whilst increasingly informative when combined with experimental data, the history of the management of models for \textit{C. acetobutylicum} tells a different story.
 
-Initially, two rival models were published containing a similar number of genes \cite{lee2008genome, senger2008genome}.
-% \textbf{TODO: describe in more detail}.
-%Discuss limitations of rival models
-This led to the further development of subsequent models, iCaC 490 \cite{mcanulty2012genome} and an automatically generated model from  the Model SEED  pipeline \cite{}.
-
-In this example we focus on the more recent contributions of iCac802 \cite{dash2014capturing} and iCac967 \cite{yoo2015quantitative}.
-iCac802 combines a significantly updated set of annotations from previously curated model versions, an improved TCA cycle, full proton balancing in reactions and reactions for more complete coverage of hexadecanoyl-acp, hexadecanoyl-CoA, crotonyl-CoA pathways as well as pathways for the uptake and biosynthesis of more compounds.
-Crucially, the model itself received significant validation against fluxonomic data from 13C-metabolic flux analysis ($13$C-MFA).
-This additional validation set considerable flux bounds upon pathways, leading to improved curation.
-
-Whilst iCac802 was informed by previous models, critically, it was started as a new curation project, a significant undertaking of work.
-Such an approach may be necessary to uncover errors or oversights in construction processes.
-However, this naturally adds a significant time burden for model curators and may be due to a lack of adaptability from previous models.
-
-The iCac 967 model presents a smaller set of changes to the annotation of the \textit{C. acetobutylicum} model.
-While the number of genes increased from 802 to 967 manual annotation actually reduced the number of reactions and metabolites from 1462 to 1137 and 1231 to 1058, respectively.
-The annotation ane experimental evidence from iCac802 \cite{dash2014capturing} to iCac967 \cite{yoo2015quantitative} is greatly expanded.
-However, it is not entirely clear, from the initial publication, the extent to which the expanded iCaC967 model preserves crucial experimental validation.
-The following sections aim to check the later model against previous validation criteria.
+\subsection{}
 
@@
 \textbf{Project construction:}
-An initial project of the \textit{C. acetobutylicum} model was created in gsmodutils.
-This is based around the iCac802 base model.
-Included are also knockout strains $\Delta adc, \Delta ack, \Delta ptb$ and $\Delta hbd$ (TODO: describe knockouts).
 
-In addition, a design listing the constraints changed in model version iCaC967 
-\footnote{We would like to note that such a significant change to a model, in many practical cases, would not be considered a design but, rather, a separate model.
-Separate models within the same project is a feature supported by gsmodutils.}.
-\\
 \textbf{Model comparison: TODO Diffs between existing models - use (and further develop) gsmodutils diff for model stats}.
@@
 \subsection{Evaluation of model validation criteria}
-Several crucial validation criteria were identified from the experimental and theoretical analysis of the models found in the original publications.
-We note that this is not an exhaustive set of criteria.
-This section is designed to serve as a use case for the gsmodutils framework, not to comprehensively review the initial publications.
-\\
-
-\textbf{TCA cycle}\\
-\textbf{Proton balanced results}\\
-\textbf{Predictions of knockouts}\\
-The inital publication for iCAC802 included validation of four gene knockout mutant strains.
-$\Delta ack$ acetate kinase.\\
-$\Delta ptb$ phosphotransbutrylase.\\
-$\Delta adc$ acetoacetate decarboxylase. \textbf{This reaction is present and the knock out works as expected but not annotated correctly.}\\
-$\Delta hdb$ hydroxybutyryl-CoA dehydrogenase. \\
-
-\textbf{Flux validation against 13C data}\\
-\textbf{Under acidogenic stress (H inhibited conditions) } \\
-\textbf{Solvent formation Cell recycling in solventogenic phase}\\
 
@@
 \subsection{Remarks}
-Reproducing the results of the initial publication of the iCAC802 model \cite{dash2014capturing} was possible.
-However, it is notworthy to point out that there were a number of inconsistencies between the annotation in the published model and the publication.
-Notably, the gene ``adc'', which codes for acetoacetate decarboxylase was missing from the initial annotation.
-We would stress that these did not significantly impact the results of the model or the claims made in the publication.
 
 \section{Related work}
+\textbf{Memote:} \cite{lieven2017memote} is tool set that has very similar ambitions to gsmodutils.
+It features a fully specified set of tests, including custom test cases and has strong version control integration with git.
+A core difference between these projects is that gsmodutils has a stronger focus on reducing the redundancy in model storage through the usage of \textit{design deltas}, as described above.
+Similarly, a core goal of gsmodutils is to allow easy import and export outside of the framework for compatibility with other modelling suites.
+It should be noted that, as memote is written in python and utilises cobrapy, it is fully compatible with gsmodutils.
+
 \textbf{cameo:} Cameo \cite{cardoso2017cameo} is a set of strain design utilities, such as genetic algorithms for finding optimal gene knock-out
@@
 \section{Discussion}
-In order to facilitate the sharing and dissemination of research good standards and software are required \cite{}.
-Naturally a great deal of effort has gone in to producing high quality systems and synthetic biology standards \cite{}.
+In order to facilitate the sharing and dissemination of high quality computational research, good standards and software are required \cite{jimenez2017four}.
+Naturally a great deal of effort has gone in to producing high quality systems and synthetic biology standards \cite{hucka2003systems, cox2018synthetic}.
 Furthermore, when research projects end it all to is common for important large models to be published and become relics lost within the literature forgotten to all but the most dedicated of individuals.
@@
 
-% Furthermore, several methods for the automated construction of gsms exist.
-% However, to our knowledge no methods include a fully automated pipeline for updating existing manually created models once constructed.
-% We would like to note that the collection of data for the validation of models and the construction of tests could be used to allow automated updates to take place.
-% For example, as GPR are curated models could automatically be updated with the inclusion of new information whilst being validated against existing criteria.
-
 As with most software development projects, gsmodutils will see expanded features.
@@
 \section*{Acknowledgements}
-We would like to thank the Oxford Brookes Cell Systems Modelling group for helpful discussions regarding this work, Rupert Norman for assistance testing and forming the idea of this software, Thomas Millat for crucial scientific advice and Phillipe Soucaille for assistance reconstructing the iCac967
-
-\section*{Funding}
-This work has been supported by the BBSRC grant XXXXX.
+We would like to thank the Oxford Brookes Cell Systems Modelling group for helpful discussions regarding this work and Thomas Millat for scientific advice.
+This work was supported by the UK Biotechnology and Biological Sciences Research Council (BBSRC) grant BB/L013940/1.
 

### 2018-05-16 gsmodutils/cli.py
Added tqdm progressbar to tests

@@ -15,6 +15,8 @@ from gsmodutils import GSMProject, load_model
 from gsmodutils.project.project_config import ProjectConfig
 from sys import exit
 
+from tqdm import tqdm
+
 
 def _load_project(project_path):
     project = None
@@ -118,7 +120,7 @@ def test(project_path, test_id, skip_default, verbose, log_path):
         exit(0)
     
     barstr = "-"*25
-    # TODO Progress bar as tests are run
+
     click.echo(
         click.style(barstr + ' gsmodutils test results ' + barstr, bg='green', bold=True)
     )
@@ -133,7 +135,8 @@ def test(project_path, test_id, skip_default, verbose, log_path):
     click.echo()
     ts = 0
     te = 0
-    for tf, log in tester.log.items():
+
+    for tf, log in tqdm(tester.log.items()):
 
         if tf == 'default_tests':
             click.echo('Default project file tests (models, designs, conditions):')

### 2018-05-16 gsmodutils/cli.py
Added tqdm progressbar to tests

@@ -130,13 +130,13 @@ def test(project_path, test_id, skip_default, verbose, log_path):
         click.echo()
 
     print('Running tests: ', end='')
-    for _ in tester.iter_tests(skip_default=skip_default):
+    for _ in tqdm(tester.iter_tests(skip_default=skip_default)):
         print('.', end='')
     click.echo()
     ts = 0
     te = 0
 
-    for tf, log in tqdm(tester.log.items()):
+    for tf, log in tester.log.items():
 
         if tf == 'default_tests':
             click.echo('Default project file tests (models, designs, conditions):')

### 2018-05-16 gsmodutils/cli.py
Added warnings for tests

@@ -52,11 +52,16 @@ def _output_child_logs(log, verbose=False, indent=4, baseindent=4):
             click.echo(
                 click.style(idt + "Asserion error: " + msg, fg='red')
             )
-            
+
+        for msg, desc in clog.warnings:
+            click.echo(
+                click.style(idt + "Warning: " + msg, fg='yellow')
+            )
+
         if verbose:
             for msg, desc in clog.success:
                 click.echo(
-                    click.style(idt + "Asserion success: " + msg, fg='green')
+                    click.style(idt + "Assertion success: " + msg, fg='green')
                 )
             
         _output_child_logs(clog, verbose=verbose, indent=indent+baseindent)
@@ -129,9 +134,9 @@ def test(project_path, test_id, skip_default, verbose, log_path):
         click.echo("verbose mode, showing successes and failures")
         click.echo()
 
-    print('Running tests: ', end='')
+    click.echo('Running tests: ')
     for _ in tqdm(tester.iter_tests(skip_default=skip_default)):
-        print('.', end='')
+        pass
     click.echo()
     ts = 0
     te = 0

### 2018-05-16 gsmodutils/test/utils.py

@@ -126,6 +126,7 @@ class TestRecord(object):
         self.parent = parent
         self.success = []
         self.error = []
+        self.warnings = []
         self.std_out = None  # Reserved for messages
         self.run_time = time.time()
         self.children = {}
@@ -145,6 +146,14 @@ class TestRecord(object):
             self.success.append((success_msg, desc))
         else:
             self.error.append((error_msg, desc))
+
+    def warning(self, statement, message, desc=''):
+        """
+        Called within test functions to capture warnings about the status of models.
+        If statement is true, the warning message will be stored.
+        """
+        if statement:
+            self.warnings.append((message, desc))
     
     def add_error(self, msg, desc=''):
         """

### 2018-05-16 tests/test_tests.py

@@ -107,6 +107,7 @@ def test_func_cove(model, project, log):
 def test_model(model, project, log):
     solution = model.solver.optimize()
     print('This is the end')
+    log.warning(True, "this is a warning")
     log.assertion(True, "Model grows", "Model does not grow")
     log.assertion(False, "Model grows", "Model does not grow")
     

### 2018-05-16 gsmodutils/cli.py
Fixed message output concat with no strings

@@ -50,18 +50,18 @@ def _output_child_logs(log, verbose=False, indent=4, baseindent=4):
         )
         for msg, desc in clog.error:
             click.echo(
-                click.style(idt + "Asserion error: " + msg, fg='red')
+                click.style(idt + "Asserion error: {}".format(msg), fg='red')
             )
 
         for msg, desc in clog.warnings:
             click.echo(
-                click.style(idt + "Warning: " + msg, fg='yellow')
+                click.style(idt + "Warning: {}".format(msg), fg='yellow')
             )
 
         if verbose:
             for msg, desc in clog.success:
                 click.echo(
-                    click.style(idt + "Assertion success: " + msg, fg='green')
+                    click.style(idt + "Assertion success: {}".format(msg), fg='green')
                 )
             
         _output_child_logs(clog, verbose=verbose, indent=indent+baseindent)

### 2018-05-22 gsmodutils/cli.py
Ordering of how errors are outputted in test reports

@@ -88,26 +88,6 @@ def test(project_path, test_id, skip_default, verbose, log_path):
     click.echo('Collecting tests...')
     tester.collect_tests()
 
-    # Display errors
-    for tf, e in tester.load_errors:
-        click.echo(
-            click.style('Error loading test {} \n {}'.format(tf, e), fg='red')
-        )
-    
-    for tf, e in tester.syntax_errors.items():
-        click.echo(
-            click.style('Error - {} has syntax errors {}'.format(tf, e), fg='red')
-        )
-    
-    for tf, entry_key, missing_fields in tester.invalid_tests:
-        click.echo(
-            click.style('---Error with formating of test {} in {} ---'.format(tf, entry_key), fg='yellow')
-        )
-        
-        click.echo('Missing fields:')
-        for field in missing_fields:
-            click.echo('\t{}'.format(field))
-    
     if test_id is not None:
         # Only run specific test id
         if test_id not in tester.test_ids:
@@ -172,6 +152,26 @@ def test(project_path, test_id, skip_default, verbose, log_path):
             )
     percent = round(((ts - te) / ts) * 100, 3)
     click.echo('Ran {} test assertions with a total of {} errors ({}% success)'.format(ts, te, percent))
+
+    # Display errors
+    for tf, e in tester.load_errors:
+        click.echo(
+            click.style('Error loading test {} \n {}'.format(tf, e), fg='red')
+        )
+
+    for tf, e in tester.syntax_errors.items():
+        click.echo(
+            click.style('Error - {} has syntax errors {}'.format(tf, e), fg='red')
+        )
+
+    for tf, entry_key, missing_fields in tester.invalid_tests:
+        click.echo(
+            click.style('---Error with formating of test {} in {} ---'.format(tf, entry_key), fg='yellow')
+        )
+
+        click.echo('Missing fields:')
+        for field in missing_fields:
+            click.echo('\t{}'.format(field))
 
     # Save report to json log file
     if log_path is not None:

### 2018-05-22 article/article.tex
Fixed progress bar for tests by adding method to the tester class.

@@
 
-\section{Example usage Clostridium autoethanogenum}
+\section{Case study usage Clostridium autoethanogenum}
 \textit{Clostridium autoethanogenum} is an anaerobic, gram positive spore forming acetogenic bacteria that is of industrial relevance \cite{}.
@@
 
-\textbf{Model comparison: TODO Diffs between existing models - use (and further develop) gsmodutils diff for model stats}.
+\subsection{Evaluation of model validation criteria}
 
-\subsection{Evaluation of model validation criteria}
+\begin{itemize}
+ \item Energetic consistency: All transporters are set to 0 lower bound and the ATPase reaction is given a fixed flux of one.
+ Providing the model has no cycles allowing ATP production without a food source, the model should never grow.
+
+ \item Minimisation of flux based tests. One conventional approach is to set an optimisation criteria of minimising flux across enzymatic reactions with a fixed biomass constraint.
+ Often known as parsimonious FBA, this was created as a design for the model allowing the creation of reversible reactions (for all model reactions) in a programmatic manner.
+
+ \item Product scans.
+Of interest to this project was the product spectrum under conditions where the organism can and cannot produce molecular Hydrogen (when growing on Carbon monoxide as an energy source).
+ Where Hydrogen production is blocked the organism should produce lactate and butanediol as by products.
+
+ \item The NFN mutant strain. This knockout mutant is shown to block production of acetate and increase the production of ethanol.
+
+ \item Lethal knockout mutants. These mutants are of interest to GSM design and development because any changes to systemic behaviour should preserve this behaviour.
+\end{itemize}
 

### 2018-05-22 gsmodutils/cli.py

@@ -15,8 +15,6 @@ from gsmodutils import GSMProject, load_model
 from gsmodutils.project.project_config import ProjectConfig
 from sys import exit
 
-from tqdm import tqdm
-
 
 def _load_project(project_path):
     project = None
@@ -115,8 +113,7 @@ def test(project_path, test_id, skip_default, verbose, log_path):
         click.echo()
 
     click.echo('Running tests: ')
-    for _ in tqdm(tester.iter_tests(skip_default=skip_default)):
-        pass
+    tester.progress_tests(skip_default=skip_default)
     click.echo()
     ts = 0
     te = 0

### 2018-05-22 gsmodutils/test/tester.py

@@ -12,6 +12,7 @@ from six import exec_
 
 import gsmodutils
 from gsmodutils.test.utils import TestRecord
+from tqdm import tqdm
 
 
 @contextlib.contextmanager
@@ -408,6 +409,19 @@ class GSMTester(object):
                 yield func(tid)
             elif func != self.iter_basetf:
                 yield func(*tid)  # generator for the test
+
+    def progress_tests(self, recollect=False, skip_default=False):
+        """
+        Run tests with a progressbar
+        :param recollect:
+        :param skip_default:
+        :return:
+        """
+        for tid, func in tqdm(self._task_execs.items()):
+            if func == self._exec_default and not skip_default:
+                func(tid)
+            elif func != self.iter_basetf:
+                func(*tid)  # generator for the test
 
     def run_all(self, recollect=False):
         """Find and run all tests for a project, executes rather than returning generator"""

### 2018-05-22 gsmodutils/project/interface.py
Wildcard model, design and conditions selection for tests. Also added run tests from project (useful for notebooks)

@@ -60,6 +60,16 @@ class GSMProject(object):
         """Creates a tester for this project instance"""
         return GSMTester(self)
 
+    def run_tests(self):
+        """
+        Returns the log output of all the tests
+        :return:
+        """
+        tester = GSMTester(self)
+        tester.collect_tests()
+        tester.progress_tests()
+        return tester.to_dict()
+
     @property
     def project_context_lock(self):
         """
@@ -113,6 +123,10 @@ class GSMProject(object):
         """
         for mdl_path in self.config.models:
             yield self.load_model(mpath=mdl_path)
+
+    @property
+    def list_models(self):
+        return self.config.models
 
     @property
     def models(self):
@@ -351,6 +365,11 @@ class GSMProject(object):
         diff['parent'] = None
         tmp_design = StrainDesign.from_dict('tmp_design', diff, self)
         return tmp_design.load()
+
+    @property
+    def list_conditions(self):
+        conditions_store = self.get_conditions(update=True)
+        return list(conditions_store["growth_conditions"].keys())
 
     def load_conditions(self, conditions_id, model=None, copy=False):
         """

### 2018-05-22 gsmodutils/test/utils.py

@@ -45,16 +45,8 @@ class ModelTestSelector(object):
             designs = []
 
         self.models = models
-        if not len(self.models):
-            self.models = [None]
-        
         self.conditions = conditions
-        if not len(self.conditions):
-            self.conditions = [None]
-        
         self.designs = designs
-        if not len(designs):
-            self.designs = [None]
 
     def __call__(self, func):
         """
@@ -64,7 +56,25 @@ class ModelTestSelector(object):
         def wrapper(*args, **kwargs):
             project = args[1]
             log = args[2]
-            
+
+            if self.models == "*":
+                self.models = project.list_models
+
+            if self.designs == "*":
+                self.designs = project.list_designs + [None]
+
+            if self.conditions == "*":
+                self.conditions = project.list_conditions + [None]
+
+            if not len(self.models):
+                self.models = [None]
+
+            if not len(self.conditions):
+                self.conditions = [None]
+
+            if not len(self.designs):
+                self.designs = [None]
+
             for mn in self.models:
                 if mn is None:
                     mn = project.config.default_model
@@ -81,7 +91,7 @@ class ModelTestSelector(object):
                             tid = (mn, cid)
                         elif did is not None:
                             tid = (mn, did)
-                        
+
                         nlog = log.create_child(tid, param_child=True)
                         
                         try:

### 2018-05-23 gsmodutils/cli.py
Changed formatting of cli utils

@@ -132,7 +132,7 @@ def test(project_path, test_id, skip_default, verbose, log_path):
         click.echo("Counted {} test assertions with {} failures".format(*lc))
         # Output base test file
         if not log.is_success:
-            click.echo(click.style('{} has errors'.format(indicator), fg='red'))
+            click.echo(click.style('{} has test errors'.format(indicator), fg='red'))
         else:
             click.echo(click.style('{} completed all tests without error'.format(indicator), fg='green'))
 
@@ -404,8 +404,9 @@ def export(file_format, filepath, project_path, model_id, design, conditions, ov
 def info(project_path):
     """ Display all the information about a gsmodutils project (list models, paths, designs etc. """
     project = _load_project(project_path)
-    click.echo('''--------------------------------------------------------------------------------------------------
-Project description - {description}
+
+    click.echo("-" * click.get_terminal_size()[0])
+    click.echo('''Project description - {description}
 Author(s): - {author}
 Author email - {author_email}
 Designs directory - {design_dir}
@@ -415,27 +416,32 @@ Tests directory - {tests_dir}
     click.echo("Models:")
     for mdl_path in project.config.models:
         model = project.load_model(mdl_path)
-        click.echo("\t* {}".format(mdl_path))
+        click.echo(click.style("\t* {}".format(mdl_path), fg="green"))
         click.echo("\t\t {}".format(model.id))
 
     click.echo("Designs:")
     for d in project.list_designs:
         try:
             design = project.get_design(d)
-            click.echo("\t* {}".format(design.id))
-            click.echo("\t\t {}".format(design.name))
-            click.echo("\t\t {}".format(design.description))
-            if design.parent is not None:
-                click.echo("\t\t Parent: {}".format(design.parent.id))
-        except DesignError:
-            click.echo(click.style("\t* Error loading design {}".format(d)), color="red")
-        except DesignOrphanError:
-            click.echo(click.style("\t* Appears to be problem with parent of design {}".format(d)), color="red")
+            click.echo("*" * click.get_terminal_size()[0])
+            click.echo(click.style("\t* {} {}".format(design.name, design.id), fg="green"))
 
+            if design.parent is not None:
+                click.echo("\tParent: {}".format(design.parent.id))
+
+            click.echo("\t\t{}".format(design.description))
+        except DesignError:
+            click.echo(click.style("\t* Error loading design {}".format(d)), fg="red")
+        except DesignOrphanError:
[…]

### 2018-06-29 article/article.tex
I've written a bit of text for the C auto case study bit.

--- 2018-05-16 article/article.tex
@@
 
-\section{Example usage Clostridium autoethanogenum}
-\textit{Clostridium autoethanogenum} is an anaerobic, gram positive spore forming acetogenic bacteria that is of industrial relevance \cite{}.
-In this section we describe how gsmodutils has been utilised to ensure that future versions of a genome scale model CLAU--- will remain functionally relevant, from the perspective of industrial biotechnology.
+\section{Case study usage Clostridium autoethanogenum}
+\textit{Clostridium autoethanogenum} is a bacterial species used for the production of commodity chemicals at industrial scale \cite{abrini1994clostridium,norman2018progress,LT}.
+A new GSMM of \textit{C.
+autoethanogenum}, `MetaCLAU', has been analysed to improve this bioprocess (Norman et al., unpublished).
+
+In this section we describe how gsmodutils has been utilised to ensure that future versions of MetaCLAU will remain functionally relevant from the perspective of industrial biotechnology.
 
@@
 
-\subsection{}
+\textit{C.
+autoethanogenum} is a strictly anaerobic, acetogenic bacterium which naturally produces ethanol and trace amounts of 2,3-butanediol (2,3BD) from carbon monoxide and water \cite{abrini1994clostridium,schuchmann2014autotrophy,norman2018progress}.
+Since carbon monoxide is readily available in the form of industrial waste gas, and 2,3BD has a global market value of \$43 billion \cite{kopke201123butanediol}, the optimisation of yields of 2,3BD from carbon monoxide is highly desirable in the context of industry \cite{norman2018progress}.
 
 \subsection{Model integration}
-\textbf{Project construction:}
 
-\textbf{Model comparison: TODO Diffs between existing models - use (and further develop) gsmodutils diff for model stats}.
+MetaCLAU was built using Pathway Tools \cite{karp2002pathway} and ScrumPy \cite{poolman2006scrumpy}, and is based on a manually annotated genome sequence of \textit{C.
+autoethanogenum} \cite{humphreys2015whole}.
+The resulting model consisted of 758 reactions, 773 metabolites and 518 genes.
+For full details of the model, see (Norman et al.
+unpublished)
+
+The model has been integrated with the gsodutils modelling framework as a test-driven project.
+The following section details specific tests used to evaluate the model at each stage of its continued development.
 
@@
 
+Energetic consistency: An important limitation of FBA is that optimal solutions may be thermodynamically infeasible if appropriate constraints are not applied \cite{Fell1197rohtua}.
+In order to identify these unwanted flux distributions and constrain the model such that they are not in the feasible soluton space, a diagnostic FBA is applied with the following constraints: 1) All transport reactions are constrained to allow no uptake (lower bound = 0), and 2) the ATPase reaction is given a fixed flux of one.
+If a (non-zero) solution to this problem exists, it must contain a thermodynamic inconsistency, which can be dealt with by manual inspection of the solution and modification of one or more of the involved reactions \cite{Fell1197rohtua}.
+
+Minimisation of flux based tests: One conventional approach in FBA is to set an optimisation criteria of minimising flux across enzymatic reactions with a fixed biomass constraint \cite{holzhutter2006thegeneralised}.
+The solution to this FBA problem represents minimal protein investment \cite{holzhutter2006thegeneralised}.
+Since execution of flux minimisation in COBRApy requires a model in which reversible reactions are split into two irreversible reactions, a gsmodutils design was created in which all reversible reactions are split in a programmatic manner.
+In the case of MetaCLAU, the flux minimal solution includes both ethanol and acetate production, which represents good qualitative agreement with experimental data (Norman et al., unpublished).
+Since any changes to this predicted phenotype must be investigated, the flux minimisation analysis has been formulated as a gsmuodutils test.
+
+Product scans: Of interest to this project were changes in the product spectrum of \textit{C.
+autoethanogenum} under conditions where the organism can and cannot produce molecular hydrogen (with carbon monoxide as sole carbon and energy source).
+The hypothesis tested in (Norman et al., unpublished), was that in the case where hydrogen production in infeasible, alternative electron sinks like lactate and 2,3BD would be produced.
+As in the previous case, the model-predicted behaviour showing both lactate and 2,3BD was deemed an important result, which model curators should be notified of if lost during model development.
+Thus the analysis was built into a gsmodutils test.
+
+Lethal knockout mutants: The prediction of lethal single-gene KO mutants through FBA of a GSMM is useful in two ways: 1) the identification of essential genes is an important first step for metabolic engineering strategies, and 2) with the advent of high-throughput TraDIS gene-essentiality data sets \cite{langridge2009simultaneous}, GSMMs can be validated by their abiity to predict essential genes.
+Furthermore, any change in the set of essential genes (particularly an increase in the number of essentials) rpresents important information for metabolic engineering.
+For these reasons, a test been built into the MetaCLAU project which enables the computation of the set of essential genes and their comparison with TraDIS data sets.
+
+$\Delta \textit{nfn}$ knockout mutant: NOT SURE ABOUT THIS. CHRIS NOW THINKS IT'S ESSENTIAL :(
+
 % Discuss these results
 \subsection{Remarks}
+
+James smells.
 

### 2018-06-29 article/article.tex
Minor changes to article text

@@
 \title{Gsmodutils: A python based framework for test driven genome scale metabolic model development}
-\author[1]{James Gilbert}
+\author[1]{James P Gilbert}
 \author[1]{Nicole Pearcy}
@@
 \affil[1]{Synthetic Biology Research Centre, University of Nottingham, Nottingham, NG7 2RD, United Kingdom}
-\affil[2]{Intelligent Modelling and Analysis group, School of Computer Science, University of Nottingham, Nottingham, NG7 2RD, United Kingdom}
+\affil[2]{Intelligent Modelling and Analysis group, School of Computer Science, University of Nottingham, Nottingham, NG8 1BB, United Kingdom}
 \affil[3]{School of Mathematical Sciences, University of Nottingham, Nottingham, NG7 2RD, United Kingdom}
@@
 Similarly, designs are self contained and will not interfere with one another allowing a form of project management.
-Figure \ref{fig:strain_hered}
+Figure \ref{fig:strain_hered} shows how this could work in a practical situation.
+Designs of this nature can also be programmatic, allowing the implementation of features such as non-standard constraints that can be dynamically loaded.
+Furtheremore, integrity of all strain designs is automatically tested as part of the default gsmodutils testing framework.
 
@@
 
-\subsection{Scientific background}
+\subsection{Scientific background and Model integration}
 
@@
 Since carbon monoxide is readily available in the form of industrial waste gas, and 2,3BD has a global market value of \$43 billion \cite{kopke201123butanediol}, the optimisation of yields of 2,3BD from carbon monoxide is highly desirable in the context of industry \cite{norman2018progress}.
-
-\subsection{Model integration}
 
@@
 
-% Discuss these results
-\subsection{Remarks}
-
-James smells.
-
 \section{Related work}
@@
 It should be noted that, as memote is written in python and utilises cobrapy, it is fully compatible with gsmodutils.
-
-\textbf{cameo:} Cameo \cite{cardoso2017cameo} is a set of strain design utilities, such as genetic algorithms for finding optimal gene knock-out
+\\
+\textbf{Cameo:} \cite{cardoso2017cameo} is a set of strain design utilities, such as genetic algorithms for finding optimal gene knock-out
 strategies, as well as other flux balance analysis utilities. Written in python and part of the cobrapy suite of applications,
@@
 Furthermore, the objective of the project is to cultivate collaboration by simplifying the process of distributing large models to different users.
-As a consequence, methods for sharing docker containers will be investigated.
+As a consequence, methods for sharing reusable docker containers will be investigated.
 

### 2018-07-03 article/article.tex
Added changes and corrections to article

@@
 \usepackage{subfigure}
-
+\usepackage{hyperref}
 %PREPRINT/MANUSCRIPT OPTIONS
@@
 As part of an ongoing effort to improve genome scale metabolic analysis, we have developed a test driven development methodology for the continuous integration of validation data from different sources.
-Contributing to the open source technology based around cobrapy, we have developed the \textit{gsmodutils} modelling framework placing an emphasis on test driven design of models through defined test cases.
+Contributing to the open source technology based around COBRApy, we have developed the \textit{gsmodutils} modelling framework placing an emphasis on test driven design of models through defined test cases.
 Crucially, different substrate conditions are configurable allowing users to examine how different designs or curation impact a wide range of system behaviours, minimising error between model versions.\\
-\textbf{Availability:} The software framework described within this paper is open source and freely available from http://github.com/nottingham\_sbrc/gsmodutils \\
+\textbf{Availability:} The software framework described within this paper is open source and freely available from \url{http://github.com/SBRCNottingham/gsmodutils} \\
 \textbf{Contact:}{james.gilbert@nottingham.ac.uk}
@@
 Furthermore, as with many areas of bioinformatic study the proliferation of available computational tools has become somewhat vast.
-This covers huge variety of software platforms including the COBRA toolbox in Matlab \cite{schellenberger2011quantitative}, ScrumPy and CobraPy in Python \cite{poolman2006scrumpy, ebrahim2013cobrapy} with additional tools and libraries such as cameo \cite{cardoso2017cameo}, OptFlux in Java \cite{rocha2010optflux}, SurreyFBA \cite{gevorgyan2010surreyfba}.
+This covers huge variety of software platforms including the COBRA toolbox in Matlab \cite{schellenberger2011quantitative}, ScrumPy and COBRApy in Python \cite{poolman2006scrumpy, ebrahim2013cobrapy} with additional tools and libraries such as cameo \cite{cardoso2017cameo}, OptFlux in Java \cite{rocha2010optflux}, SurreyFBA \cite{gevorgyan2010surreyfba}.
 Whilst most of these tools are Open Source and follow standards, such as SBML \cite{finney2003systems}, there are often few options that make analysis easy to replicate.
@@
 By this we mean the notion that, as a model is curated to represent biological phenomena, much of the validation can be turned into specific test cases that can be repeated between model versions.
-We provide an example test case for \textit{C.
-acetobutylicum} an organism that has had considerable focus in terms of genome scale models and how a working methodology using the software presented here can reduce repetition of work and improve the reproducibility of results.
+We provide an example test case for \textit{Clostridium autoethanogenum}, an organism that has had considerable focus in terms of genome scale models and how a working methodology using the software presented here can reduce repetition of work and improve the reproducibility of results.
 This article aims to summarise the main objectives of the gsmodutils software and we refer the reader to the software user guide for a more detailed exploration features.
@@
 
-By automatically integrating the COBRAPy \cite{ebrahim2013cobrapy} users can easily write convenient test cases following examples given in the user guide.
+By automatically integrating the COBRApy \cite{ebrahim2013cobrapy} users can easily write convenient test cases following examples given in the user guide.
 A standard test case, ensuring that a given model grows on media is given in Box \ref{fig:test_case}.
@@
 
-Energetic consistency: An important limitation of FBA is that optimal solutions may be thermodynamically infeasible if appropriate constraints are not applied \cite{Fell1197rohtua}.
-In order to identify these unwanted flux distributions and constrain the model such that they are not in the feasible soluton space, a diagnostic FBA is applied with the following constraints: 1) All transport reactions are constrained to allow no uptake (lower bound = 0), and 2) the ATPase reaction is given a fixed flux of one.
+\textbf{Energetic consistency:} 
+An important limitation of FBA is that optimal solutions may be thermodynamically infeasible if appropriate constraints are not applied \cite{Fell1197rohtua}.
+In order to identify these unwanted flux distributions and constrain the model such that they are not in the feasible soluton space, a diagnostic FBA is applied with the following constraints:
+1) All transport reactions are constrained to allow no uptake (lower bound = 0), and 2) the ATPase reaction is given a fixed flux of one.
 If a (non-zero) solution to this problem exists, it must contain a thermodynamic inconsistency, which can be dealt with by manual inspection of the solution and modification of one or more of the involved reactions \cite{Fell1197rohtua}.
 
-Minimisation of flux based tests: One conventional approach in FBA is to set an optimisation criteria of minimising flux across enzymatic reactions with a fixed biomass constraint \cite{holzhutter2006thegeneralised}.
+\textbf{Minimisation of flux based tests:}
+One conventional approach in FBA is to set an optimisation criteria of minimising flux across enzymatic reactions with a fixed biomass constraint \cite{holzhutter2006thegeneralised}. 
 The solution to this FBA problem represents minimal protein investment \cite{holzhutter2006thegeneralised}.
@@
 
-Product scans: Of interest to this project were changes in the product spectrum of \textit{C.
+\textbf{Product scans:}
+Of interest to this project were changes in the product spectrum of \textit{C.
 autoethanogenum} under conditions where the organism can and cannot produce molecular hydrogen (with carbon monoxide as sole carbon and energy source).
@@
 
-Lethal knockout mutants: The prediction of lethal single-gene KO mutants through FBA of a GSMM is useful in two ways: 1) the identification of essential genes is an important first step for metabolic engineering strategies, and 2) with the advent of high-throughput TraDIS gene-essentiality data sets \cite{langridge2009simultaneous}, GSMMs can be validated by their abiity to predict essential genes.
-Furthermore, any change in the set of essential genes (particularly an increase in the number of essentials) rpresents important information for metabolic engineering.
+\textbf{Lethal knockout mutants:} 
+The prediction of lethal single-gene KO mutants through FBA of a GSMM is useful in two ways: 
+1) the identification of essential genes is an important first step for metabolic engineering strategies, 
+and 2) with the advent of high-throughput TraDIS gene-essentiality data sets \cite{langridge2009simultaneous}, GSMMs can be validated by their abiity to predict essential genes.
+Furthermore, any change in the set of essential genes (particularly an increase in the number of essentials) represents important information for metabolic engineering.
 For these reasons, a test been built into the MetaCLAU project which enables the computation of the set of essential genes and their comparison with TraDIS data sets.
 
-$\Delta \textit{nfn}$ knockout mutant: NOT SURE ABOUT THIS. CHRIS NOW THINKS IT'S ESSENTIAL :(
+% $\Delta \textit{nfn}$ knockout mutant: NOT SURE ABOUT THIS. CHRIS NOW THINKS IT'S ESSENTIAL :(
+These test cases are all fully reproducible and available as an example within the gsmodutils software repository and we encourage the reader to explore these examples to understand how to utilise gsmodutils within their own project.
 
@@
 Similarly, a core goal of gsmodutils is to allow easy import and export outside of the framework for compatibility with other modelling suites.
-It should be noted that, as memote is written in python and utilises cobrapy, it is fully compatible with gsmodutils.
+It should be noted that, as memote is written in python and utilises COBRApy, it is fully compatible with gsmodutils.
 \\
 \textbf{Cameo:} \cite{cardoso2017cameo} is a set of strain design utilities, such as genetic algorithms for finding optimal gene knock-out
-strategies, as well as other flux balance analysis utilities. Written in python and part of the cobrapy suite of applications,
+strategies, as well as other flux balance analysis utilities. Written in python and part of the COBRApy suite of applications,
 gsmodutils integrates easily with cameo.

### 2018-08-13 article/article.tex
Article authors (added Charlie)
Article draft changes

@@
-\documentclass[a4paper,10pt]{article}
+\documentclass[a4paper,11pt]{article}
 \usepackage[utf8]{inputenc}
@@
 \title{Gsmodutils: A python based framework for test driven genome scale metabolic model development}
-\author[1]{James P Gilbert}
-\author[1]{Nicole Pearcy}
+\author[1,2]{James P Gilbert}
+\author[1,2]{Nicole Pearcy}
 \author[1]{Rupert Norman}
@@
 \author[3]{John King}
+\author[4]{Charlie Hodgman}
 \author[1]{Nigel Minton}
@@
 \affil[3]{School of Mathematical Sciences, University of Nottingham, Nottingham, NG7 2RD, United Kingdom}
+\affil[4]{School of Biosciences, University of Nottingham, Sutton Bonington, Loughborough, LE12 5RD, United Kingdom}
 
@@
 Given its flexibility, low barrier to entry and requirement only on minimal knowledge regarding the stoichiometry of metabolic networks this structural approach has become an extremely popular method for modelling steady state behaviour of large, biochemical networks \cite{kauffman2003advances}.
-Such large scale reconstructions are often refereed to as genome scale metabolic models (GSMMs), as the processes is significantly aided through the advent of relatively inexpensive genome sequencing \cite{o2015using, land2015insights}.
+Such large scale reconstructions are often referred to as genome scale metabolic models (GSMMs), as the processes is significantly aided through the advent of relatively inexpensive genome sequencing \cite{o2015using, land2015insights}.
 Indeed, owing to their ability to model complex aspects of metabolism, GSMMs have been widely adopted as a standard to elucidate and optimise of industrial biotechnology processes \cite{kim2017current}.
@@
 
-Furthermore, as with many areas of bioinformatic study the proliferation of available computational tools has become somewhat vast.
-This covers huge variety of software platforms including the COBRA toolbox in Matlab \cite{schellenberger2011quantitative}, ScrumPy and COBRApy in Python \cite{poolman2006scrumpy, ebrahim2013cobrapy} with additional tools and libraries such as cameo \cite{cardoso2017cameo}, OptFlux in Java \cite{rocha2010optflux}, SurreyFBA \cite{gevorgyan2010surreyfba}.
+Furthermore, as with many areas of bioinformatic study the number available computational tools has become vast.
+This covers huge variety of software platforms including the COBRA toolbox in Matlab \cite{schellenberger2011quantitative}, ScrumPy and COBRApy in Python \cite{poolman2006scrumpy, ebrahim2013cobrapy} with additional tools and libraries such as cameo \cite{cardoso2017cameo}, OptFlux in Java \cite{rocha2010optflux} and SurreyFBA \cite{gevorgyan2010surreyfba}.
 Whilst most of these tools are Open Source and follow standards, such as SBML \cite{finney2003systems}, there are often few options that make analysis easy to replicate.
-
 The archetype design, build and test cycle of synthetic biology heavily relies on the usage of bioinformatics software and modelling to improve the production of natural products \cite{carbonell2016bioinformatics}.
@@
 
-For example, many tools such as RetroPath \cite{carbonell2013retropath}, XTMS \cite{carbonell2014xtms} and GEM-Path \cite{campodonico2014generation} suggest thousands of potential hetrologous pathways as well as providing predictions for flux distributions.
+For example, many tools such as RetroPath \cite{carbonell2013retropath}, XTMS \cite{carbonell2014xtms} and GEM-Path \cite{campodonico2014generation} suggest thousands of potential heterologous pathways as well as providing predictions for flux distributions.
 Many of these tools significantly increase the value of genome scale models, for example by coupling commodity production to an organism's growth \cite{feist2010model}.
 These tools all suggest major changes to wild-type strains must be tracked and compared to allow models to remain relevant.
-In effect, mechanisms are required to related modified models to modified test and production strains.
+In effect, mechanisms are required to relate modified models to modified test and production strains.
 
@@
 This is borne out of a number significant constraints in the development of genome scale models.
-The waterfall approach for model is characterised in Figure \ref{fig:methodology_approaches} where requirements for a model are captured at an early stage and are hard to refine during the production process.
+The waterfall approach for modelling is characterised in Figure \ref{fig:methodology_approaches} where requirements for a model are captured at an early stage and are hard to refine during the production process.
 
@@
 This modelling formation costs an in depth understanding of dynamic behaviour. 
-However, capturing steady state phenomena still provides a good understanding of systemic properties \cite{o2015using}.
+However, capturing steady state phenomena still provides a good understanding of system properties \cite{o2015using}.
 
@@
 \begin{figure}[ht]
-\includegraphics[width=\textwidth]{Waterfall.png}
+\includegraphics[width=\textwidth]{waterfall.eps}
 \caption{Conventional waterfall methodology for model development.
@@
 
-We feel that while many of the current software tools for genome scale models do an excellent job of facilitating answers to crucial research and design questions.
-However, their is a major gap in terms of the reliability and re-usability of models due to a lack of standardisation and software tools to aid such processes.
+We feel that many of the current software tools for genome scale models do an excellent job of facilitating answers to crucial research and design questions.
+However, there is a major gap in terms of the reliability and re-usability of models due to a lack of standardisation and software tools to aid such processes.
 The following sections provide an overview of the gsmodutils software framework.
@@
 
-By automatically integrating the COBRApy \cite{ebrahim2013cobrapy} users can easily write convenient test cases following examples given in the user guide.
-A standard test case, ensuring that a given model grows on media is given in Box \ref{fig:test_case}.
+By automatically integrating COBRApy \cite{ebrahim2013cobrapy} users can easily write convenient test cases following examples given in the user guide.
+A standard test case, ensuring that a given model grows on media is given in Figure \ref{fig:test_case}.
 When a new model repository is created with the gsmodutils tool, a number of pre-written test cases are automatically added to a file.
@@
 
-The software provides a number of features such as import an export of models in different formats and the generation of test reports through use of the command line. 
-The use of flat files for easy integration with version control software such as git and mercurial.
-
-In addition, projects are easy to export using portable standardised docker images \cite{merkel2014docker}. 
-The idea being to allow users to share models as quickly and easily as possible without concern for custom system configurations.
+The software provides a number of features such as import and export of models in different formats and the generation of test reports through use of the command line. 
+The use of flat files enables easy integration with version control software such as git and mercurial.
+In addition, projects are easy to export using portable standardised docker images \cite{merkel2014docker}, the idea being to allow users to share models as quickly and easily as possible without concern for custom system configurations.
 
@@
 Consequently, such designs are hereditary in nature.
-By taking the delta between the constraints applied to an initial model and subsequent modifications, gsmodutils allows users to easily reuse and export models with this design.
+By taking the \textit{model delta} between the constraints applied to an initial model and subsequent modifications, gsmodutils allows users to easily reuse and export models with this design.
 As designs inherit from a base model, future curation to a wild type model will automatically be included in the designs.
@@
 Figure \ref{fig:strain_hered} shows how this could work in a practical situation.
+
 Designs of this nature can also be programmatic, allowing the implementation of features such as non-standard constraints that can be dynamically loaded.
-Furtheremore, integrity of all strain designs is automatically tested as part of the default gsmodutils testing framework.
+An example of this is shown in Figure \ref{fig:programmatic_design}.
+This example converts an existing model to one based on an MILP, as the reaction names do not need to be specified, should reactions be added or removed the process will still apply.
+Alternative examples could include reductions of models through methods such as elementary flux modes or minimum cut sets, which can change dramatically with only small changes to stoichiometry.
+Furthermore, integrity of all strain designs is automatically tested as part of the default gsmodutils testing framework.
+
+\begin{figure}[ht]
+\lstset{language=Python}   
+\begin{lstlisting}
+def gsmdesign_example_milp(model, project):
+    for reaction in model.reactions:
+	reaction.variable_kind = "integer"
+    return model
+
+\end{lstlisting}
+\label{fig:programmatic_design} 
+\caption{An example gsmodutils programmatic design written in python. 
+This design converts reactions to integer type, allowing an MILP formation.}
+\end{figure}
 
@@
   Each design stores the delta between the wild-type base model, any parents and the changes to constraints the design contains.
-  In the example presented above a hetrologous production pathway is combined with a reusable set of knock-outs.
+  In the example presented above a heterologous production pathway is combined with a reusable set of knock-outs.
   In practice, ideally, these designs should relate to real constructs and strains evaluated by wet lab biologists.
@@
 The outlined workflow is that the user writes a formal test case for some modelling goal, perhaps driven by captured experimental data, that fits a specific form of validation criteria.
-We note that, in principle, this test case should be written \textit{a priori} of changes to a model.
-
-% \section{Example usage Clostridium acetobutylicum models}
-% \subsection{Scientific background}
-% \textit{C. acetobutylicum} ATCC 824 is a widely researched organism that has found use in the production of solvents for decades.
-% More recently, \textit{C. acetobutylicum} has been applied for the production of alcohols such as ethanol and n-butanols.
-% The industrial relevance and strong background of experimental data has led to the application of many genome scale models \cite{}.
-% Whilst increasingly informative when combined with experimental data, the history of the management of models for \textit{C. acetobutylicum} tells a different story.
-% 
-% Initially, two rival models were published containing a similar number of genes \cite{lee2008genome, senger2008genome}.
-% % \textbf{TODO: describe in more detail}.
-% %Discuss limitations of rival models
-% This led to the further development of subsequent models, iCaC 490 \cite{mcanulty2012genome} and an automatically generated model from  the Model SEED  pipeline \cite{}.
-% 
-% In this example we focus on the more recent contributions of iCac802 \cite{dash2014capturing} and iCac967 \cite{yoo2015quantitative}.
-% iCac802 combines a significantly updated set of annotations from previously curated model versions, an improved TCA cycle, full proton balancing in reactions and reactions for more complete coverage of hexadecanoyl-acp, hexadecanoyl-CoA, crotonyl-CoA pathways as well as pathways for the uptake and biosynthesis of more compounds.
-% Crucially, the model itself received significant validation against fluxonomic data from 13C-metabolic flux analysis ($13$C-MFA).
-% This additional validation set considerable flux bounds upon pathways, leading to improved curation.
-% 
-% Whilst iCac802 was informed by previous models, critically, it was started as a new curation project, a significant undertaking of work.
-% Such an approach may be necessary to uncover errors or oversights in construction processes.
-% However, this naturally adds a significant time burden for model curators and may be due to a lack of adaptability from previous models.
-% 
-% The iCac 967 model presents a smaller set of changes to the annotation of the \textit{C. acetobutylicum} model.
-% While the number of genes increased from 802 to 967 manual annotation actually reduced the number of reactions and metabolites from 1462 to 1137 and 1231 to 1058, respectively.
-% The annotation ane experimental evidence from iCac802 \cite{dash2014capturing} to iCac967 \cite{yoo2015quantitative} is greatly expanded.
-% However, it is not entirely clear, from the initial publication, the extent to which the expanded iCaC967 model preserves crucial experimental validation.
-% The following sections aim to check the later model against previous validation criteria.
-
-% \subsection{Model integration}
-% \textbf{Project construction:}
-% An initial project of the \textit{C. acetobutylicum} model was created in gsmodutils.
-% This is based around the iCac802 base model.
-% Included are also knockout strains $\Delta adc, \Delta ack, \Delta ptb$ and $\Delta hbd$ (TODO: describe knockouts).
-% 
-% In addition, a design listing the constraints changed in model version iCaC967 
-% \footnote{We would like to note that such a significant change to a model, in many practical cases, would not be considered a design but, rather, a separate model.
-% Separate models within the same project is a feature supported by gsmodutils.}.
-% \\
-% \textbf{Model comparison: TODO Diffs between existing models - use (and further develop) gsmodutils diff for model stats}.
-% 
-% \subsection{Evaluation of model validation criteria}
-% Several crucial validation criteria were identified from the experimental and theoretical analysis of the models found in the original publications.
-% We note that this is not an exhaustive set of criteria.
-% This section is designed to serve as a use case for the gsmodutils framework, not to comprehensively review the initial publications.
-% \\
-% 
-% \textbf{TCA cycle}\\
-% \textbf{Proton balanced results}\\
-% \textbf{Predictions of knockouts}\\
-% The inital publication for iCAC802 included validation of four gene knockout mutant strains.
-% $\Delta ack$ acetate kinase.\\
-% $\Delta ptb$ phosphotransbutrylase.\\
-% $\Delta adc$ acetoacetate decarboxylase. \textbf{This reaction is present and the knock out works as expected but not annotated correctly.}\\
-% $\Delta hdb$ hydroxybutyryl-CoA dehydrogenase. \\
-% 
-% \textbf{Flux validation against 13C data}\\
-% \textbf{Under acidogenic stress (H inhibited conditions) } \\
-% \textbf{Solvent formation Cell recycling in solventogenic phase}\\
+We note that, in principle, this test case should be written before changes to a model.
 
@@
 A new GSMM of \textit{C.
-autoethanogenum}, `MetaCLAU', has been analysed to improve this bioprocess (Norman et al., unpublished).
-
+autoethanogenum}, `MetaCLAU', has been analysed to improve this bioprocess (Norman et al., submitted for peer review).
 In this section we describe how gsmodutils has been utilised to ensure that future versions of MetaCLAU will remain functionally relevant from the perspective of industrial biotechnology.
@@
 Since carbon monoxide is readily available in the form of industrial waste gas, and 2,3BD has a global market value of \$43 billion \cite{kopke201123butanediol}, the optimisation of yields of 2,3BD from carbon monoxide is highly desirable in the context of industry \cite{norman2018progress}.
-
 MetaCLAU was built using Pathway Tools \cite{karp2002pathway} and ScrumPy \cite{poolman2006scrumpy}, and is based on a manually annotated genome sequence of \textit{C.
@@
 For full details of the model, see (Norman et al.
-unpublished)
-
+submitted for peer review).
 The model has been integrated with the gsodutils modelling framework as a test-driven project.
@@
 \subsection{Evaluation of model validation criteria}
-
 \textbf{Energetic consistency:} 
@@
 \textbf{Minimisation of flux based tests:}
-One conventional approach in FBA is to set an optimisation criteria of minimising flux across enzymatic reactions with a fixed biomass constraint \cite{holzhutter2006thegeneralised}. 
+One conventional approach in FBA is to set an optimisation criterion of minimising flux across enzymatic reactions with a fixed biomass constraint \cite{holzhutter2006thegeneralised}. 
 The solution to this FBA problem represents minimal protein investment \cite{holzhutter2006thegeneralised}.
@@
 1) the identification of essential genes is an important first step for metabolic engineering strategies, 
-and 2) with the advent of high-throughput TraDIS gene-essentiality data sets \cite{langridge2009simultaneous}, GSMMs can be validated by their abiity to predict essential genes.
+and 2) with the advent of high-throughput TraDIS gene-essentiality data sets \cite{langridge2009simultaneous}, GSMMs can be validated by their ability to predict essential genes.
 Furthermore, any change in the set of essential genes (particularly an increase in the number of essentials) represents important information for metabolic engineering.
-For these reasons, a test been built into the MetaCLAU project which enables the computation of the set of essential genes and their comparison with TraDIS data sets.
+For these reasons, a test has been built into the MetaCLAU project which enables the computation of the set of essential genes and their comparison with TraDIS data sets.
 
@@
 It features a fully specified set of tests, including custom test cases and has strong version control integration with git.
-A core difference between these projects is that gsmodutils has a stronger focus on reducing the redundancy in model storage through the usage of \textit{design deltas}, as described above.
+A core difference between these projects is that gsmodutils has a stronger focus on reducing the redundancy in model storage through the use of \textit{design deltas}, as described above.
 Similarly, a core goal of gsmodutils is to allow easy import and export outside of the framework for compatibility with other modelling suites.
@@
 Naturally a great deal of effort has gone in to producing high quality systems and synthetic biology standards \cite{hucka2003systems, cox2018synthetic}.
-Furthermore, when research projects end it all to is common for important large models to be published and become relics lost within the literature forgotten to all but the most dedicated of individuals.
+Furthermore, when research projects end it all to is common for important large models to be published and become relics lost within the literature, forgotten to all but the most dedicated of individuals.
 As GSMMs grow in terms of the metabolism the contain as well as the biological problems they are used to solve, problems with annotation and curation naturally accumulate as a product of human error.
@@
 Indeed, a core objective of this framework is to ensure that good practices are followed in model development that help scientists to better trust the results discovered by their models.
-In an ideal world, we would envision a methodology such as our becoming a pre-requisite for GSMMs to pass peer review.
+In an ideal world, we would envision a methodology such as ours becoming a pre-requisite for GSMMs to pass peer review.
 
@@
 We would like to thank the Oxford Brookes Cell Systems Modelling group for helpful discussions regarding this work and Thomas Millat for scientific advice.
-This work was supported by the UK Biotechnology and Biological Sciences Research Council (BBSRC) grant BB/L013940/1.
+This work was supported by the UK Biotechnology and Biological Sciences Research Council (BBSRC) grants BB/L013940/1,  BB/K00283X/1 and BB/L502030/1.
 

### 2018-08-29 article/article.tex
ACS journal format, fixed some refs

@@
-\documentclass[a4paper,11pt]{article}
+\documentclass[journal=asbcd6,10pt]{achemso}
 \usepackage[utf8]{inputenc}
@@
 \usepackage{color}
-\usepackage{authblk}
 \usepackage{listings}
@@
 %PREPRINT/MANUSCRIPT OPTIONS
-\usepackage{setspace}
+% \usepackage{setspace}
 \usepackage{lineno}
 \linenumbers
-\doublespacing
+% \doublespacing
 
@@
 
+\author{James P Gilbert}
+\email{james.gilbert@nottingham.ac.uk}
+\author{Nicole Pearcy}
+\author{Rupert Norman}
+\affiliation[sbrc]{Synthetic Biology Research Centre, University of Nottingham, Nottingham, NG7 2RD, United Kingdom}
+\author{Jamie Twycross}
+\affiliation[IMA]{Intelligent Modelling and Analysis group, School of Computer Science, University of Nottingham, Nottingham, NG8 1BB, United Kingdom}
+\author{John King}
+\affiliation[Maths]{School of Mathematical Sciences, University of Nottingham, Nottingham, NG7 2RD, United Kingdom}
+\author{Charlie Hodgman}
+\affiliation[Biosciences]{School of Biosciences, University of Nottingham, Sutton Bonington, Loughborough, LE12 5RD, United Kingdom}
+\author{Nigel Minton}
+\affiliation[sbrc]{Synthetic Biology Research Centre, University of Nottingham, Nottingham, NG7 2RD, United Kingdom}
+
 \title{Gsmodutils: A python based framework for test driven genome scale metabolic model development}
-\author[1,2]{James P Gilbert}
-\author[1,2]{Nicole Pearcy}
-\author[1]{Rupert Norman}
-\author[2]{Jamie Twycross}
-\author[3]{John King}
-\author[4]{Charlie Hodgman}
-\author[1]{Nigel Minton}
-\affil[1]{Synthetic Biology Research Centre, University of Nottingham, Nottingham, NG7 2RD, United Kingdom}
-\affil[2]{Intelligent Modelling and Analysis group, School of Computer Science, University of Nottingham, Nottingham, NG8 1BB, United Kingdom}
-\affil[3]{School of Mathematical Sciences, University of Nottingham, Nottingham, NG7 2RD, United Kingdom}
-\affil[4]{School of Biosciences, University of Nottingham, Sutton Bonington, Loughborough, LE12 5RD, United Kingdom}
 
 \begin{document}
+\include{contents}
 
-\maketitle
-
-\begin{abstract}
-\textbf{Motivation:} Genome scale metabolic models (GSMMs) are increasingly important for systems biology and metabolic engineering research as they are capable of simulating complex steady state phenotypic behaviour.
-Constraints based models of this form can include thousands of reactions and metabolites, with many crucial pathways that only become activated in specific simulation settings.
-However, despite their widespread use, power and the availability of tools to aid with the construction and analysis of large scale models little methodology is suggested for the management of continually curated large scale models.
-This is quickly becoming an issue for industrial systems and synthetic biotechnology applications, which require good quality reusable models integral to the design, build and test cycle.
-\\
-\textbf{Results:} 
-As part of an ongoing effort to improve genome scale metabolic analysis, we have developed a test driven development methodology for the continuous integration of validation data from different sources.
-Contributing to the open source technology based around COBRApy, we have developed the \textit{gsmodutils} modelling framework placing an emphasis on test driven design of models through defined test cases.
-Crucially, different substrate conditions are configurable allowing users to examine how different designs or curation impact a wide range of system behaviours, minimising error between model versions.\\
-\textbf{Availability:} The software framework described within this paper is open source and freely available from \url{http://github.com/SBRCNottingham/gsmodutils} \\
-\textbf{Contact:}{james.gilbert@nottingham.ac.uk}
-\end{abstract}
-
-\section{Introduction}
-Stoichiometric constraints based modelling for biological systems has been a mainstay of systems biology for several decades \cite{fell1986fat, varma1994stoichiometric}.
-Given its flexibility, low barrier to entry and requirement only on minimal knowledge regarding the stoichiometry of metabolic networks this structural approach has become an extremely popular method for modelling steady state behaviour of large, biochemical networks \cite{kauffman2003advances}.
-Such large scale reconstructions are often referred to as genome scale metabolic models (GSMMs), as the processes is significantly aided through the advent of relatively inexpensive genome sequencing \cite{o2015using, land2015insights}.
-Indeed, owing to their ability to model complex aspects of metabolism, GSMMs have been widely adopted as a standard to elucidate and optimise of industrial biotechnology processes \cite{kim2017current}.
-
-The reconstruction of GSMMs is an arduous process that follows a complex protocol to ensure model validity \cite{thiele2010protocol}.
-Whilst many popular automated methods exists to construct GSMMs from reference genomes \cite{henry2010high, poolman2006scrumpy}, there is a still a significant amount of manual curation.
-However, treating the creation of models as an isolated ``one-off'' event ignores the significant amount of curation that is required for applications such as biotechnology.
-
-As a consequence, a significant amount of work has gone into the management of genome scale models.
-The BiGG models database \cite{king2015bigg}, for example, exists to provide a standardised repository of validated models that can be shared and reused.
-However, little focus is placed upon the collaborative design aspect of such models with few mechanisms existing to store the \textit{model delta} based tools provided in our software.
-Similarly, the MetaNetX \cite{moretti2016metanetx} system exists to provide a standardised namespace and toolchain for GSM analysis.
-However, such tools often make it difficult to understand the design decisions made by the initial model authors.
-
-Furthermore, as with many areas of bioinformatic study the number available computational tools has become vast.
-This covers huge variety of software platforms including the COBRA toolbox in Matlab \cite{schellenberger2011quantitative}, ScrumPy and COBRApy in Python \cite{poolman2006scrumpy, ebrahim2013cobrapy} with additional tools and libraries such as cameo \cite{cardoso2017cameo}, OptFlux in Java \cite{rocha2010optflux} and SurreyFBA \cite{gevorgyan2010surreyfba}.
-Whilst most of these tools are Open Source and follow standards, such as SBML \cite{finney2003systems}, there are often few options that make analysis easy to replicate.
-The archetype design, build and test cycle of synthetic biology heavily relies on the usage of bioinformatics software and modelling to improve the production of natural products \cite{carbonell2016bioinformatics}.
-In order to speed up the use of bioinformatics tools to produce high value platform chemicals, genome scale models are often used to discover methods for process optimisation \textit{in silico}.
-
-For example, many tools such as RetroPath \cite{carbonell2013retropath}, XTMS \cite{carbonell2014xtms} and GEM-Path \cite{campodonico2014generation} suggest thousands of potential heterologous pathways as well as providing predictions for flux distributions.
-Many of these tools significantly increase the value of genome scale models, for example by coupling commodity production to an organism's growth \cite{feist2010model}.
-These tools all suggest major changes to wild-type strains must be tracked and compared to allow models to remain relevant.
-In effect, mechanisms are required to relate modified models to modified test and production strains.
-
-Similarly, many conventional applications of genome scale models in systems biology have often suffered from unnecessary replication of work due to a lack of standards.
-For example, there are now many independently developed models of \textit{Clostridium acetobutylicum} \cite{lee2008genome, senger2008genome, mcanulty2012genome, dash2014capturing, yoo2015quantitative}, an organism used in the production of solvents for over a century.
-These models all exist to solve similar biological problems some being updates to the inital base models. 
-However, there has been disagreement over fundamental biochemical properties of these anaerobic organisms, notably with the focus on redox balancing \cite{dash2014capturing}.
-Indeed, such models also include updates based on improved genome annotations and the inclusion of fluxomic, transcriptomic and metabolomic characterisations \cite{yoo2015quantitative}.
-
-Many of the results presented in original sources are difficult to compare due to a lack of standardisation of reference identifiers for reaction names.
-Where models are shared they are often in non-standard spreadsheet formats, rather than SBML models.
-Indeed, even in the case of valid SBML models being made available at the time of publication few details are given as to how to run such models.
-
-% Perhaps more worryingly, from an industrial perspective, is that the use of models in bioproduction has not escaped regulatory notice \cite{}.
-Furthermore, as models are often used to explain issues with online production and facilitate the optimisation of bioprocessing, it stands to reason that model correctness should attract oversight.
-Consequently, methods are needed to ensure that the models themselves are tightly version controlled.
-In this paper, we argue for the use of a form of \textit{test driven} model of development, a conceptual approach borrowed from software engineering.
-Given the availability of widespread open source virtualisation tools that allow the standardisation of platforms \cite{merkel2014docker}, modellers have the opportunity to share highly reproducible research \cite{chamberlain2014using}.
-
-In this paper we present a software framework geared toward \textit{test driven} genome scale model development, a concept that is taken directly from good software development practices.
-By this we mean the notion that, as a model is curated to represent biological phenomena, much of the validation can be turned into specific test cases that can be repeated between model versions.
-We provide an example test case for \textit{Clostridium autoethanogenum}, an organism that has had considerable focus in terms of genome scale models and how a working methodology using the software presented here can reduce repetition of work and improve the reproducibility of results.
-This article aims to summarise the main objectives of the gsmodutils software and we refer the reader to the software user guide for a more detailed exploration features.
-
-\section{Improving the design phase of industrial biotechnology}
-
-The design phase of industrial synthetic biotechnology, due to historical and regulatory convention, often follows a waterfall model for development of genome scale models.
-This is borne out of a number significant constraints in the development of genome scale models.
-The waterfall approach for modelling is characterised in Figure \ref{fig:methodology_approaches} where requirements for a model are captured at an early stage and are hard to refine during the production process.
-
-In contrast, an \textit{agile} methodology for the development of models places the focus on adapting work to new requirements \cite{martin2002agile}.
-Such an approach best fits genome scale models because they are rarely created to investigate individual processes and, instead, capture the complexity of large systems.
-Constraints based models are intrinsically related to available genome annotations.
-Such annotations rely heavily on automated matches to related species, with the characterisations of individual genes often being left to a few of critical interest \cite{seemann2014prokka}.
-This modelling formation costs an in depth understanding of dynamic behaviour. 
-However, capturing steady state phenomena still provides a good understanding of system properties \cite{o2015using}.
-
-As such approaches often leave models with missing reactions, incorrect gene-reaction rules \cite{thiele2010protocol} or with pathways based on gap filling methods that add reactions that may not be present within the organism \cite{benedict2014likelihood}.
-When attempting to understand specific natural phenomena, genome annotations are frequently updated and models are often corrected in an \textit{ad hoc} manner.
-
-\begin{figure}[ht]
-\includegraphics[width=\textwidth]{waterfall.eps}
-\caption{Conventional waterfall methodology for model development.
-Each process is considered to be an isolated aspect of model development.}
-\label{fig:methodology_approaches}
-\end{figure}
-
-Therefore, models undergo significant manual annotation - a process which naturally leads to errors.
-In this work we advocate a test driven approach to model development highlighted in Figure \ref{fig:test_driven}.
-Here, the model is changed to achieve core research goals that are dynamic in response to the changes of a project.
-In order to meet this objective, validation criteria for a model such as growth conditions or the impact of gene knock outs, should be formally set.
-When a model is changed, all such validation criteria should be retested to ensure that models do not regress to previous states.
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width=0.5\textwidth]{Testdriven.png}
-\caption{Workflow for model development proposed for gsmodutils.
-The objective of this approach is to simultaneously capture research questions, model validation criteria and minimise the impact of changes on previously completed models.
-}
-\label{fig:test_driven}
-\end{figure}
-
-We feel that many of the current software tools for genome scale models do an excellent job of facilitating answers to crucial research and design questions.
-However, there is a major gap in terms of the reliability and re-usability of models due to a lack of standardisation and software tools to aid such processes.
-The following sections provide an overview of the gsmodutils software framework.
-Gsmodutils aims to provide a basis for test driven, version controlled agile model development.
-All software is open source and is designed to be interoperable with platforms widely used in the domain of constraints based modelling.
-
-\section{Software}
-
-\subsection{Outline and features}
-Test driven development is driven by the idea of clearly defined test cases written before significant changes are made to any underlying architecture.
-In the case of genome scale models, errors occur as a product of human curation to better represent newly discovered aspects of metabolism.
-
-By automatically integrating COBRApy \cite{ebrahim2013cobrapy} users can easily write convenient test cases following examples given in the user guide.
-A standard test case, ensuring that a given model grows on media is given in Figure \ref{fig:test_case}.
-When a new model repository is created with the gsmodutils tool, a number of pre-written test cases are automatically added to a file.
-However, we stress that the vast majority of individual use cases for a model must be specific to a given biological problem.
-
-The software provides a number of features such as import and export of models in different formats and the generation of test reports through use of the command line. 
-The use of flat files enables easy integration with version control software such as git and mercurial.
-In addition, projects are easy to export using portable standardised docker images \cite{merkel2014docker}, the idea being to allow users to share models as quickly and easily as possible without concern for custom system configurations.
-
-\begin{figure}[ht]
-
-\lstset{language=Python}   
-\begin{lstlisting}
-def test_func(model, project, log):
-    status = model.solver.optimize()
-    log.assertion(status == 'optimal',
-                  ``Works'', # Success message
-		  ``Does not work'', # Failure message
-                  ``Test description'')
-\end{lstlisting}
-\label{fig:test_case} 
-\caption{An example gsmodutils test case written in python}
-\end{figure}
-
-\subsection{Strain designs}
-A core aspect behind the implementation of gsmodutils is the concept of a \textit{design}, this encompasses a simple set of changes to a ``wild-type'' model that are required for analysis.
-For example a mutant strain that has undergone several gene deletions is not sufficiently different from the original wild-type to merit having its own genome scale model.
-However, it is often the case that such deletions are of scientific or industrial interest and, as such, the strain will be used in future work.
-Consequently, such designs are hereditary in nature.
-By taking the \textit{model delta} between the constraints applied to an initial model and subsequent modifications, gsmodutils allows users to easily reuse and export models with this design.
-As designs inherit from a base model, future curation to a wild type model will automatically be included in the designs.
-Similarly, designs are self contained and will not interfere with one another allowing a form of project management.
-Figure \ref{fig:strain_hered} shows how this could work in a practical situation.
-
-Designs of this nature can also be programmatic, allowing the implementation of features such as non-standard constraints that can be dynamically loaded.
-An example of this is shown in Figure \ref{fig:programmatic_design}.
-This example converts an existing model to one based on an MILP, as the reaction names do not need to be specified, should reactions be added or removed the process will still apply.
-Alternative examples could include reductions of models through methods such as elementary flux modes or minimum cut sets, which can change dramatically with only small changes to stoichiometry.
-Furthermore, integrity of all strain designs is automatically tested as part of the default gsmodutils testing framework.
-
-\begin{figure}[ht]
-\lstset{language=Python}   
-\begin{lstlisting}
-def gsmdesign_example_milp(model, project):
-    for reaction in model.reactions:
-	reaction.variable_kind = "integer"
-    return model
-
-\end{lstlisting}
-\label{fig:programmatic_design} 
-\caption{An example gsmodutils programmatic design written in python. 
-This design converts reactions to integer type, allowing an MILP formation.}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\textwidth]{inheritence.png}
-  \label{fig:strain_hered}
-  \caption{Examples of gsmodutils design inheritance.
-  Each design stores the delta between the wild-type base model, any parents and the changes to constraints the design contains.
-  In the example presented above a heterologous production pathway is combined with a reusable set of knock-outs.
-  In practice, ideally, these designs should relate to real constructs and strains evaluated by wet lab biologists.
-  }
-\end{figure}
-
-\subsection{Development workflow}
-In this section we propose a method for the development of genome scale models that integrates gsmodutils with version control systems.
-Figure \ref{fig:test_case} highlights the notion of test cases, taken from test driven development.
-The outlined workflow is that the user writes a formal test case for some modelling goal, perhaps driven by captured experimental data, that fits a specific form of validation criteria.
-We note that, in principle, this test case should be written before changes to a model.
-
-\section{Case study usage Clostridium autoethanogenum}
-\textit{Clostridium autoethanogenum} is a bacterial species used for the production of commodity chemicals at industrial scale \cite{abrini1994clostridium,norman2018progress,LT}.
-A new GSMM of \textit{C.
-autoethanogenum}, `MetaCLAU', has been analysed to improve this bioprocess (Norman et al., submitted for peer review).
-In this section we describe how gsmodutils has been utilised to ensure that future versions of MetaCLAU will remain functionally relevant from the perspective of industrial biotechnology.
-
-\subsection{Scientific background and Model integration}
-
-\textit{C.
-autoethanogenum} is a strictly anaerobic, acetogenic bacterium which naturally produces ethanol and trace amounts of 2,3-butanediol (2,3BD) from carbon monoxide and water \cite{abrini1994clostridium,schuchmann2014autotrophy,norman2018progress}.
-Since carbon monoxide is readily available in the form of industrial waste gas, and 2,3BD has a global market value of \$43 billion \cite{kopke201123butanediol}, the optimisation of yields of 2,3BD from carbon monoxide is highly desirable in the context of industry \cite{norman2018progress}.
-MetaCLAU was built using Pathway Tools \cite{karp2002pathway} and ScrumPy \cite{poolman2006scrumpy}, and is based on a manually annotated genome sequence of \textit{C.
-autoethanogenum} \cite{humphreys2015whole}.
-The resulting model consisted of 758 reactions, 773 metabolites and 518 genes.
-For full details of the model, see (Norman et al.
-submitted for peer review).
-The model has been integrated with the gsodutils modelling framework as a test-driven project.
-The following section details specific tests used to evaluate the model at each stage of its continued development.
-
-\subsection{Evaluation of model validation criteria}
-\textbf{Energetic consistency:} 
-An important limitation of FBA is that optimal solutions may be thermodynamically infeasible if appropriate constraints are not applied \cite{Fell1197rohtua}.
-In order to identify these unwanted flux distributions and constrain the model such that they are not in the feasible soluton space, a diagnostic FBA is applied with the following constraints:
-1) All transport reactions are constrained to allow no uptake (lower bound = 0), and 2) the ATPase reaction is given a fixed flux of one.
-If a (non-zero) solution to this problem exists, it must contain a thermodynamic inconsistency, which can be dealt with by manual inspection of the solution and modification of one or more of the involved reactions \cite{Fell1197rohtua}.
-
-\textbf{Minimisation of flux based tests:}
-One conventional approach in FBA is to set an optimisation criterion of minimising flux across enzymatic reactions with a fixed biomass constraint \cite{holzhutter2006thegeneralised}. 
-The solution to this FBA problem represents minimal protein investment \cite{holzhutter2006thegeneralised}.
-Since execution of flux minimisation in COBRApy requires a model in which reversible reactions are split into two irreversible reactions, a gsmodutils design was created in which all reversible reactions are split in a programmatic manner.
-In the case of MetaCLAU, the flux minimal solution includes both ethanol and acetate production, which represents good qualitative agreement with experimental data (Norman et al., unpublished).
-Since any changes to this predicted phenotype must be investigated, the flux minimisation analysis has been formulated as a gsmuodutils test.
-
-\textbf{Product scans:}
-Of interest to this project were changes in the product spectrum of \textit{C.
-autoethanogenum} under conditions where the organism can and cannot produce molecular hydrogen (with carbon monoxide as sole carbon and energy source).
-The hypothesis tested in (Norman et al., unpublished), was that in the case where hydrogen production in infeasible, alternative electron sinks like lactate and 2,3BD would be produced.
-As in the previous case, the model-predicted behaviour showing both lactate and 2,3BD was deemed an important result, which model curators should be notified of if lost during model development.
-Thus the analysis was built into a gsmodutils test.
-
-\textbf{Lethal knockout mutants:} 
-The prediction of lethal single-gene KO mutants through FBA of a GSMM is useful in two ways: 
-1) the identification of essential genes is an important first step for metabolic engineering strategies, 
-and 2) with the advent of high-throughput TraDIS gene-essentiality data sets \cite{langridge2009simultaneous}, GSMMs can be validated by their ability to predict essential genes.
-Furthermore, any change in the set of essential genes (particularly an increase in the number of essentials) represents important information for metabolic engineering.
-For these reasons, a test has been built into the MetaCLAU project which enables the computation of the set of essential genes and their comparison with TraDIS data sets.
-
-% $\Delta \textit{nfn}$ knockout mutant: NOT SURE ABOUT THIS. CHRIS NOW THINKS IT'S ESSENTIAL :(
-These test cases are all fully reproducible and available as an example within the gsmodutils software repository and we encourage the reader to explore these examples to understand how to utilise gsmodutils within their own project.
-
-\section{Related work}
-\textbf{Memote:} \cite{lieven2017memote} is tool set that has very similar ambitions to gsmodutils.
-It features a fully specified set of tests, including custom test cases and has strong version control integration with git.
-A core difference between these projects is that gsmodutils has a stronger focus on reducing the redundancy in model storage through the use of \textit{design deltas}, as described above.
-Similarly, a core goal of gsmodutils is to allow easy import and export outside of the framework for compatibility with other modelling suites.
-It should be noted that, as memote is written in python and utilises COBRApy, it is fully compatible with gsmodutils.
-\\
-\textbf{Cameo:} \cite{cardoso2017cameo} is a set of strain design utilities, such as genetic algorithms for finding optimal gene knock-out
-strategies, as well as other flux balance analysis utilities. Written in python and part of the COBRApy suite of applications,
-gsmodutils integrates easily with cameo.
-Indeed, making a set of design changes can easily be integrated into gsmodutils through the python API.
-\\
-\textbf{Model repositories:} models are frequently shared, at the time of publication through services such as BiGG \cite{king2015bigg} and Biomodels \cite{chelliah2013biomodels}. 
-Whilst these repositories encourage the reuse of models and the reproducibility of \textit{in silico} predictions they are not designed to improved collaboration.
-The software presented here is designed with the notion that genome scale models are never finished, \textit{per se}, but under continuous development.
-The cornerstone of this is the use of test cases, which formalise modelling validation criteria.
-
-\section{Discussion}
-In order to facilitate the sharing and dissemination of high quality computational research, good standards and software are required \cite{jimenez2017four}.
-Naturally a great deal of effort has gone in to producing high quality systems and synthetic biology standards \cite{hucka2003systems, cox2018synthetic}.
-Furthermore, when research projects end it all to is common for important large models to be published and become relics lost within the literature, forgotten to all but the most dedicated of individuals.
-As GSMMs grow in terms of the metabolism the contain as well as the biological problems they are used to solve, problems with annotation and curation naturally accumulate as a product of human error.
-Software that facilitates actively improving how researchers develop and apply models to new phenomena is required.
-
-We have presented a framework with a number of features taken from the software development world specifically designed to improve collaboration and minimise such error.
-However, it is important to stress the difference between defined behaviour expected from pre-written test cases and novel predictions made by a model.
-Indeed, a core objective of this framework is to ensure that good practices are followed in model development that help scientists to better trust the results discovered by their models.
-In an ideal world, we would envision a methodology such as ours becoming a pre-requisite for GSMMs to pass peer review.
-
-As with most software development projects, gsmodutils will see expanded features.
-Initially this will include tighter integration with version control systems such as git and mercurial.
-Furthermore, the objective of the project is to cultivate collaboration by simplifying the process of distributing large models to different users.
-As a consequence, methods for sharing reusable docker containers will be investigated.
-
-\section*{Acknowledgements}
+\begin{acknowledgement}
 We would like to thank the Oxford Brookes Cell Systems Modelling group for helpful discussions regarding this work and Thomas Millat for scientific advice.
 This work was supported by the UK Biotechnology and Biological Sciences Research Council (BBSRC) grants BB/L013940/1,  BB/K00283X/1 and BB/L502030/1.
+\end{acknowledgement}
 
-\bibliographystyle{unsrt}
+\bibliographystyle{ieeetrans}
 \bibliography{references}
-
 \end{document}
-\grid

### 2018-08-29 article/contents.tex
Moved article content to different tex file

\maketitle

\begin{abstract}
\textbf{Motivation:} Genome scale metabolic models (GSMMs) are increasingly important for systems biology and metabolic engineering research as they are capable of simulating complex steady state phenotypic behaviour.
Constraints based models of this form can include thousands of reactions and metabolites, with many crucial pathways that only become activated in specific simulation settings.
However, despite their widespread use, power and the availability of tools to aid with the construction and analysis of large scale models little methodology is suggested for the management of continually curated large scale models.
This is quickly becoming an issue for industrial systems and synthetic biotechnology applications, which require good quality reusable models integral to the design, build and test cycle.
\\
\textbf{Results:} 
As part of an ongoing effort to improve genome scale metabolic analysis, we have developed a test driven development methodology for the continuous integration of validation data from different sources.
Contributing to the open source technology based around COBRApy, we have developed the \textit{gsmodutils} modelling framework placing an emphasis on test driven design of models through defined test cases.
Crucially, different substrate conditions are configurable allowing users to examine how different designs or curation impact a wide range of system behaviours, minimising error between model versions.\\
\textbf{Availability:} The software framework described within this paper is open source and freely available from \url{http://github.com/SBRCNottingham/gsmodutils} \\
\textbf{Contact:}{james.gilbert@nottingham.ac.uk}
\end{abstract}

\section{Introduction}
Stoichiometric constraints based modelling for biological systems has been a mainstay of systems biology for several decades \cite{fell1986fat, varma1994stoichiometric}.
Given its flexibility, low barrier to entry and requirement only on minimal knowledge regarding the stoichiometry of metabolic networks this structural approach has become an extremely popular method for modelling steady state behaviour of large, biochemical networks \cite{kauffman2003advances}.
Such large scale reconstructions are often referred to as genome scale metabolic models (GSMMs), as the processes is significantly aided through the advent of relatively inexpensive genome sequencing \cite{o2015using, land2015insights}.
Indeed, owing to their ability to model complex aspects of metabolism, GSMMs have been widely adopted as a standard to elucidate and optimise of industrial biotechnology processes \cite{kim2017current}.

The reconstruction of GSMMs is an arduous process that follows a complex protocol to ensure model validity \cite{thiele2010protocol}.
Whilst many popular automated methods exists to construct GSMMs from reference genomes \cite{henry2010high, poolman2006scrumpy}, there is a still a significant amount of manual curation.
However, treating the creation of models as an isolated ``one-off'' event ignores the significant amount of curation that is required for applications such as biotechnology.

As a consequence, a significant amount of work has gone into the management of genome scale models.
The BiGG models database \cite{king2015bigg}, for example, exists to provide a standardised repository of validated models that can be shared and reused.
However, little focus is placed upon the collaborative design aspect of such models with few mechanisms existing to store the \textit{model delta} based tools provided in our software.
Similarly, the MetaNetX \cite{moretti2016metanetx} system exists to provide a standardised namespace and toolchain for GSM analysis.
However, such tools often make it difficult to understand the design decisions made by the initial model authors.

Furthermore, as with many areas of bioinformatic study the number available computational tools has become vast.
This covers huge variety of software platforms including the COBRA toolbox in Matlab \cite{schellenberger2011quantitative}, ScrumPy and COBRApy in Python \cite{poolman2006scrumpy, ebrahim2013cobrapy} with additional tools and libraries such as cameo \cite{cardoso2017cameo}, OptFlux in Java \cite{rocha2010optflux} and SurreyFBA \cite{gevorgyan2010surreyfba}.
Whilst most of these tools are Open Source and follow standards, such as SBML \cite{finney2003systems}, there are often few options that make analysis easy to replicate.
The archetype design, build and test cycle of synthetic biology heavily relies on the usage of bioinformatics software and modelling to improve the production of natural products \cite{carbonell2016bioinformatics}.
In order to speed up the use of bioinformatics tools to produce high value platform chemicals, genome scale models are often used to discover methods for process optimisation \textit{in silico}.

For example, many tools such as RetroPath \cite{carbonell2013retropath}, XTMS \cite{carbonell2014xtms} and GEM-Path \cite{campodonico2014generation} suggest thousands of potential heterologous pathways as well as providing predictions for flux distributions.
Many of these tools significantly increase the value of genome scale models, for example by coupling commodity production to an organism's growth \cite{feist2010model}.
These tools all suggest major changes to wild-type strains must be tracked and compared to allow models to remain relevant.
In effect, mechanisms are required to relate modified models to modified test and production strains.

Similarly, many conventional applications of genome scale models in systems biology have often suffered from unnecessary replication of work due to a lack of standards.
For example, there are now many independently developed models of \textit{Clostridium acetobutylicum} \cite{lee2008genome, senger2008genome, mcanulty2012genome, dash2014capturing, yoo2015quantitative}, an organism used in the production of solvents for over a century.
These models all exist to solve similar biological problems some being updates to the inital base models. 
However, there has been disagreement over fundamental biochemical properties of these anaerobic organisms, notably with the focus on redox balancing \cite{dash2014capturing}.
Indeed, such models also include updates based on improved genome annotations and the inclusion of fluxomic, transcriptomic and metabolomic characterisations \cite{yoo2015quantitative}.

Many of the results presented in original sources are difficult to compare due to a lack of standardisation of reference identifiers for reaction names.
Where models are shared they are often in non-standard spreadsheet formats, rather than SBML models.
Indeed, even in the case of valid SBML models being made available at the time of publication few details are given as to how to run such models.

% Perhaps more worryingly, from an industrial perspective, is that the use of models in bioproduction has not escaped regulatory notice \cite{}.
Furthermore, as models are often used to explain issues with online production and facilitate the optimisation of bioprocessing, it stands to reason that model correctness should attract oversight.
Consequently, methods are needed to ensure that the models themselves are tightly version controlled.
In this paper, we argue for the use of a form of \textit{test driven} model of development, a conceptual approach borrowed from software engineering.
Given the availability of widespread open source virtualisation tools that allow the standardisation of platforms \cite{merkel2014docker}, modellers have the opportunity to share highly reproducible research \cite{chamberlain2014using}.

In this paper we present a software framework geared toward \textit{test driven} genome scale model development, a concept that is taken directly from good software development practices.
By this we mean the notion that, as a model is curated to represent biological phenomena, much of the validation can be turned into specific test cases that can be repeated between model versions.
We provide an example test case for \textit{Clostridium autoethanogenum}, an organism that has had considerable focus in terms of genome scale models and how a working methodology using the software presented here can reduce repetition of work and improve the reproducibility of results.
This article aims to summarise the main objectives of the gsmodutils software and we refer the reader to the software user guide for a more detailed exploration features.

\section{Improving the design phase of industrial biotechnology}

The design phase of industrial synthetic biotechnology, due to historical and regulatory convention, often follows a waterfall model for development of genome scale models.
This is borne out of a number significant constraints in the development of genome scale models.
The waterfall approach for modelling is characterised in Figure \ref{fig:methodology_approaches} where requirements for a model are captured at an early stage and are hard to refine during the production process.

In contrast, an \textit{agile} methodology for the development of models places the focus on adapting work to new requirements \cite{martin2002agile}.
Such an approach best fits genome scale models because they are rarely created to investigate individual processes and, instead, capture the complexity of large systems.
Constraints based models are intrinsically related to available genome annotations.
Such annotations rely heavily on automated matches to related species, with the characterisations of individual genes often being left to a few of critical interest \cite{seemann2014prokka}.
This modelling formation costs an in depth understanding of dynamic behaviour. 
However, capturing steady state phenomena still provides a good understanding of system properties \cite{o2015using}.

As such approaches often leave models with missing reactions, incorrect gene-reaction rules \cite{thiele2010protocol} or with pathways based on gap filling methods that add reactions that may not be present within the organism \cite{benedict2014likelihood}.
When attempting to understand specific natural phenomena, genome annotations are frequently updated and models are often corrected in an \textit{ad hoc} manner.

\begin{figure}[ht]
\includegraphics[width=\textwidth]{waterfall.eps}
\caption{Conventional waterfall methodology for model development.
Each process is considered to be an isolated aspect of model development.}
\label{fig:methodology_approaches}
\end{figure}

Therefore, models undergo significant manual annotation - a process which naturally leads to errors.
In this work we advocate a test driven approach to model development highlighted in Figure \ref{fig:test_driven}.
Here, the model is changed to achieve core research goals that are dynamic in response to the changes of a project.
In order to meet this objective, validation criteria for a model such as growth conditions or the impact of gene knock outs, should be formally set.
When a model is changed, all such validation criteria should be retested to ensure that models do not regress to previous states.

\begin{figure}[ht]
\centering
\includegraphics[width=0.5\textwidth]{Testdriven.png}
\caption{Workflow for model development proposed for gsmodutils.
The objective of this approach is to simultaneously capture research questions, model validation criteria and minimise the impact of changes on previously completed models.
}
\label{fig:test_driven}
\end{figure}

We feel that many of the current software tools for genome scale models do an excellent job of facilitating answers to crucial research and design questions.
However, there is a major gap in terms of the reliability and re-usability of models due to a lack of standardisation and software tools to aid such processes.
The following sections provide an overview of the gsmodutils software framework.
Gsmodutils aims to provide a basis for test driven, version controlled agile model development.
All software is open source and is designed to be interoperable with platforms widely used in the domain of constraints based modelling.

\section{Software}

\subsection{Outline and features}
Test driven development is driven by the idea of clearly defined test cases written before significant changes are made to any underlying architecture.
In the case of genome scale models, errors occur as a product of human curation to better represent newly discovered aspects of metabolism.

By automatically integrating COBRApy \cite{ebrahim2013cobrapy} users can easily write convenient test cases following examples given in the user guide.
A standard test case, ensuring that a given model grows on media is given in Figure \ref{fig:test_case}.
When a new model repository is created with the gsmodutils tool, a number of pre-written test cases are automatically added to a file.
However, we stress that the vast majority of individual use cases for a model must be specific to a given biological problem.

The software provides a number of features such as import and export of models in different formats and the generation of test reports through use of the command line. 
The use of flat files enables easy integration with version control software such as git and mercurial.
In addition, projects are easy to export using portable standardised docker images \cite{merkel2014docker}, the idea being to allow users to share models as quickly and easily as possible without concern for custom system configurations.

\begin{figure}[ht]

\lstset{language=Python}   
\begin{lstlisting}
def test_func(model, project, log):
    status = model.solver.optimize()
    log.assertion(status == 'optimal',
                  ``Works'', # Success message
		  ``Does not work'', # Failure message
                  ``Test description'')
\end{lstlisting}
\label{fig:test_case} 
\caption{An example gsmodutils test case written in python}
\end{figure}

\subsection{Strain designs}
A core aspect behind the implementation of gsmodutils is the concept of a \textit{design}, this encompasses a simple set of changes to a ``wild-type'' model that are required for analysis.
For example a mutant strain that has undergone several gene deletions is not sufficiently different from the original wild-type to merit having its own genome scale model.
However, it is often the case that such deletions are of scientific or industrial interest and, as such, the strain will be used in future work.
Consequently, such designs are hereditary in nature.
By taking the \textit{model delta} between the constraints applied to an initial model and subsequent modifications, gsmodutils allows users to easily reuse and export models with this design.
As designs inherit from a base model, future curation to a wild type model will automatically be included in the designs.
Similarly, designs are self contained and will not interfere with one another allowing a form of project management.
Figure \ref{fig:strain_hered} shows how this could work in a practical situation.

Designs of this nature can also be programmatic, allowing the implementation of features such as non-standard constraints that can be dynamically loaded.
An example of this is shown in Figure \ref{fig:programmatic_design}.
This example converts an existing model to one based on an MILP, as the reaction names do not need to be specified, should reactions be added or removed the process will still apply.
Alternative examples could include reductions of models through methods such as elementary flux modes or minimum cut sets, which can change dramatically with only small changes to stoichiometry.
Furthermore, integrity of all strain designs is automatically tested as part of the default gsmodutils testing framework.

\begin{figure}[ht]
\lstset{language=Python}   
\begin{lstlisting}
def gsmdesign_example_milp(model, project):
    for reaction in model.reactions:
	reaction.variable_kind = "integer"
    return model

\end{lstlisting}
\label{fig:programmatic_design} 
\caption{An example gsmodutils programmatic design written in python. 
This design converts reactions to integer type, allowing an MILP formation.}
\end{figure}

\begin{figure}
  \includegraphics[width=\textwidth]{inheritence.png}
  \label{fig:strain_hered}
  \caption{Examples of gsmodutils design inheritance.
  Each design stores the delta between the wild-type base model, any parents and the changes to constraints the design contains.
  In the example presented above a heterologous production pathway is combined with a reusable set of knock-outs.
  In practice, ideally, these designs should relate to real constructs and strains evaluated by experiment.
  }
\end{figure}

\subsection{Development workflow}
In this section we propose a method for the development of genome scale models that integrates gsmodutils with version control systems.
Figure \ref{fig:test_case} highlights the notion of test cases, taken from test driven development.
The outlined workflow is that the user writes a formal test case for some modelling goal, perhaps driven by captured experimental data, that fits a specific form of validation criteria.
We note that, in principle, this test case should be written before changes to a model.

\section{Case study usage Clostridium autoethanogenum}
\textit{Clostridium autoethanogenum} is a bacterial species used for the production of commodity chemicals at industrial scale \cite{abrini1994clostridium,norman2018progress,LT}. A new GSMM of \textit{C. autoethanogenum}, `MetaCLAU', has been analysed to improve this bioprocess (Norman et al., submitted for peer review).
In this section we describe how gsmodutils has been utilised to ensure that future versions of MetaCLAU will remain functionally relevant from the perspective of industrial biotechnology.

\subsection{Scientific background and Model integration}

\textit{C. autoethanogenum} is a strictly anaerobic, acetogenic bacterium which naturally produces ethanol and trace amounts of 2,3-butanediol (2,3BD) from carbon monoxide and water \cite{abrini1994clostridium,schuchmann2014autotrophy,norman2018progress}. Since carbon monoxide is readily available in the form of industrial waste gas, and 2,3BD has a global market value of \$43 billion \cite{kopke201123butanediol}, the optimisation of yields of 2,3BD from carbon monoxide is highly desirable in the context of industry \cite{norman2018progress}.   
MetaCLAU was built using Pathway Tools \cite{karp2002pathway} and ScrumPy \cite{poolman2006scrumpy}, and is based on a manually annotated genome sequence of \textit{C. autoethanogenum} \cite{humphreys2015whole}. The resulting model consisted of 758 reactions, 773 metabolites and 518 genes. For full details of the model, see (Norman et al. submitted for peer review).
The model has been integrated with the gsodutils modelling framework as a test-driven project. The following section details specific tests used to evaluate the model at each stage of its continued development.

\subsection{Evaluation of model validation criteria}
\textbf{Energetic consistency:} 
An important limitation of FBA is that optimal solutions may be thermodynamically infeasible if appropriate constraints are not applied \cite{Fell1197rohtua}.
In order to identify these unwanted flux distributions and constrain the model such that they are not in the feasible soluton space, a diagnostic FBA is applied with the following constraints:
1) All transport reactions are constrained to allow no uptake (lower bound = 0), and 2) the ATPase reaction is given a fixed flux of one.
If a (non-zero) solution to this problem exists, it must contain a thermodynamic inconsistency, which can be dealt with by manual inspection of the solution and modification of one or more of the involved reactions \cite{Fell1197rohtua}.

\textbf{Minimisation of flux based tests:}
One conventional approach in FBA is to set an optimisation criterion of minimising flux across enzymatic reactions with a fixed biomass constraint \cite{holzhutter2006thegeneralised}. 
The solution to this FBA problem represents minimal protein investment \cite{holzhutter2006thegeneralised}.
Since execution of flux minimisation in COBRApy requires a model in which reversible reactions are split into two irreversible reactions, a gsmodutils design was created in which all reversible reactions are split in a programmatic manner.
In the case of MetaCLAU, the flux minimal solution includes both ethanol and acetate production, which represents good qualitative agreement with experimental data (Norman et al., unpublished).
Since any changes to this predicted phenotype must be investigated, the flux minimisation analysis has been formulated as a gsmuodutils test.

\textbf{Product scans:}
Of interest to this project were changes in the product spectrum of \textit{C. autoethanogenum} under conditions where the organism can and cannot produce molecular hydrogen (with carbon monoxide as sole carbon and energy source).
The hypothesis tested in (Norman et al., unpublished), was that in the case where hydrogen production in infeasible, alternative electron sinks like lactate and 2,3BD would be produced.
As in the previous case, the model-predicted behaviour showing both lactate and 2,3BD was deemed an important result, which model curators should be notified of if lost during model development.
Thus the analysis was built into a gsmodutils test.

\textbf{Lethal knockout mutants:} 
The prediction of lethal single-gene KO mutants through FBA of a GSMM is useful in two ways: 
1) the identification of essential genes is an important first step for metabolic engineering strategies, 
and 2) with the advent of high-throughput TraDIS gene-essentiality data sets \cite{langridge2009simultaneous}, GSMMs can be validated by their ability to predict essential genes.
Furthermore, any change in the set of essential genes (particularly an increase in the number of essentials) represents important information for metabolic engineering.
For these reasons, a test has been built into the MetaCLAU project which enables the computation of the set of essential genes and their comparison with TraDIS data sets.

% $\Delta \textit{nfn}$ knockout mutant: NOT SURE ABOUT THIS. CHRIS NOW THINKS IT'S ESSENTIAL :(
These test cases are all fully reproducible and available as an example within the gsmodutils software repository and we encourage the reader to explore these examples to understand how to utilise gsmodutils within their own project.

\section{Related work}
\textbf{Memote:} \cite{lieven2017memote} is tool set that has very similar ambitions to gsmodutils.
It features a fully specified set of tests, including custom test cases and has strong version control integration with git.
A core difference between these projects is that gsmodutils has a stronger focus on reducing the redundancy in model storage through the use of \textit{design deltas}, as described above.
Similarly, a core goal of gsmodutils is to allow easy import and export outside of the framework for compatibility with other modelling suites.
It should be noted that, as memote is written in python and utilises COBRApy, it is fully compatible with gsmodutils.
\\
\textbf{Cameo:} \cite{cardoso2017cameo} is a set of strain design utilities, such as genetic algorithms for finding optimal gene knock-out
strategies, as well as other flux balance analysis utilities. Written in python and part of the COBRApy suite of applications,
gsmodutils integrates easily with cameo.
Indeed, making a set of design changes can easily be integrated into gsmodutils through the python API.
\\
\textbf{Model repositories:} models are frequently shared, at the time of publication through services such as BiGG \cite{king2015bigg} and Biomodels \cite{chelliah2013biomodels}. 
Whilst these repositories encourage the reuse of models and the reproducibility of \textit{in silico} predictions they are not designed to improved collaboration.
The software presented here is designed with the notion that genome scale models are never finished, \textit{per se}, but under continuous development.
The cornerstone of this is the use of test cases, which formalise modelling validation criteria.

\section{Discussion}
In order to facilitate the sharing and dissemination of high quality computational research, good standards and software are required \cite{jimenez2017four}.
Naturally a great deal of effort has gone in to producing high quality systems and synthetic biology standards \cite{hucka2003systems, cox2018synthetic}.
Furthermore, when research projects end it all to is common for important large models to be published and become relics lost within the literature, forgotten to all but the most dedicated of individuals.
As GSMMs grow in terms of the metabolism the contain as well as the biological problems they are used to solve, problems with annotation and curation naturally accumulate as a product of human error.
Software that facilitates actively improving how researchers develop and apply models to new phenomena is required.

We have presented a framework with a number of features taken from the software development world specifically designed to improve collaboration and minimise such error.
However, it is important to stress the difference between defined behaviour expected from pre-written test cases and novel predictions made by a model.
Indeed, a core objective of this framework is to ensure that good practices are followed in model development that help scientists to better trust the results discovered by their models.
In an ideal world, we would envision a methodology such as ours becoming a pre-requisite for GSMMs to pass peer review.

As with most software development projects, gsmodutils will see expanded features.
Initially this will include tighter integration with version control systems such as git and mercurial.
Furthermore, the objective of the project is to cultivate collaboration by simplifying the process of distributing large models to different users.
As a consequence, methods for sharing reusable docker containers will be investigated.

### 2018-08-29 gsmodutils/utils/scrumpy.py
Added tox testing

@@ -5,9 +5,8 @@ import json
 import os
 
 import cobra
-#from gsmodutils.utils.io import load_medium
 import re
-from io import StringIO
+from six import StringIO
 
 
 class ParseError(Exception):
@@ -75,7 +74,8 @@ def load_scrumpy_model(filepath_or_string, name=None, model_id=None, media=None,
             added_reactions.append(r.id)
 
     if media is not None:
-        load_medium(model, media)
+        for ex_reaction in model.exchanges:
+            ex_reaction.lower_bound = media.get(ex_reaction.id, 0)
 
     if fixed_fluxes is not None:
         for rid, flux in fixed_fluxes.items():

### 2018-08-30 article/article.tex
EPS figures

--- 2018-08-13 article/article.tex
@@
 
+\author{James P Gilbert}
+\email{james.gilbert@nottingham.ac.uk}
+\author{Nicole Pearcy}
+\author{Rupert Norman}
+\affiliation[sbrc]{Synthetic Biology Research Centre, University of Nottingham, Nottingham, NG7 2RD, United Kingdom}
+\author{Jamie Twycross}
+\affiliation[IMA]{Intelligent Modelling and Analysis group, School of Computer Science, University of Nottingham, Nottingham, NG8 1BB, United Kingdom}
+\author{John King}
+\affiliation[Maths]{School of Mathematical Sciences, University of Nottingham, Nottingham, NG7 2RD, United Kingdom}
+\author{Charlie Hodgman}
+\affiliation[Biosciences]{School of Biosciences, University of Nottingham, Sutton Bonington, Loughborough, LE12 5RD, United Kingdom}
+\author{Nigel Minton}
+\affiliation[sbrc]{Synthetic Biology Research Centre, University of Nottingham, Nottingham, NG7 2RD, United Kingdom}
+
 \title{Gsmodutils: A python based framework for test driven genome scale metabolic model development}
-\author[1,2]{James P Gilbert}
-\author[1,2]{Nicole Pearcy}
-\author[1]{Rupert Norman}
-\author[2]{Jamie Twycross}
-\author[3]{John King}
-\author[4]{Charlie Hodgman}
-\author[1]{Nigel Minton}
-\affil[1]{Synthetic Biology Research Centre, University of Nottingham, Nottingham, NG7 2RD, United Kingdom}
-\affil[2]{Intelligent Modelling and Analysis group, School of Computer Science, University of Nottingham, Nottingham, NG8 1BB, United Kingdom}
-\affil[3]{School of Mathematical Sciences, University of Nottingham, Nottingham, NG7 2RD, United Kingdom}
-\affil[4]{School of Biosciences, University of Nottingham, Sutton Bonington, Loughborough, LE12 5RD, United Kingdom}
 
 \begin{document}
-
 \maketitle
@@
 \centering
-\includegraphics[width=0.5\textwidth]{Testdriven.png}
+\includegraphics[width=0.8\textwidth]{testdriven.eps}
 \caption{Workflow for model development proposed for gsmodutils.
@@
 \begin{figure}
-  \includegraphics[width=\textwidth]{inheritence.png}
+  \includegraphics[width=\textwidth]{inheritence2.eps}
   \label{fig:strain_hered}
@@
   In the example presented above a heterologous production pathway is combined with a reusable set of knock-outs.
-  In practice, ideally, these designs should relate to real constructs and strains evaluated by wet lab biologists.
+  In practice, ideally, these designs should relate to real constructs and strains evaluated by experiment.
   }
@@
 
-\section*{Acknowledgements}
+\begin{acknowledgement}
 We would like to thank the Oxford Brookes Cell Systems Modelling group for helpful discussions regarding this work and Thomas Millat for scientific advice.
 This work was supported by the UK Biotechnology and Biological Sciences Research Council (BBSRC) grants BB/L013940/1,  BB/K00283X/1 and BB/L502030/1.
+\end{acknowledgement}
 
-\bibliographystyle{unsrt}
+\bibliographystyle{ieeetrans}
 \bibliography{references}
-
 \end{document}
-\grid

### 2018-08-30 tests/test_designs.py
Fixed error in unit test

@@ -180,5 +180,5 @@ gsmdesign_invalid_parent_id.parent = "thanos"
         assert mdl3.reactions.get_by_id("RBPC").lower_bound == 0
 
         # This design shouldn't load as the parent id ref is wrong
-        with pytest.raises(DesignOrphanError):
+        with pytest.raises(DesignError):
             project.get_design("t1_invalid_parent_id")

### 2018-08-30 gsmodutils/project/interface.py
Fixed large bug with designs not being reloaded if stored in project context. This was caused by an issue with the usage of GSMProject._designs_store which caches designs for speedups (particularly with complex heritage). Removing this fixes the bug as a workaround. A longer term solution should be to use file timestamps.

@@ -290,16 +290,14 @@ class GSMProject(object):
         if design not in self.list_designs:
             raise DesignNotFoundError("Design of name {} not found in project".format(design))
 
-        if design not in self._designs_store:
-
-            if design in self._json_designs:
-                des_path = os.path.join(self._project_path, self.config.design_dir, '{}.json'.format(design))
-                self._designs_store[design] = StrainDesign.from_json(design, des_path, self)
-            else:
-                try:
-                    self._designs_store[design] = self._load_py_design(design)
-                except Exception as exp:
-                    raise DesignError(str(exp))
+        if design in self._json_designs:
+            des_path = os.path.join(self._project_path, self.config.design_dir, '{}.json'.format(design))
+            self._designs_store[design] = StrainDesign.from_json(design, des_path, self)
+        else:
+            try:
+                self._designs_store[design] = self._load_py_design(design)
+            except Exception as exp:
+                raise DesignError(str(exp))
 
         return self._designs_store[design]
 

### 2018-08-30 tests/test_cli.py

@@ -204,6 +204,9 @@ def test_import_designs():
             '--parent', did
         ]
 
+        name = 'test m'
+        description = 'test desc'
+
         result = runner.invoke(gsmodutils.cli.dimport, opts, input='{}\n{}\n'.format(name, description))
 
         assert result.exit_code == 0
@@ -221,25 +224,23 @@ def test_import_designs():
         assert cdes.name == name
         assert cdes.description == description
 
+        desc_n = "overwrite valid test"
+        name_n = "overwrite"
+        # Test overwrite - valid
         opts = [
             save_path, cdid,
             '--project_path', ctx.path,
             '--parent', did,
+            '--description', desc_n,
+            '--name', name_n,
             '--overwrite'
         ]
-
-        result = runner.invoke(gsmodutils.cli.dimport, opts, input='{}\n{}\n'.format(name, description))
+        result = runner.invoke(gsmodutils.cli.dimport, opts)
         assert result.exit_code == 0
 
-        opts = [
-            save_path, cdid,
-            '--project_path', ctx.path,
-            '--parent', did
-        ]
-
-        # Test overwrite existing fail and pass
-        result = runner.invoke(gsmodutils.cli.dimport, opts)
-        assert result.exit_code == -2
+        odes = project.get_design(cdid)
+        assert odes.name == name_n
+        assert odes.description == desc_n
 
         # Test non-existent parent
         opts = [
@@ -249,8 +250,19 @@ def test_import_designs():
         ]
 
         result = runner.invoke(gsmodutils.cli.dimport, opts)
-        assert result.exit_code == -1
+        assert result.exit_code == -3
         assert 'should_fail' not in project.list_designs
+
+        # Test overwrite existing fail and pass
+        opts = [
+            save_path, cdid,
+            '--project_path', ctx.path,
+            '--parent', did
+        ]
+
+        # Test overwrite existing fail and pass
+        result = runner.invoke(gsmodutils.cli.dimport, opts)
+        assert result.exit_code == -2
 
 
 def test_addmodel():

### 2018-08-31 gsmodutils/cli.py
Fixed bug in test cli, handling of bad tests

@@ -161,14 +161,12 @@ def test(project_path, test_id, skip_default, verbose, log_path):
             click.style('Error - {} has syntax errors {}'.format(tf, e), fg='red')
         )
 
-    for tf, entry_key, missing_fields in tester.invalid_tests:
+    for tf, entry_key, exception in tester.invalid_tests:
         click.echo(
             click.style('---Error with formating of test {} in {} ---'.format(tf, entry_key), fg='yellow')
         )
 
-        click.echo('Missing fields:')
-        for field in missing_fields:
-            click.echo('\t{}'.format(field))
+        click.echo('\t{}'.format(exception))
 
     # Save report to json log file
     if log_path is not None:

### 2018-08-31 gsmodutils/cli.py
Better display for badly formatted tests

@@ -166,7 +166,8 @@ def test(project_path, test_id, skip_default, verbose, log_path):
             click.style('---Error with formating of test {} in {} ---'.format(tf, entry_key), fg='yellow')
         )
 
-        click.echo('\t{}'.format(exception))
+        click.echo('\tError found in file:')
+        click.echo('\t{}'.format(exception.message))
 
     # Save report to json log file
     if log_path is not None:

### 2018-08-31 gsmodutils/test/tester.py
Better display for badly formatted tests

@@ -4,11 +4,11 @@ import json
 import os
 import sys
 from collections import defaultdict
-from io import StringIO
 
 from cobra.core import get_solution
 from cobra.exceptions import Infeasible
 from six import exec_
+from gsmodutils.utils import StringIO
 
 import gsmodutils
 from gsmodutils.test.utils import TestRecord

### 2018-08-31 gsmodutils/utils/__init__.py

@@ -1,6 +1,17 @@
 import cobra
 from cobra.exceptions import Infeasible
-from six import string_types
+from six import string_types, StringIO
+import sys
+
+# Hack for python version 2 to allow context management of StringIO
+if sys.version_info[0] == 2:
+
+    class StringIO(StringIO):
+        def __enter__(self):
+            return self
+
+        def __exit__(self, *args):
+            self.close()
 
 
 class FrozenDict(dict):

### 2018-08-31 tests/test_tests.py

@@ -14,6 +14,7 @@ from gsmodutils.utils.io import load_medium
 from click.testing import CliRunner
 import gsmodutils.cli
 from gsmodutils.test.tester import stdout_ctx
+from six import exec_
 
 
 def test_json_tests():
@@ -249,6 +250,6 @@ def test_std_out_capture_in_exec():
 
     with stdout_ctx() as context:
         compiled_code = compile(exec_code, '', 'exec')
-        exec(compiled_code, {})
+        exec_(compiled_code, {})
 
     assert context.getvalue() == output_message+"\n"

### 2018-09-24 gsmodutils/project/design.py
Removed redundant code

@@ -538,10 +538,6 @@ class StrainDesign(object):
             except Exception as ex:
                 raise DesignError("Function execution error {}".format(ex))
 
-            if not isinstance(model, cobra.Model):
-                raise DesignError(
-                    "Function does not return cobra.Model instance. Got {} type instead".format(type(model)))
-
             return mdl
 
         # Add new or changed metabolites to model

### 2018-09-24 tests/test_model.py

@@ -1,14 +1,37 @@
 from tutils import FakeProjectContext
 from gsmodutils import GSMProject, GSModutilsModel
 import cobra
+import pytest
 
 
 def test_load_model():
+    """
+    Most of this is for code coverage
+    :return:
+    """
     with FakeProjectContext() as ctx:
+        ctx.add_fake_conditions()
+        ctx.add_fake_designs()
+
         project = GSMProject(ctx.path)
         assert project.project_path == ctx.path
         model = GSModutilsModel(project)
         assert isinstance(model, cobra.Model)
+
+        model.load_conditions("xyl_src")
+
+        model.diff()
+        model.diff(GSModutilsModel(project, design="cbb_cycle"))
+
+        with pytest.raises(TypeError):
+            model.diff("should break")
+
+        # Test loading non design fails
+        with pytest.raises(TypeError):
+            model = project.model
+            model.set_design({})
+
+        model.save_model()
 
 
 def test_copy():

### 2018-09-24 tests/test_scrumpy.py

@@ -1,4 +1,5 @@
 from gsmodutils.utils.io import load_model
+from gsmodutils.utils.scrumpy import load_scrumpy_model
 from tutils import scrumpy_model_path, scrumpy_biomass_path
 import tempfile
 import cobra
@@ -7,6 +8,7 @@ import gsmodutils
 import os
 import string
 import random
+import pytest
 
 
 def test_cli_tool():
@@ -41,3 +43,12 @@ def test_cli_tool():
     assert len(model.metabolites) == 1130 # 1129
     # Number of transporters + metabolites should equal scrumpy's number of metabolites
     assert len(model.exchanges) == 71  # 67
+
+
+def test_load_model():
+    # Test loading of scrumpy models
+    model = load_scrumpy_model(scrumpy_model_path, media={"NH3_tx": -1000})
+    assert isinstance(model, cobra.Model)
+
+    with pytest.raises(KeyError):
+        load_scrumpy_model(scrumpy_model_path, fixed_fluxes={"NOTREAL": 1000})

### 2018-09-24 article/article.tex
Article changes

@@
 \affiliation[sbrc]{Synthetic Biology Research Centre, University of Nottingham, Nottingham, NG7 2RD, United Kingdom}
-\author{Jamie Twycross}
-\affiliation[IMA]{Intelligent Modelling and Analysis group, School of Computer Science, University of Nottingham, Nottingham, NG8 1BB, United Kingdom}
+\author{Thomas Millat}
+\affiliation[sbrc]{Synthetic Biology Research Centre, University of Nottingham, Nottingham, NG7 2RD, United Kingdom}
+\author{Klaus Winzer}
+\affiliation[sbrc]{Synthetic Biology Research Centre, University of Nottingham, Nottingham, NG7 2RD, United Kingdom}
 \author{John King}
@@
 \affiliation[sbrc]{Synthetic Biology Research Centre, University of Nottingham, Nottingham, NG7 2RD, United Kingdom}
+\author{Jamie Twycross}
+\affiliation[IMA]{Intelligent Modelling and Analysis group, School of Computer Science, University of Nottingham, Nottingham, NG8 1BB, United Kingdom}
 
@@
 \begin{acknowledgement}
-We would like to thank the Oxford Brookes Cell Systems Modelling group for helpful discussions regarding this work and Thomas Millat for scientific advice.
+We would like to thank the Oxford Brookes Cell Systems Modelling group for helpful discussions regarding this work and for scientific advice.
 This work was supported by the UK Biotechnology and Biological Sciences Research Council (BBSRC) grants BB/L013940/1,  BB/K00283X/1 and BB/L502030/1.

### 2018-09-24 tests/test_diff.py
Code coverage for diff html display in jupyter

@@ -16,6 +16,7 @@ def test_model_ident():
     assert len(diff['removed_metabolites']) == 0
     assert len(diff['reactions']) == 0
     assert len(diff['metabolites']) == 0
+    diff._repr_html_()
 
 
 def test_reaction_lb_change():
@@ -34,6 +35,7 @@ def test_reaction_lb_change():
     assert len(diff['reactions']) == 1
     assert diff['reactions'][0]['id'] == model_b.reactions.ATPM.id
     assert diff['reactions'][0]['lower_bound'] == model_b.reactions.ATPM.lower_bound
+    diff._repr_html_()
 
 
 def test_metabolite_formula_change():
@@ -55,6 +57,7 @@ def test_metabolite_formula_change():
     assert len(diff['metabolites']) == 1
     assert diff['metabolites'][0]['id'] == model_a.metabolites.h2o_c.id
     assert diff['metabolites'][0]['formula'] == model_b.metabolites.h2o_c.formula
+    diff._repr_html_()
 
 
 def test_gene_removal():
@@ -65,6 +68,7 @@ def test_gene_removal():
     diff = gsmodutils.model_diff.model_diff(model_a, model_b)
 
     assert len(diff['removed_genes']) == 1
+    diff._repr_html_()
 
     model_a = load_model(_CORE_MODEL_PATH)
     model_b = load_model(_CORE_MODEL_PATH)
@@ -72,6 +76,7 @@ def test_gene_removal():
     cobra.manipulation.remove_genes(model_a, ['b0008'], remove_reactions=False)
     diff = gsmodutils.model_diff.model_diff(model_a, model_b)
     assert len(diff['genes']) == 1
+    diff._repr_html_()
 
 
 def test_model_error():

### 2018-09-24 gsmodutils/utils/scrumpy.py
Cyclic scipy files unittest

@@ -41,16 +41,19 @@ def load_scrumpy_model(filepath_or_string, name=None, model_id=None, media=None,
     if os.path.isfile(filepath_or_string):
         rel_path = '/'.join(os.path.abspath(filepath_or_string).split('/')[:-1])
         fp = os.path.abspath(filepath_or_string).split('/')[-1]
-        reactions, metabolites = parse_file(fp, rel_path=rel_path)
+        reactions, metabolites, externals = parse_file(fp, rel_path=rel_path)
     else:
         rel_path = '.'
-        reactions, metabolites = parse_string(filepath_or_string, rel_path=rel_path)
+        reactions, metabolites, externals = parse_string(filepath_or_string, rel_path=rel_path)
 
     model = cobra.Model()
     for mid in metabolites:
-        m = cobra.Metabolite(id=mid)
+        compartment = 'e'
+        if mid[:2] == "x_" or mid in externals:
+            compartment = 'e'
+        m = cobra.Metabolite(id=mid, compartment=compartment)  # ScrumPy does not use compartments
         model.add_metabolites([m])
-    
+
     added_reactions = []
     for reaction in reactions:
         if reaction['id'] not in added_reactions:
@@ -64,7 +67,6 @@ def load_scrumpy_model(filepath_or_string, name=None, model_id=None, media=None,
     # We need to add transporters for external metabolites not defined with the "External" directive
     for metabolite in model.metabolites:
         if metabolite.id[:2] == "x_":
-
             r = cobra.Reaction("EX_{}".format(metabolite.id[2:]))
             model.add_reactions([r])
             r.lower_bound = -1000.0
@@ -164,16 +166,16 @@ def parse_file(filepath, fp_stack=None, rel_path=''):
         fp_stack.append(filepath)
 
     with open(os.path.join(rel_path, filepath)) as infile:
-        reactions, metabolites = parse_fobj(infile, fp_stack, rel_path, filepath)
-    return reactions, metabolites
+        reactions, metabolites, externals = parse_fobj(infile, fp_stack, rel_path, filepath)
+    return reactions, metabolites, externals
 
 
 def parse_string(spy_string, rel_path='.'):
     with StringIO() as fstr:
         fstr.write(spy_string)
         fstr.seek(0)
-        reactions, metabolites = parse_fobj(fstr, [], rel_path, "scrumpy_string")
-    return reactions, metabolites
[…]

### 2018-09-24 tests/test_scrumpy.py

@@ -1,6 +1,6 @@
 from gsmodutils.utils.io import load_model
-from gsmodutils.utils.scrumpy import load_scrumpy_model
-from tutils import scrumpy_model_path, scrumpy_biomass_path
+from gsmodutils.utils.scrumpy import load_scrumpy_model, ParseError
+from tutils import scrumpy_model_path, scrumpy_biomass_path, CleanUpDir
 import tempfile
 import cobra
 from click.testing import CliRunner
@@ -47,8 +47,29 @@ def test_cli_tool():
 
 def test_load_model():
     # Test loading of scrumpy models
-    model = load_scrumpy_model(scrumpy_model_path, media={"NH3_tx": -1000})
+    model = load_scrumpy_model(scrumpy_model_path, media={"NH3_tx": -1000}, fixed_fluxes={"NOTREAL": 1000})
     assert isinstance(model, cobra.Model)
 
-    with pytest.raises(KeyError):
-        load_scrumpy_model(scrumpy_model_path, fixed_fluxes={"NOTREAL": 1000})
+
+def test_cyclic_files():
+    """ Shouldn't allow cyclic file definitions """
+
+    spy_file_1 = """
+Include(File2.spy)
+    """
+    spy_file_2 = """
+Include(File1.spy)
+    """
+    # Write to temp dirs
+
+    with CleanUpDir() as dir:
+        filepath_1 = os.path.join(dir.path, "File1.spy")
+        filepath_2 = os.path.join(dir.path, "File1.spy")
+        with open(filepath_1, "w+") as file1:
+            file1.write(spy_file_1)
+
+        with open(filepath_2, "w+") as file2:
+            file2.write(spy_file_2)
+
+        with pytest.raises(ParseError):
+            load_scrumpy_model(filepath_1)

### 2018-09-24 gsmodutils/utils/scrumpy.py
More extensive scrumpy tests

@@ -103,7 +103,7 @@ def load_scrumpy_model(filepath_or_string, name=None, model_id=None, media=None,
     return model
 
 
-def get_tokens(line_dt):
+def get_tokens(line):
     """
     Goes through each charachter in scrumpy file attempting to find tokens
 
@@ -112,6 +112,8 @@ def get_tokens(line_dt):
     :param line_dt:
     :return:
     """
+
+    line_dt = line.strip().split('#')[0]
 
     tokens = []
     quoted = False
@@ -193,9 +195,7 @@ def parse_fobj(infile, fp_stack, rel_path, source_name):
 
     for linecount, line in enumerate(infile):
         # Ignore anything after comments
-        line_dt = line.strip().split('#')[0]
-
-        tokens = get_tokens(line_dt)
+        tokens = get_tokens(line)
         prev_token = ''
         # print tokens
         for token in tokens:

### 2018-09-24 tests/helpers/tutils.py

@@ -10,6 +10,7 @@ _CORE_MODEL_PATH = os.path.join(os.path.dirname(__file__), 'e_coli_core.json')
 
 scrumpy_model_path = os.path.join(os.path.dirname(__file__), 'scrumpy', 'MetaSal.spy')
 scrumpy_biomass_path = os.path.join(os.path.dirname(__file__), 'scrumpy', 'biomass_comp.json')
+scrumpy_media_path = os.path.join(os.path.dirname(__file__), 'scrumpy', 'media.json')
 
 
 class CleanUpFile(object):

### 2018-09-24 tests/test_scrumpy.py

@@ -1,6 +1,6 @@
 from gsmodutils.utils.io import load_model
-from gsmodutils.utils.scrumpy import load_scrumpy_model, ParseError
-from tutils import scrumpy_model_path, scrumpy_biomass_path, CleanUpDir
+from gsmodutils.utils.scrumpy import load_scrumpy_model, ParseError, get_tokens
+from tutils import scrumpy_model_path, scrumpy_biomass_path, scrumpy_media_path, CleanUpDir
 import tempfile
 import cobra
 from click.testing import CliRunner
@@ -44,6 +44,15 @@ def test_cli_tool():
     # Number of transporters + metabolites should equal scrumpy's number of metabolites
     assert len(model.exchanges) == 71  # 67
 
+    result = runner.invoke(gsmodutils.utils.scrumpy.scrumpy_to_cobra, [
+        scrumpy_model_path,
+        'MetaSal',
+        '--output', output_file_path,
+        '--media', scrumpy_media_path,
+        '--name', 'samonella scrumpy model',
+        '--objective', 'Glc_tx',
+    ])
+
 
 def test_load_model():
     # Test loading of scrumpy models
@@ -73,3 +82,23 @@ Include(File1.spy)
 
         with pytest.raises(ParseError):
             load_scrumpy_model(filepath_1)
+
+
+def test_get_tokens():
+    tokens = get_tokens("")
+    assert len(tokens) == 0
+
+    line = "A ->B"
+    tokens = get_tokens(line)
+    assert len(tokens) == 3
+    assert tuple(tokens) == ("A", "->", "B")
+
+    line = "\"FOO\"<>\"BOO\""
+    tokens = get_tokens(line)
+    assert len(tokens) == 3
+    assert tuple(tokens) == ("\"FOO\"", "<>", "\"BOO\"")
+
+    line = "\"FOO\"<- \"POO"
+    tokens = get_tokens(line)
+    assert len(tokens) == 3
+    assert tuple(tokens) == ("\"FOO\"", "<-", "\"POO")

### 2018-09-24 gsmodutils/utils/scrumpy.py
More extensive scrumpy tests

@@ -270,10 +270,7 @@ def parse_fobj(infile, fp_stack, rel_path, source_name):
                     externals += exset
                 continue
 
-            if token == '':
-                continue
-
-            elif token == 'External':
+            if token == 'External':
                 in_external = True
 
             elif token == 'Include':

### 2018-09-24 tests/test_model.py

@@ -21,15 +21,15 @@ def test_load_model():
         model.load_conditions("xyl_src")
 
         model.diff()
-        model.diff(GSModutilsModel(project, design="cbb_cycle"))
+
+        model.diff(model)
 
         with pytest.raises(TypeError):
             model.diff("should break")
 
         # Test loading non design fails
         with pytest.raises(TypeError):
-            model = project.model
-            model.set_design({})
+            model = GSModutilsModel(project, design={})
 
         model.save_model()
 

### 2018-09-24 tests/test_scrumpy.py

@@ -85,8 +85,26 @@ Include(File1.spy)
 
 
 def test_get_tokens():
+    # Should never return empty tokens
     tokens = get_tokens("")
     assert len(tokens) == 0
+
+    tokens = get_tokens(" ")
+    assert len(tokens) == 0
+
+    tokens = get_tokens("     ")
+    assert len(tokens) == 0
+
+    tokens = get_tokens("   A  ")
+    assert len(tokens) == 1
+
+    tokens = get_tokens("   A  BB")
+    assert tuple(tokens) == ("A", "BB")
+    assert len(tokens) == 2
+
+    tokens = get_tokens("\"\'A\'\"  BB")
+    assert tuple(tokens) == ("\"\'A\'\"", "BB")
+    assert len(tokens) == 2
 
     line = "A ->B"
     tokens = get_tokens(line)
@@ -102,3 +120,5 @@ def test_get_tokens():
     tokens = get_tokens(line)
     assert len(tokens) == 3
     assert tuple(tokens) == ("\"FOO\"", "<-", "\"POO")
+
+

### 2018-09-24 gsmodutils/utils/__init__.py
Test coverage

@@ -6,12 +6,12 @@ import sys
 # Hack for python version 2 to allow context management of StringIO
 if sys.version_info[0] == 2:
 
-    class StringIO(StringIO):
-        def __enter__(self):
-            return self
+    class StringIO(StringIO):  # pragma: no cover
+        def __enter__(self):  # pragma: no cover
+            return self   # pragma: no cover
 
-        def __exit__(self, *args):
-            self.close()
+        def __exit__(self, *args):  # pragma: no cover
+            self.close()  # pragma: no cover
 
 
 class FrozenDict(dict):

### 2018-09-24 gsmodutils/cli.py
Test coverage

@@ -30,7 +30,7 @@ def _load_project(project_path):
 @click.group()
 def cli():
     """Command line tools for management of gsmodutils genome scale model projects"""
-    pass
+    pass  # pragma: no cover
 
 
 def _output_child_logs(log, verbose=False, indent=4, baseindent=4):
@@ -487,12 +487,12 @@ def docker(project_path, overwrite, build, save, tag, save_path):
     client = docker.from_env()
     try:
         client.info()
-    except ConnectionError:
-        click.echo(
+    except ConnectionError:  # pragma: no cover
+        click.echo(  # pragma: no cover
             click.style('Error: Docker is either not installed or not configured on your system'
                         'please consult the documentation at docs.docker.com', fg='red')
         )
-        exit(-1)
+        exit(-1)  # pragma: no cover
 
     ttag = project.config.name
     if tag is not None:
@@ -509,49 +509,49 @@ def docker(project_path, overwrite, build, save, tag, save_path):
             client.images.build(path=project.project_path, tag=ttag)
             click.echo('Image built')
 
-        except BuildError as e:
-            click.echo(
+        except BuildError as e:  # pragma: no cover
+            click.echo(  # pragma: no cover
                 click.style('Error building project:\n{}'.format(e), fg='red')
             )
 
-        except APIError as e:
-            click.echo(
+        except APIError as e:  # pragma: no cover
+            click.echo( # pragma: no cover
                 click.style('Docker returned an error:\n{}'.format(e), fg='red')
             )
-            exit(-1)
+            exit(-1)  # pragma: no cover
 
     if save:
         # ensure that build has been completed first
-        image = None
-        try:
-            image = client.images.get(ttag)
-        except ImageNotFound:
-            click.echo(
+        image = None  # pragma: no cover
+        try:  # pragma: no cover
+            image = client.images.get(ttag)  # pragma: no cover
+        except ImageNotFound:  # pragma: no cover
+            click.echo(  # pragma: no cover
                 click.style('Image not found try building image first', fg='red')
             )
-            exit(-1)
-        except APIError as e:
-            click.echo(
+            exit(-1)  # pragma: no cover
+        except APIError as e:  # pragma: no cover
+            click.echo(  # pragma: no cover
                 click.style('Docker returned an error:\n{}'.format(e), fg='red')
+            )  # pragma: no cover
+            exit(-1)  # pragma: no cover
+
+        if save_path is None:  # pragma: no cover
[…]

### 2018-09-24 tests/test_utils.py

@@ -1,6 +1,6 @@
 from tutils import FakeProjectContext
 import pytest
-from gsmodutils.utils.io import load_model
+from gsmodutils.utils.io import load_model, load_medium
 import tempfile
 from gsmodutils.utils.validator import validate_model
 import cobra
@@ -14,8 +14,18 @@ def test_load_model():
 
     with pytest.raises(TypeError):
         # Create a fake file, format is not valid
-        with tempfile.TemporaryFile() as fp:
-            load_model(fp, file_format="foo")
+        with tempfile.NamedTemporaryFile() as fp:
+            load_model(fp.name, file_format="foo")
+
+    with FakeProjectContext() as ctx:
+        model = ctx.project.model
+
+        load_medium(model, {}, copy=True)
+        with pytest.raises(TypeError):
+            load_medium(model, set())
+
+        with pytest.raises(TypeError):
+            load_medium(set(), dict())
 
 
 def test_validator():

### 2018-09-24 tests/test_model.py
Test coverage

@@ -32,6 +32,8 @@ def test_load_model():
             model = GSModutilsModel(project, design={})
 
         model.save_model()
+        cpy = model.to_cobra_model()
+        assert not isinstance(cpy, GSModutilsModel)
 
 
 def test_copy():

### 2018-09-25 gsmodutils/project/model.py
Fixed some bugs in the model saving designs and added test coverage

@@ -56,6 +56,9 @@ class GSModutilsModel(cobra.Model):
     @property
     def model_path(self):
         """ Models path on disk """
+        if self.mpath is None:
+            return None
+
         return os.path.join(self.project.project_path, self.mpath)
 
     def _load_cobra_model(self):
@@ -63,7 +66,7 @@ class GSModutilsModel(cobra.Model):
         Loads the cobra model
         :return: cobra.Model instance of self
         """
-        if not os.path.exists(self.model_path):
+        if self.model_path is not None and not os.path.exists(self.model_path):
             raise IOError("Model file not found")
 
         if self.design is not None:
@@ -107,7 +110,8 @@ class GSModutilsModel(cobra.Model):
             if self.design.is_pydesign:
                 raise NotImplementedError("It is not possible to save python based designs in this manner")
 
-            self.save_as_design(self, self.design.id, self.design.name, self.design.description)
+            self.project.save_design(self, self.design.id, self.design.name, description=self.design.description,
+                                     base_model=self.mpath, parent=self.design.parent, overwrite=True)
         else:
             with self.project.project_context_lock:
                 model_type = self.mpath.split(".")[-1]

### 2018-09-25 tests/test_model.py

@@ -29,11 +29,27 @@ def test_load_model():
 
         # Test loading non design fails
         with pytest.raises(TypeError):
-            model = GSModutilsModel(project, design={})
+            GSModutilsModel(project, design={})
+
+        with pytest.raises(TypeError):
+            GSModutilsModel({})
+
+        with pytest.raises(IOError):
+            GSModutilsModel(project, mpath="/this/is/a/fake/path")
 
         model.save_model()
         cpy = model.to_cobra_model()
         assert not isinstance(cpy, GSModutilsModel)
+
+
+def test_save_design():
+    with FakeProjectContext() as ctx:
+        ctx.add_fake_conditions()
+        ctx.add_fake_designs()
+        project = GSMProject(ctx.path)
+        design = project.get_design("mevalonate_cbb")
+        mdl = GSModutilsModel(project, design=design)
+        mdl.save_model()
 
 
 def test_copy():

### 2018-09-25 tests/test_cli.py
Test coverage

@@ -37,6 +37,8 @@ def test_docker():
 
 def test_info():
     with FakeProjectContext() as ctx:
+        ctx.add_fake_designs()
+        ctx.add_fake_conditions()
         runner = CliRunner()
         result = runner.invoke(gsmodutils.cli.info, ['--project_path', ctx.path])
         assert result.exit_code == 0

### 2018-09-25 gsmodutils/project/model.py
Testing for using trying to save python based designs

@@ -108,6 +108,9 @@ class GSModutilsModel(cobra.Model):
         """
         if self.design is not None:
             if self.design.is_pydesign:
+                logger.error("{} is a python based design".format(self.design.id) +
+                             "Trying to save a python based design in this manner will not work." +
+                             "Consider saving differences as a child using GSModutilsModel.save_as_design")
                 raise NotImplementedError("It is not possible to save python based designs in this manner")
 
             self.project.save_design(self, self.design.id, self.design.name, description=self.design.description,

### 2018-09-25 tests/helpers/tutils.py

@@ -207,3 +207,15 @@ class FakeProjectContext(object):
 
         project.save_design(model, 'mevalonate_cbb', 'mevalonate production', parent='cbb_cycle',
                             description='Reactions for the production of mevalonate', overwrite=True)
+
+        py_design = """
+def gsmdesign_testpy(model, project):
+    reaction = model.reactions.get_by_id("ATPM")
+    reaction.bounds = (-999, 999)
+    return model
+
+gsmdesign_testpy.parent = "mevalonate_cbb"
+        """
+        ndpath = os.path.join(project.design_path, 'design_fake.py')
+        with open(ndpath, 'w+') as desf:
+            desf.write(py_design)

### 2018-09-25 tests/test_model.py

@@ -51,6 +51,12 @@ def test_save_design():
         mdl = GSModutilsModel(project, design=design)
         mdl.save_model()
 
+        # Can't save py designs that have been modified
+        with pytest.raises(NotImplementedError):
+            design = project.get_design("fake_testpy")
+            mdl = GSModutilsModel(project, design=design)
+            mdl.save_model()
+
 
 def test_copy():
     with FakeProjectContext() as ctx:

### 2018-09-25 tests/test_project.py
Test coverage

@@ -295,6 +295,9 @@ def test_design_parent():
             project.save_design(model, 'mevalonate_cbbd', 'mevalonate production', parent=des,
                                 description='Reactions for the production of mevalonate')
 
+        project.list_models()
+        project.list_condtions()
+
 
 def test_save_infeasible_design():
 

### 2018-09-25 tests/test_tests.py

@@ -169,6 +169,8 @@ def test_model(model, project, log):
 
         assert result.exit_code == -1
 
+        project.run_tests()
+
 
 def test_conditions():
     with FakeProjectContext() as fp:

### 2018-09-25 tests/test_project.py
Test coverage

@@ -295,8 +295,8 @@ def test_design_parent():
             project.save_design(model, 'mevalonate_cbbd', 'mevalonate production', parent=des,
                                 description='Reactions for the production of mevalonate')
 
-        project.list_models()
-        project.list_condtions()
+        _ = project.list_models
+        _ = project.list_conditions
 
 
 def test_save_infeasible_design():

### 2018-09-25 gsmodutils/utils/metacyc.py
Metacyc tests (90% coverage)

@@ -94,12 +94,13 @@ def parse_metacyc_file(fpath, unique_fields):
     return db
 
 
-def add_pathway(model, enzyme_ids=None, reaction_ids=None, db_path=None, copy=False):
+def add_pathway(model, enzyme_ids=None, reaction_ids=None, compartment="c", db_path=None, copy=False):
     """
     For a given model add enzymes from metacyc database
     :param model: cobra model object
     :param enzyme_ids: list of EC entires in format ["EC-x.x.x.x", ...]
     :param reaction_ids: set of reaction ids to add to model
+    :param compartment: compartment pathway is in
     :param db_path: path to the metacyc dat files on disk
     :param copy:
     :return:
@@ -124,7 +125,7 @@ def add_pathway(model, enzyme_ids=None, reaction_ids=None, db_path=None, copy=Fa
     added_metabolites = []
     for rid in set(reaction_ids):
         if rid not in model.reactions:
-            react, ametabolites = add_reaction(model, rid, db)
+            react, ametabolites = add_reaction(model, rid, db, compartment=compartment)
             added_reactions.append(react)
             added_metabolites += added_metabolites
 
@@ -165,12 +166,13 @@ def get_enzyme_reactions(eid, db):
     return erids
 
 
-def add_reaction(model, reaction_id, db):
+def add_reaction(model, reaction_id, db, compartment='c'):
     """
     Add a metacyc reaction id to a cobrapy model
     :param model: cobrapy model instance
     :param reaction_id: reaction identifier in metacyc
     :param db: dictionary db object
+    :param compartment: compartment reactions takeplace in (default is "c")
     :return: tuple(reaction, added_metabolites) cobrapy Reaction and Metabolite instances
     """
     metabolites = dict()
@@ -185,7 +187,7 @@ def add_reaction(model, reaction_id, db):
         mid = mid.replace('|', '')
         metabolites[mid] = -1 * coef
 
-    model.add_reaction(reaction)
+    model.add_reactions([reaction])
     added_metabolites = []
     for mid in metabolites:
         if mid not in model.metabolites:
@@ -198,6 +200,7 @@ def add_reaction(model, reaction_id, db):
             m = Metabolite(id=mid)
             m.name =cpd['COMMON-NAME']
             m.annotation = dict(metacyc_data=cpd)
+            m.compartment = compartment
             model.add_metabolites([m])
             added_metabolites.append(mid)
 
@@ -217,7 +220,7 @@ def add_reaction(model, reaction_id, db):
[…]

### 2018-09-25 tests/test_metacyc.py

@@ -1,0 +1,23 @@
+"""
+Tests for metacyc integration
+
+Requires the use of fake database as real metacyc databases require a license
+"""
+from tutils import METACYC_DB_PATH, FakeProjectContext
+from gsmodutils.utils import metacyc
+import os
+
+
+def test_db():
+
+    with FakeProjectContext() as ctx:
+        cache_location = os.path.join(ctx.path, '.gsmod_metacyc_db.json')
+        db = metacyc.parse_db(METACYC_DB_PATH)
+        db = metacyc.parse_db(METACYC_DB_PATH)
+        _ = metacyc.build_universal_model(METACYC_DB_PATH, use_cache=True, cache_location=cache_location)
+        _ = metacyc.build_universal_model(METACYC_DB_PATH, use_cache=True, cache_location=cache_location)
+        _ = metacyc.build_universal_model(METACYC_DB_PATH, use_cache=False)
+        model = ctx.model
+        metacyc.add_pathway(model, reaction_ids=["ALCOHOL-DEHYDROG-RXN"], db_path=METACYC_DB_PATH)
+        model = ctx.model
+        metacyc.add_pathway(model, enzyme_ids=["EC-1.1.1.1"], db_path=METACYC_DB_PATH)

### 2018-09-25 setup.py
LICENSE and python 3.7 support

@@ -46,6 +46,7 @@ setup(
         'Programming Language :: Python :: 3.4',
         'Programming Language :: Python :: 3.5',
         'Programming Language :: Python :: 3.6',
+        'Programming Language :: Python :: 3.7',
         'Programming Language :: Python :: Implementation :: CPython',
         'Topic :: Scientific/Engineering',
         'Topic :: Scientific/Engineering :: Bio-Informatics'

### 2018-09-25 setup.py
LICENSE and python 3.7 support

@@ -24,7 +24,7 @@ setup(
     install_requires=requirements,
     author_email="james.gilbert@nottingham.ac.uk",
     url="",
-    license="MIT",
+    license="Apache 2.0",
     packages=find_packages(),
     entry_points={
         'console_scripts': [

### 2018-09-27 article/article.tex
Article updates

@@
 
-\title{Gsmodutils: A python based framework for test driven genome scale metabolic model development}
+\title{Gsmodutils: A python based framework for test-driven genome scale metabolic model development}
 
@@
 \begin{abstract}
-\textbf{Motivation:} Genome scale metabolic models (GSMMs) are increasingly important for systems biology and metabolic engineering research as they are capable of simulating complex steady state phenotypic behaviour.
+\textbf{Motivation:} Genome scale metabolic models (GSMMs) are increasingly important for systems biology and metabolic engineering research as they are capable of simulating complex steady-state behaviour.
 Constraints based models of this form can include thousands of reactions and metabolites, with many crucial pathways that only become activated in specific simulation settings.
-However, despite their widespread use, power and the availability of tools to aid with the construction and analysis of large scale models little methodology is suggested for the management of continually curated large scale models.
+However, despite their widespread use, power and the availability of tools to aid with the construction and analysis of large scale models, little methodology is suggested for the continued management of curated large scale models.
+For example, when genome annotations are updated or new understanding regarding behaviour of is discovered models may need to be updated.
 This is quickly becoming an issue for industrial systems and synthetic biotechnology applications, which require good quality reusable models integral to the design, build and test cycle.
@@
 \textbf{Results:} 
-As part of an ongoing effort to improve genome scale metabolic analysis, we have developed a test driven development methodology for the continuous integration of validation data from different sources.
-Contributing to the open source technology based around COBRApy, we have developed the \textit{gsmodutils} modelling framework placing an emphasis on test driven design of models through defined test cases.
-Crucially, different substrate conditions are configurable allowing users to examine how different designs or curation impact a wide range of system behaviours, minimising error between model versions.\\
+As part of an ongoing effort to improve genome scale metabolic analysis, we have developed a test-driven development methodology for the continuous integration of validation data from different sources.
+Contributing to the open source technology based around COBRApy, we have developed the \textit{gsmodutils} modelling framework placing an emphasis on test-driven design of models through defined test cases.
+Crucially, different substrates are configurable allowing users to examine how different designs or curation impact a wide range of system behaviours, minimising error between model versions.\\
 \textbf{Availability:} The software framework described within this paper is open source and freely available from \url{http://github.com/SBRCNottingham/gsmodutils} \\
@@
 Stoichiometric constraints based modelling for biological systems has been a mainstay of systems biology for several decades \cite{fell1986fat, varma1994stoichiometric}.
-Given its flexibility, low barrier to entry and requirement only on minimal knowledge regarding the stoichiometry of metabolic networks this structural approach has become an extremely popular method for modelling steady state behaviour of large, biochemical networks \cite{kauffman2003advances}.
+Given its flexibility, low barrier to entry and requirement only on minimal knowledge regarding the stoichiometry of metabolic networks this structural approach has become an extremely popular method for modelling steady-state behaviour of large, biochemical networks \cite{kauffman2003advances}.
 Such large scale reconstructions are often referred to as genome scale metabolic models (GSMMs), as the processes is significantly aided through the advent of relatively inexpensive genome sequencing \cite{o2015using, land2015insights}.
@@
 The BiGG models database \cite{king2015bigg}, for example, exists to provide a standardised repository of validated models that can be shared and reused.
-However, little focus is placed upon the collaborative design aspect of such models with few mechanisms existing to store the \textit{model delta} based tools provided in our software.
+However, little focus is placed upon the collaborative design aspect of such models with few mechanisms existing to capture the differences between two model versions, \textit{model deltas}.
 Similarly, the MetaNetX \cite{moretti2016metanetx} system exists to provide a standardised namespace and toolchain for GSM analysis.
@@
 Furthermore, as with many areas of bioinformatic study the number available computational tools has become vast.
-This covers huge variety of software platforms including the COBRA toolbox in Matlab \cite{schellenberger2011quantitative}, ScrumPy and COBRApy in Python \cite{poolman2006scrumpy, ebrahim2013cobrapy} with additional tools and libraries such as cameo \cite{cardoso2017cameo}, OptFlux in Java \cite{rocha2010optflux} and SurreyFBA \cite{gevorgyan2010surreyfba}.
-Whilst most of these tools are Open Source and follow standards, such as SBML \cite{finney2003systems}, there are often few options that make analysis easy to replicate.
-The archetype design, build and test cycle of synthetic biology heavily relies on the usage of bioinformatics software and modelling to improve the production of natural products \cite{carbonell2016bioinformatics}.
+This covers huge variety of software platforms including the COBRA toolbox for Matlab \cite{schellenberger2011quantitative}, ScrumPy and COBRApy in python \cite{poolman2006scrumpy, ebrahim2013cobrapy} with additional tools and libraries such as cameo \cite{cardoso2017cameo}, OptFlux in Java \cite{rocha2010optflux} and SurreyFBA \cite{gevorgyan2010surreyfba}.
+Whilst most of these tools are Open Source and follow standards, such as SBML \cite{finney2003systems}, it is often challenging to replicate the initial modelling efforts conducted by authors of papers.
+Consequently, we feel that software tools are urgently needed to address this issue.
+
+Similarly, The archetype design, build and test cycle of synthetic biology heavily relies on the usage of bioinformatics software and modelling to improve the production of natural products \cite{carbonell2016bioinformatics}.
 In order to speed up the use of bioinformatics tools to produce high value platform chemicals, genome scale models are often used to discover methods for process optimisation \textit{in silico}.
 
-For example, many tools such as RetroPath \cite{carbonell2013retropath}, XTMS \cite{carbonell2014xtms} and GEM-Path \cite{campodonico2014generation} suggest thousands of potential heterologous pathways as well as providing predictions for flux distributions.
+For example, many tools such as RetroPath \cite{carbonell2013retropath}, XTMS \cite{carbonell2014xtms} and GEM-Path \cite{campodonico2014generation} suggest thousands of potential heterologous pathways.
 Many of these tools significantly increase the value of genome scale models, for example by coupling commodity production to an organism's growth \cite{feist2010model}.
 These tools all suggest major changes to wild-type strains must be tracked and compared to allow models to remain relevant.
-In effect, mechanisms are required to relate modified models to modified test and production strains.
+In effect, mechanisms are required to relate to modified test and production strains.
 
-Similarly, many conventional applications of genome scale models in systems biology have often suffered from unnecessary replication of work due to a lack of standards.
-For example, there are now many independently developed models of \textit{Clostridium acetobutylicum} \cite{lee2008genome, senger2008genome, mcanulty2012genome, dash2014capturing, yoo2015quantitative}, an organism used in the production of solvents for over a century.
+Similarly, many conventional applications of genome scale models in systems biology have often suffered from unnecessary replication of work due to a lack of adherence to standards \cite{monk2014}.
+For example, there are now many independently developed models of \textit{Clostridium acetobutylicum} \cite{lee2008genome, senger2008genome, mcanulty2012genome, dash2014capturing, yoo2015quantitative}, an organism used in the production of solvents for around a century \cite{moon2016one, 1919charles}.
 These models all exist to solve similar biological problems some being updates to the inital base models. 
-However, there has been disagreement over fundamental biochemical properties of these anaerobic organisms, notably with the focus on redox balancing \cite{dash2014capturing}.
-Indeed, such models also include updates based on improved genome annotations and the inclusion of fluxomic, transcriptomic and metabolomic characterisations \cite{yoo2015quantitative}.
+However, there has been disagreement over fundamental biochemical properties of this anaerobic organism, notably with the focus on redox balancing \cite{dash2014capturing}.
+Such models also include updates based on improved genome annotations and the inclusion of fluxomic, transcriptomic and metabolomic characterisations \cite{yoo2015quantitative}.
 
-Many of the results presented in original sources are difficult to compare due to a lack of standardisation of reference identifiers for reaction names.
-Where models are shared they are often in non-standard spreadsheet formats, rather than SBML models.
-Indeed, even in the case of valid SBML models being made available at the time of publication few details are given as to how to run such models.
+%Many of the results presented in original sources are difficult to compare due to a lack of standardisation of reference identifiers for reaction names.
 
-% Perhaps more worryingly, from an industrial perspective, is that the use of models in bioproduction has not escaped regulatory notice \cite{}.
-Furthermore, as models are often used to explain issues with online production and facilitate the optimisation of bioprocessing, it stands to reason that model correctness should attract oversight.
-Consequently, methods are needed to ensure that the models themselves are tightly version controlled.
-In this paper, we argue for the use of a form of \textit{test driven} model of development, a conceptual approach borrowed from software engineering.
-Given the availability of widespread open source virtualisation tools that allow the standardisation of platforms \cite{merkel2014docker}, modellers have the opportunity to share highly reproducible research \cite{chamberlain2014using}.
+Unfortunatley, many of the results reported in \cite{lee2008genome, senger2008genome, mcanulty2012genome, dash2014capturing, yoo2015quantitative} are difficult to compare or reproduce as the result of a number of issues.
+Often, the models lack a standardisation of identifiers for reaction names, which makes direct comparisson of model structure as well as differences between reactions a challenge.
+Where models are shared, it is often in non-standard spreadsheet formats, rather than SBML models.
+Indeed, even in the case of valid SBML models being made available at the time of publication few details are given as to how to run such models for conditions discussed in original articles.
 
-In this paper we present a software framework geared toward \textit{test driven} genome scale model development, a concept that is taken directly from good software development practices.
+In this paper, we present a software framework geared toward \textit{test-driven} genome scale model development, a concept that is taken directly from good software development practices \cite{martin2002agile}.
 By this we mean the notion that, as a model is curated to represent biological phenomena, much of the validation can be turned into specific test cases that can be repeated between model versions.
 We provide an example test case for \textit{Clostridium autoethanogenum}, an organism that has had considerable focus in terms of genome scale models and how a working methodology using the software presented here can reduce repetition of work and improve the reproducibility of results.
-This article aims to summarise the main objectives of the gsmodutils software and we refer the reader to the software user guide for a more detailed exploration features.
+This article aims to summarise the main objectives of the \textit{gsmodutils} software and we refer the reader to the software user guide for a more detailed exploration features.
 
@@
 Such an approach best fits genome scale models because they are rarely created to investigate individual processes and, instead, capture the complexity of large systems.
-Constraints based models are intrinsically related to available genome annotations.
+Genome scale models are intrinsically related to available genome annotations.
 Such annotations rely heavily on automated matches to related species, with the characterisations of individual genes often being left to a few of critical interest \cite{seemann2014prokka}.
 This modelling formation costs an in depth understanding of dynamic behaviour. 
-However, capturing steady state phenomena still provides a good understanding of system properties \cite{o2015using}.
+However, capturing steady-state phenomena still provides a good understanding of system properties \cite{o2015using}.
 
@@
 
-Therefore, models undergo significant manual annotation - a process which naturally leads to errors.
-In this work we advocate a test driven approach to model development highlighted in Figure \ref{fig:test_driven}.
+Therefore, models undergo significant manual annotation and curation; a process which has a high chance of error.
+In this work we advocate a test-driven approach to model development highlighted in Figure \ref{fig:test_driven}.
 Here, the model is changed to achieve core research goals that are dynamic in response to the changes of a project.
@@
 However, there is a major gap in terms of the reliability and re-usability of models due to a lack of standardisation and software tools to aid such processes.
-The following sections provide an overview of the gsmodutils software framework.
-Gsmodutils aims to provide a basis for test driven, version controlled agile model development.
+The following sections provide an overview of the \textit{gsmodutils} software framework.
+\textit{gsmodutils} aims to provide a basis for test-driven, version controlled agile model development.
 All software is open source and is designed to be interoperable with platforms widely used in the domain of constraints based modelling.
@@
 \subsection{Outline and features}
-Test driven development is driven by the idea of clearly defined test cases written before significant changes are made to any underlying architecture.
+Test-driven development is driven by the idea of clearly defined test cases written before significant changes are made to any underlying architecture.
 In the case of genome scale models, errors occur as a product of human curation to better represent newly discovered aspects of metabolism.
@@
 A standard test case, ensuring that a given model grows on media is given in Figure \ref{fig:test_case}.
-When a new model repository is created with the gsmodutils tool, a number of pre-written test cases are automatically added to a file.
+When a new model repository is created with the \textit{gsmodutils} tool, a number of pre-written test cases are automatically added to a file.
 However, we stress that the vast majority of individual use cases for a model must be specific to a given biological problem.
@@
 
-\lstset{language=Python}   
+\lstset{language=python}   
 \begin{lstlisting}
@@
 \end{lstlisting}
+\caption{An example \textit{gsmodutils} test case written in python}
 \label{fig:test_case} 
-\caption{An example gsmodutils test case written in python}
 \end{figure}
@@
 \subsection{Strain designs}
-A core aspect behind the implementation of gsmodutils is the concept of a \textit{design}, this encompasses a simple set of changes to a ``wild-type'' model that are required for analysis.
+A core aspect behind the implementation of \textit{gsmodutils} is the concept of a \textit{design}, this encompasses a simple set of changes to a ``wild-type'' model that are required for analysis.
 For example a mutant strain that has undergone several gene deletions is not sufficiently different from the original wild-type to merit having its own genome scale model.
@@
 Consequently, such designs are hereditary in nature.
-By taking the \textit{model delta} between the constraints applied to an initial model and subsequent modifications, gsmodutils allows users to easily reuse and export models with this design.
+By taking the \textit{model delta} between the constraints applied to an initial model and subsequent modifications, \textit{gsmodutils} allows users to easily reuse and export models with this design.
 As designs inherit from a base model, future curation to a wild type model will automatically be included in the designs.
@@
 An example of this is shown in Figure \ref{fig:programmatic_design}.
-This example converts an existing model to one based on an MILP, as the reaction names do not need to be specified, should reactions be added or removed the process will still apply.
+This example converts an existing model to one based on a Mixed Integer Linear Program (MILP), as the reaction names do not need to be specified, should reactions be added or removed the process will still apply.
 Alternative examples could include reductions of models through methods such as elementary flux modes or minimum cut sets, which can change dramatically with only small changes to stoichiometry.
-Furthermore, integrity of all strain designs is automatically tested as part of the default gsmodutils testing framework.
+Furthermore, integrity of all strain designs is automatically tested as part of the default \textit{gsmodutils} testing framework.
 
 \begin{figure}[ht]
-\lstset{language=Python}   
+\lstset{language=python}   
 \begin{lstlisting}
@@
     return model
-
 \end{lstlisting}
+\caption{An example \textit{gsmodutils} programmatic design written in python. 
+This design converts reactions to integer type, allowing an MILP formation.}
 \label{fig:programmatic_design} 
-\caption{An example gsmodutils programmatic design written in python. 
-This design converts reactions to integer type, allowing an MILP formation.}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{inheritence2.eps}
-  \label{fig:strain_hered}
-  \caption{Examples of gsmodutils design inheritance.
+  \caption{Examples of \textit{gsmodutils} design inheritance.
   Each design stores the delta between the wild-type base model, any parents and the changes to constraints the design contains.
@@
   }
+    \label{fig:strain_hered}
 \end{figure}
@@
 \subsection{Development workflow}
-In this section we propose a method for the development of genome scale models that integrates gsmodutils with version control systems.
-Figure \ref{fig:test_case} highlights the notion of test cases, taken from test driven development.
+In this section we propose a method for the development of genome scale models that integrates \textit{gsmodutils} with version control systems.
+Figure \ref{fig:test_case} highlights the notion of test cases, taken from test-driven development.
 The outlined workflow is that the user writes a formal test case for some modelling goal, perhaps driven by captured experimental data, that fits a specific form of validation criteria.
@@
 autoethanogenum}, `MetaCLAU', has been analysed to improve this bioprocess (Norman et al., submitted for peer review).
-In this section we describe how gsmodutils has been utilised to ensure that future versions of MetaCLAU will remain functionally relevant from the perspective of industrial biotechnology.
+In this section we describe how \textit{gsmodutils} has been utilised to ensure that future versions of MetaCLAU will remain functionally relevant from the perspective of industrial biotechnology.
 
@@
 The solution to this FBA problem represents minimal protein investment \cite{holzhutter2006thegeneralised}.
-Since execution of flux minimisation in COBRApy requires a model in which reversible reactions are split into two irreversible reactions, a gsmodutils design was created in which all reversible reactions are split in a programmatic manner.
+Since execution of flux minimisation in COBRApy requires a model in which reversible reactions are split into two irreversible reactions, a \textit{gsmodutils} design was created in which all reversible reactions are split in a programmatic manner.
 In the case of MetaCLAU, the flux minimal solution includes both ethanol and acetate production, which represents good qualitative agreement with experimental data (Norman et al., unpublished).
@@
 As in the previous case, the model-predicted behaviour showing both lactate and 2,3BD was deemed an important result, which model curators should be notified of if lost during model development.
-Thus the analysis was built into a gsmodutils test.
+Thus the analysis was built into a \textit{gsmodutils} test.
 
@@
 % $\Delta \textit{nfn}$ knockout mutant: NOT SURE ABOUT THIS. CHRIS NOW THINKS IT'S ESSENTIAL :(
-These test cases are all fully reproducible and available as an example within the gsmodutils software repository and we encourage the reader to explore these examples to understand how to utilise gsmodutils within their own project.
+These test cases are all fully reproducible and available as an example within the \textit{gsmodutils} software repository and we encourage the reader to explore these examples to understand how to utilise \textit{gsmodutils} within their own project.
 
@@
 It features a fully specified set of tests, including custom test cases and has strong version control integration with git.
-A core difference between these projects is that gsmodutils has a stronger focus on reducing the redundancy in model storage through the use of \textit{design deltas}, as described above.
-Similarly, a core goal of gsmodutils is to allow easy import and export outside of the framework for compatibility with other modelling suites.
+A core difference between these projects is that \textit{gsmodutils} has a stronger focus on reducing the redundancy in model storage through the use of \textit{design deltas}, as described above.
+Similarly, a core goal of \textit{gsmodutils} is to allow easy import and export outside of the framework for compatibility with other modelling suites.
 It should be noted that, as memote is written in python and utilises COBRApy, it is fully compatible with gsmodutils.
@@
 gsmodutils integrates easily with cameo.
-Indeed, making a set of design changes can easily be integrated into gsmodutils through the python API.
+Indeed, making a set of design changes can easily be integrated into \textit{gsmodutils} through the python API.
 \\
@@
 
-As with most software development projects, gsmodutils will see expanded features.
+As with most software development projects, \textit{gsmodutils} will see expanded features.
 Initially this will include tighter integration with version control systems such as git and mercurial.

### 2018-09-27 article/article.tex
More minor article changes

@@
 \author{James P Gilbert}
-\email{james.gilbert@nottingham.ac.uk}
 \author{Nicole Pearcy}
@@
 \author{Jamie Twycross}
+\email{jamie.twycross@nottingham.ac.uk}
 \affiliation[IMA]{Intelligent Modelling and Analysis group, School of Computer Science, University of Nottingham, Nottingham, NG8 1BB, United Kingdom}
@@
 Genome scale models are intrinsically related to available genome annotations.
-Such annotations rely heavily on automated matches to related species, with the characterisations of individual genes often being left to a few of critical interest \cite{seemann2014prokka}.
+Such annotations rely heavily on automated matches to related species, with the characterisations of individual genes or changes in cofactors and substrates for specific reactions often being left to a few of critical interest \cite{seemann2014prokka}.
 This modelling formation costs an in depth understanding of dynamic behaviour. 
@@
 
-As such approaches often leave models with missing reactions, incorrect gene-reaction rules \cite{thiele2010protocol} or with pathways based on gap filling methods that add reactions that may not be present within the organism \cite{benedict2014likelihood}.
+As such, approaches often leave models with missing reactions, incorrect gene-reaction rules \cite{thiele2010protocol} or with pathways based on gap filling methods that add reactions that may not actually be catalysed by the organism in question \cite{benedict2014likelihood}.
 When attempting to understand specific natural phenomena, genome annotations are frequently updated and models are often corrected in an \textit{ad hoc} manner.
@@
 Therefore, models undergo significant manual annotation and curation; a process which has a high chance of error.
-In this work we advocate a test-driven approach to model development highlighted in Figure \ref{fig:test_driven}.
-Here, the model is changed to achieve core research goals that are dynamic in response to the changes of a project.
+In this work, we advocate a test-driven approach to model development highlighted in Figure \ref{fig:test_driven}.
+Here, the model is changed to achieve research goals that are dynamic in response to the changes of a project.
 In order to meet this objective, validation criteria for a model such as growth conditions or the impact of gene knock outs, should be formally set.
@@
 \textit{gsmodutils} aims to provide a basis for test-driven, version controlled agile model development.
-All software is open source and is designed to be interoperable with platforms widely used in the domain of constraints based modelling.
+All software and packages are open source and is designed to be interoperable with platforms widely used in the domain of constraints based modelling.
 
@@
 Test-driven development is driven by the idea of clearly defined test cases written before significant changes are made to any underlying architecture.
-In the case of genome scale models, errors occur as a product of human curation to better represent newly discovered aspects of metabolism.
+In the case of genome scale models, errors can easily occur as a product of human curation designed to better represent newly discovered aspects of metabolism.
 
@@
 \textit{C.
-autoethanogenum} is a strictly anaerobic, acetogenic bacterium which naturally produces ethanol and trace amounts of 2,3-butanediol (2,3BD) from carbon monoxide and water \cite{abrini1994clostridium,schuchmann2014autotrophy,norman2018progress}.
-Since carbon monoxide is readily available in the form of industrial waste gas, and 2,3BD has a global market value of \$43 billion \cite{kopke201123butanediol}, the optimisation of yields of 2,3BD from carbon monoxide is highly desirable in the context of industry \cite{norman2018progress}.
+autoethanogenum} is a strictly anaerobic, acetogenic bacterium which naturally produces ethanol and trace amounts of \textit{2,3-butanediol} (\textit{2,3 BD}) from carbon monoxide and water \cite{abrini1994clostridium,schuchmann2014autotrophy,norman2018progress}.
+Since carbon monoxide is readily available in the form of industrial waste gas, and \textit{2,3 BD} has a global market value of \$43 billion \cite{kopke201123butanediol}, the optimisation of yields of \textit{2,3 BD} from carbon monoxide is highly desirable in the context of industry \cite{norman2018progress}.
 MetaCLAU was built using Pathway Tools \cite{karp2002pathway} and ScrumPy \cite{poolman2006scrumpy}, and is based on a manually annotated genome sequence of \textit{C.
@@
 autoethanogenum} under conditions where the organism can and cannot produce molecular hydrogen (with carbon monoxide as sole carbon and energy source).
-The hypothesis tested in (Norman et al., unpublished), was that in the case where hydrogen production in infeasible, alternative electron sinks like lactate and 2,3BD would be produced.
-As in the previous case, the model-predicted behaviour showing both lactate and 2,3BD was deemed an important result, which model curators should be notified of if lost during model development.
+The hypothesis tested in (Norman et al., unpublished), was that in the case where hydrogen production infeasible, alternative electron sinks like lactate and \textit{2,3 BD} would be produced.
+As in the previous case, the model-predicted behaviour showing both lactate and \textit{2,3 BD} was deemed an important result, which model curators should be notified of if lost during model development.
 Thus the analysis was built into a \textit{gsmodutils} test.

### 2018-09-27 article/article.tex
missing figure

@@
 We would like to thank the Oxford Brookes Cell Systems Modelling group for helpful discussions regarding this work and for scientific advice.
-This work was supported by the UK Biotechnology and Biological Sciences Research Council (BBSRC) grants BB/L013940/1,  BB/K00283X/1 and BB/L502030/1.
+This work was supported by the UK Biotechnology and Biological Sciences Research Council (BBSRC) and Engineering and Physical Sciences Research Council (EPSRC) grants BB/L013940/1,  BB/K00283X/1 and BB/L502030/1.
 \end{acknowledgement}

### 2018-09-27 article/article.tex
article changes

@@
 However, despite their widespread use, power and the availability of tools to aid with the construction and analysis of large scale models, little methodology is suggested for the continued management of curated large scale models.
-For example, when genome annotations are updated or new understanding regarding behaviour of is discovered models may need to be updated.
-This is quickly becoming an issue for industrial systems and synthetic biotechnology applications, which require good quality reusable models integral to the design, build and test cycle.
+For example, when genome annotations are updated or new understanding regarding behaviour of is discovered, models often need to be altered to reflect this.
+This is quickly becoming an issue for industrial systems and synthetic biotechnology applications, which require good quality reusable models integral to the design, build and test cycle in biprodcessing and related fields.
 \\
@@
 Contributing to the open source technology based around COBRApy, we have developed the \textit{gsmodutils} modelling framework placing an emphasis on test-driven design of models through defined test cases.
-Crucially, different substrates are configurable allowing users to examine how different designs or curation impact a wide range of system behaviours, minimising error between model versions.\\
+Crucially, different conditions are configurable allowing users to examine how different designs or curation impact a wide range of system behaviours, minimising error between model versions.\\
 \textbf{Availability:} The software framework described within this paper is open source and freely available from \url{http://github.com/SBRCNottingham/gsmodutils} \\
@@
 \includegraphics[width=0.8\textwidth]{testdriven.eps}
-\caption{Workflow for model development proposed for gsmodutils.
+\caption{Workflow for model development proposed for \textit{gsmodutils}.
 The objective of this approach is to simultaneously capture research questions, model validation criteria and minimise the impact of changes on previously completed models.
@@
 submitted for peer review).
-The model has been integrated with the gsodutils modelling framework as a test-driven project.
+The model has been integrated with the \textit{gsmodutils} modelling framework as a test-driven project.
 The following section details specific tests used to evaluate the model at each stage of its continued development.
@@
 \section{Related work}
-\textbf{Memote:} \cite{lieven2017memote} is tool set that has very similar ambitions to gsmodutils.
+\textbf{Memote:} \cite{lieven2017memote} is tool set that has very similar ambitions to \textit{gsmodutils}.
 It features a fully specified set of tests, including custom test cases and has strong version control integration with git.
@@
 Similarly, a core goal of \textit{gsmodutils} is to allow easy import and export outside of the framework for compatibility with other modelling suites.
-It should be noted that, as memote is written in python and utilises COBRApy, it is fully compatible with gsmodutils.
+It should be noted that, as memote is written in python, utilises COBRApy and, at the time of writing, it is fully compatible with \textit{gsmodutils}.
 \\

### 2018-09-27 article/article.tex
typo

@@
 \textbf{Model repositories:} models are frequently shared, at the time of publication through services such as BiGG \cite{king2015bigg} and Biomodels \cite{chelliah2013biomodels}. 
-Whilst these repositories encourage the reuse of models and the reproducibility of \textit{in silico} predictions they are not designed to improved collaboration.
+Whilst these repositories encourage the reuse of models and the reproducibility of \textit{in silico} predictions they are not designed to improve collaboration.
 The software presented here is designed with the notion that genome scale models are never finished, \textit{per se}, but under continuous development.

### 2018-09-28 article/article.tex
Article ready for bioArxiv submission

@@
-\documentclass[journal=asbcd6,10pt]{achemso}
+\documentclass[journal=asbcd6]{achemso}
+\setkeys{acs}{articletitle=true}
 \usepackage[utf8]{inputenc}
@@
 \author{James P Gilbert}
+\affiliation[sbrc]{Synthetic Biology Research Centre, University of Nottingham, Nottingham, NG7 2RD, United Kingdom}
 \author{Nicole Pearcy}
+\affiliation[sbrc]{Synthetic Biology Research Centre, University of Nottingham, Nottingham, NG7 2RD, United Kingdom}
 \author{Rupert Norman}
 \affiliation[sbrc]{Synthetic Biology Research Centre, University of Nottingham, Nottingham, NG7 2RD, United Kingdom}
+\alsoaffiliation[Biosciences]{School of Biosciences, University of Nottingham, Sutton Bonington, Loughborough, LE12 5RD, United Kingdom}
 \author{Thomas Millat}
@@
 \begin{abstract}
-\textbf{Motivation:} Genome scale metabolic models (GSMMs) are increasingly important for systems biology and metabolic engineering research as they are capable of simulating complex steady-state behaviour.
+\textbf{Motivation:}
+Genome scale metabolic models (GSMMs) are increasingly important for systems biology and metabolic engineering research as they are capable of simulating complex steady-state behaviour.
 Constraints based models of this form can include thousands of reactions and metabolites, with many crucial pathways that only become activated in specific simulation settings.
@@
 For example, when genome annotations are updated or new understanding regarding behaviour of is discovered, models often need to be altered to reflect this.
-This is quickly becoming an issue for industrial systems and synthetic biotechnology applications, which require good quality reusable models integral to the design, build and test cycle in biprodcessing and related fields.
+This is quickly becoming an issue for industrial systems and synthetic biotechnology applications, which require good quality reusable models integral to the design, build and test cycle.
 \\
@@
 For example, there are now many independently developed models of \textit{Clostridium acetobutylicum} \cite{lee2008genome, senger2008genome, mcanulty2012genome, dash2014capturing, yoo2015quantitative}, an organism used in the production of solvents for around a century \cite{moon2016one, 1919charles}.
-These models all exist to solve similar biological problems some being updates to the inital base models. 
+These models all exist to solve similar biological problems some being updates to the initial base models. 
 However, there has been disagreement over fundamental biochemical properties of this anaerobic organism, notably with the focus on redox balancing \cite{dash2014capturing}.
@@
 
-Unfortunatley, many of the results reported in \cite{lee2008genome, senger2008genome, mcanulty2012genome, dash2014capturing, yoo2015quantitative} are difficult to compare or reproduce as the result of a number of issues.
-Often, the models lack a standardisation of identifiers for reaction names, which makes direct comparisson of model structure as well as differences between reactions a challenge.
+Unfortunately, many of the results reported in \cite{lee2008genome, senger2008genome, mcanulty2012genome, dash2014capturing, yoo2015quantitative} are difficult to compare or reproduce as the result of a number of issues.
+Often, the models lack a standardisation of identifiers for reaction names, which makes direct comparison of model structure as well as differences between reactions a challenge.
 Where models are shared, it is often in non-standard spreadsheet formats, rather than SBML models.
@@
 An important limitation of FBA is that optimal solutions may be thermodynamically infeasible if appropriate constraints are not applied \cite{Fell1197rohtua}.
-In order to identify these unwanted flux distributions and constrain the model such that they are not in the feasible soluton space, a diagnostic FBA is applied with the following constraints:
+In order to identify these unwanted flux distributions and constrain the model such that they are not in the feasible solution space, a diagnostic FBA is applied with the following constraints:
 1) All transport reactions are constrained to allow no uptake (lower bound = 0), and 2) the ATPase reaction is given a fixed flux of one.
@@
 In the case of MetaCLAU, the flux minimal solution includes both ethanol and acetate production, which represents good qualitative agreement with experimental data (Norman et al., unpublished).
-Since any changes to this predicted phenotype must be investigated, the flux minimisation analysis has been formulated as a gsmuodutils test.
+Since any changes to this predicted phenotype must be investigated, the flux minimisation analysis has been formulated as a \textit{gsmodutils} test.
 
@@
 
-% $\Delta \textit{nfn}$ knockout mutant: NOT SURE ABOUT THIS. CHRIS NOW THINKS IT'S ESSENTIAL :(
 These test cases are all fully reproducible and available as an example within the \textit{gsmodutils} software repository and we encourage the reader to explore these examples to understand how to utilise \textit{gsmodutils} within their own project.
@@
 \textbf{Cameo:} \cite{cardoso2017cameo} is a set of strain design utilities, such as genetic algorithms for finding optimal gene knock-out
-strategies, as well as other flux balance analysis utilities. Written in python and part of the COBRApy suite of applications,
-gsmodutils integrates easily with cameo.
+strategies, as well as other flux balance analysis utilities. Written in python and part of the COBRApy suite of applications, \textit{gsmodutils} works well with cameo.
 Indeed, making a set of design changes can easily be integrated into \textit{gsmodutils} through the python API.
@@
 
-\bibliographystyle{ieeetrans}
 \bibliography{references}

### 2018-09-28 article/article.tex
Article ready for bioArxiv submission
Minor whitespace change
Submission copy

@@
 \begin{abstract}
-\textbf{Motivation:}
-Genome scale metabolic models (GSMMs) are increasingly important for systems biology and metabolic engineering research as they are capable of simulating complex steady-state behaviour.
+\textbf{Motivation:} Genome scale metabolic models (GSMMs) are increasingly important for systems biology and metabolic engineering research as they are capable of simulating complex steady-state behaviour.
 Constraints based models of this form can include thousands of reactions and metabolites, with many crucial pathways that only become activated in specific simulation settings.
@@
 \textbf{Availability:} The software framework described within this paper is open source and freely available from \url{http://github.com/SBRCNottingham/gsmodutils} \\
-\textbf{Contact:}{james.gilbert@nottingham.ac.uk}
 \end{abstract}
@@
 A new GSMM of \textit{C.
-autoethanogenum}, `MetaCLAU', has been analysed to improve this bioprocess (Norman et al., submitted for peer review).
+autoethanogenum}, `MetaCLAU', has been analysed to improve this bioprocess (\cite{norman2018progress}, submitted for peer review).
 In this section we describe how \textit{gsmodutils} has been utilised to ensure that future versions of MetaCLAU will remain functionally relevant from the perspective of industrial biotechnology.
@@
 The resulting model consisted of 758 reactions, 773 metabolites and 518 genes.
-For full details of the model, see (Norman et al.
+For full details of the model, see (\cite{norman2018progress}.
 submitted for peer review).
@@
 Since execution of flux minimisation in COBRApy requires a model in which reversible reactions are split into two irreversible reactions, a \textit{gsmodutils} design was created in which all reversible reactions are split in a programmatic manner.
-In the case of MetaCLAU, the flux minimal solution includes both ethanol and acetate production, which represents good qualitative agreement with experimental data (Norman et al., unpublished).
+In the case of MetaCLAU, the flux minimal solution includes both ethanol and acetate production, which represents good qualitative agreement with experimental data (\cite{norman2018progress}, under review).
 Since any changes to this predicted phenotype must be investigated, the flux minimisation analysis has been formulated as a \textit{gsmodutils} test.
@@
 autoethanogenum} under conditions where the organism can and cannot produce molecular hydrogen (with carbon monoxide as sole carbon and energy source).
-The hypothesis tested in (Norman et al., unpublished), was that in the case where hydrogen production infeasible, alternative electron sinks like lactate and \textit{2,3 BD} would be produced.
+The hypothesis tested in (\cite{norman2018progress}, under review), was that in the case where hydrogen production infeasible, alternative electron sinks like lactate and \textit{2,3 BD} would be produced.
 As in the previous case, the model-predicted behaviour showing both lactate and \textit{2,3 BD} was deemed an important result, which model curators should be notified of if lost during model development.
@@
 For these reasons, a test has been built into the MetaCLAU project which enables the computation of the set of essential genes and their comparison with TraDIS data sets.
-
-These test cases are all fully reproducible and available as an example within the \textit{gsmodutils} software repository and we encourage the reader to explore these examples to understand how to utilise \textit{gsmodutils} within their own project.
 

### 2019-02-01 article/inheritence.tex
Updated manuscript

\documentclass[tikz,border=3pt]{standalone}
\usetikzlibrary{shapes.geometric}
% \usetikzlibrary{arrows.meta}
\usetikzlibrary{decorations.text}
\definecolor{mygray}{RGB}{143,188,143}
\definecolor{boxgray}{RGB}{200,200,200}

\definecolor{testblue}{RGB}{200,200,200}
\definecolor{designyellow}{RGB}{200,200,200}

\definecolor{darkgreen}{RGB}{0,100,0}
\newcommand*{\mytextstyle}{\sffamily\Large\bfseries\color{black!85}}

\newcommand*{\boxtextstyle}{\sffamily\small\bfseries\color{black!85}}

\begin{document}
\begin{tikzpicture}[square/.style={regular polygon,regular polygon sides=4}]

    \node[fill=gray!20, text width = 7cm,draw=black]  at (-1,3.5)[
      font  = \boxtextstyle,
      align = left](design)
      {Design deltas - reusable, inheritable changes};

    \node[fill=blue!20, text width = 3cm,draw=black]  at (-3,2.5)[
      font  = \boxtextstyle,
      align = left](design)
      {Wild-type strain};

    \node[fill=red!20, text width = 3cm, draw=black]  at (3, 2.5) [
      font  = \boxtextstyle,
      align = left] (learn) {Knock-out strain};

    \node[fill=green!20, text width = 3cm, draw=black]  at (9, 2.5) [
      font  = \boxtextstyle,
      align = left] (learn) {Production strains};

    %design

    \node at (-3,0) {\includegraphics[width=4cm]{model.png}};

    \node at (3,0) {\includegraphics[width=4cm]{model.png}};

    \node at (3.65,0.3) {\includegraphics[width=0.25cm]{cross.png}};
    \node at (2.9,-0.74) {\includegraphics[width=0.25cm]{cross.png}};

    \node at (9,0) {\includegraphics[width=4cm]{model.png}};

    \node at (9.65,0.3) {\includegraphics[width=0.25cm]{cross.png}};
    \node at (8.9,-0.74) {\includegraphics[width=0.25cm]{cross.png}};

    \node at (9.1,-1.5) {\includegraphics[width=0.9cm, angle=300]{pathway.eps}};

    \node at (9,-4) {\includegraphics[width=4cm]{model.png}};
     \node at (9.1,-5.5) {\includegraphics[width=0.9cm, angle=300]{pathway.eps}};

    \path [->,very thick]  (-1, 0 ) edge (1, 0);

    \node [fill=gray!20] (CRA) at (0,0.4) {\includegraphics[width=0.25cm]{cross.png} \includegraphics[width=0.25cm]{cross.png}};

    \path [->,very thick]  (5, 0 ) edge (7, 0);

    \node [fill=gray!20]  at (6,0.4) {\includegraphics[width=0.25cm]{cross.png} \includegraphics[width=0.25cm]{cross.png}};

    \node [fill=gray!20] (PA) at (6,-0.4) {\includegraphics[width=1cm]{pathway.eps}};

    \path [-,very thick]  (-3, -2) edge (-3, -4.02);

    \node [fill=gray!20] (PB) at (6,-3.6) {\includegraphics[width=1cm]{pathway.eps}};
    \path [->,very thick]  (-3, -4 ) edge (7, -4);

    \node[fill=gray!20, text width = 6.5cm,draw=black]  at (1,-5)[
      font  = \boxtextstyle,
      align = left](dbox)
      {Only store differences: \\ Relate changes to real strains};

   \node [fill=gray!20] at (3.6,-5) {\includegraphics[width=1cm]{pathway.eps}};
   \node [fill=gray!20] at (2.4,-5) {\includegraphics[width=0.25cm]{cross.png}};
   \node [fill=gray!20] at (2.8,-5) {\includegraphics[width=0.25cm]{cross.png}};

\end{tikzpicture}
\end{document}

### 2019-02-01 article/pathway.tex
Updated manuscript

--- 2019-02-01 article/inheritence.tex
@@
 \usetikzlibrary{decorations.text}
-\definecolor{mygray}{RGB}{143,188,143}
-\definecolor{boxgray}{RGB}{200,200,200}
-
-\definecolor{testblue}{RGB}{200,200,200}
-\definecolor{designyellow}{RGB}{200,200,200}
-
 \definecolor{darkgreen}{RGB}{0,100,0}
-\newcommand*{\mytextstyle}{\sffamily\Large\bfseries\color{black!85}}
-
-\newcommand*{\boxtextstyle}{\sffamily\small\bfseries\color{black!85}}
 
@@
 
-    \node[fill=gray!20, text width = 7cm,draw=black]  at (-1,3.5)[
-      font  = \boxtextstyle,
-      align = left](design)
-      {Design deltas - reusable, inheritable changes};
+    \node[circle, fill=darkgreen, draw=black, text width=0.3cm]  at (0, 0) (A) {};
+    \node[circle, fill=darkgreen, draw=black, text width=0.3cm]  at (2, 0) (B) {};
 
-    \node[fill=blue!20, text width = 3cm,draw=black]  at (-3,2.5)[
-      font  = \boxtextstyle,
-      align = left](design)
-      {Wild-type strain};
-
-    \node[fill=red!20, text width = 3cm, draw=black]  at (3, 2.5) [
-      font  = \boxtextstyle,
-      align = left] (learn) {Knock-out strain};
-
-    \node[fill=green!20, text width = 3cm, draw=black]  at (9, 2.5) [
-      font  = \boxtextstyle,
-      align = left] (learn) {Production strains};
-
-    %design
-
-    \node at (-3,0) {\includegraphics[width=4cm]{model.png}};
-
-    \node at (3,0) {\includegraphics[width=4cm]{model.png}};
-
-    \node at (3.65,0.3) {\includegraphics[width=0.25cm]{cross.png}};
-    \node at (2.9,-0.74) {\includegraphics[width=0.25cm]{cross.png}};
-
-    \node at (9,0) {\includegraphics[width=4cm]{model.png}};
-
-    \node at (9.65,0.3) {\includegraphics[width=0.25cm]{cross.png}};
-    \node at (8.9,-0.74) {\includegraphics[width=0.25cm]{cross.png}};
-
-    \node at (9.1,-1.5) {\includegraphics[width=0.9cm, angle=300]{pathway.eps}};
-
-    \node at (9,-4) {\includegraphics[width=4cm]{model.png}};
-     \node at (9.1,-5.5) {\includegraphics[width=0.9cm, angle=300]{pathway.eps}};
-
-    \path [->,very thick]  (-1, 0 ) edge (1, 0);
-
-    \node [fill=gray!20] (CRA) at (0,0.4) {\includegraphics[width=0.25cm]{cross.png} \includegraphics[width=0.25cm]{cross.png}};
-
-    \path [->,very thick]  (5, 0 ) edge (7, 0);
-
-    \node [fill=gray!20]  at (6,0.4) {\includegraphics[width=0.25cm]{cross.png} \includegraphics[width=0.25cm]{cross.png}};
-
-    \node [fill=gray!20] (PA) at (6,-0.4) {\includegraphics[width=1cm]{pathway.eps}};
-
-    \path [-,very thick]  (-3, -2) edge (-3, -4.02);
-
-    \node [fill=gray!20] (PB) at (6,-3.6) {\includegraphics[width=1cm]{pathway.eps}};
-    \path [->,very thick]  (-3, -4 ) edge (7, -4);
-
-    \node[fill=gray!20, text width = 6.5cm,draw=black]  at (1,-5)[
-      font  = \boxtextstyle,
-      align = left](dbox)
-      {Only store differences: \\ Relate changes to real strains};
-
-   \node [fill=gray!20] at (3.6,-5) {\includegraphics[width=1cm]{pathway.eps}};
-   \node [fill=gray!20] at (2.4,-5) {\includegraphics[width=0.25cm]{cross.png}};
-   \node [fill=gray!20] at (2.8,-5) {\includegraphics[width=0.25cm]{cross.png}};
+    \draw (-2,0) edge[->, very thick] (-0.4,0) ;
+    \draw (A) edge[->, very thick] (1.6,0) ;
 

### 2019-02-01 article/test_case.tex
Updated manuscript

\documentclass[border={35pt 10pt 150pt 60pt}, % left bottom right top
  svgnames]{standalone}
\usepackage{tikz}
\usetikzlibrary{positioning}
\usepackage{listings}
\lstset{%
  frame            = tb,    % draw frame at top and bottom of code block
  tabsize          = 1,     % tab space width
  numbers          = left,  % display line numbers on the left
  framesep         = 3pt,   % expand outward
  framerule        = 0.4pt, % expand outward 
  commentstyle     = \color{Green},      % comment color
  keywordstyle     = \color{blue},       % keyword color
  stringstyle      = \color{DarkRed},    % string color
  backgroundcolor  = \color{WhiteSmoke}, % backgroundcolor color
  showstringspaces = false,              % do not mark spaces in strings
}
\begin{document}
\begin{lstlisting}[language = python, numbers = none, escapechar = !,
    basicstyle = \ttfamily\bfseries, linewidth = \linewidth] 
        !
 \tikz[remember picture] \node [] (a) {}; !
@ModelTestSelector(designs=["calvin_cycle"]!\tikz[remember picture]\node[](b){};!)
def test_fluxomic_data(model, log, project)!\tikz[remember picture] \node [] (c){};!
    """ 
    Example test comparing 13C fluxomic data to
    model predicted values.
    """
    path = project.project_path + "data/13C_flux_dt.csv"
    flux_data = pandas.read_csv(path) !\tikz[remember picture] \node [] (load_external_data){};!

    # Call to external cobrapy flux variability function
    fva_result = fva(model) !\tikz[remember picture] \node [] (cobra_fva){};!

    for row in flux_data.iterrows():
        # Get reactions found in the data file
        rxn = model.reactions.get_by_id(row["rxn_id"])
        # Compare against fva
        stmt = fva_result[rxn.id].minimum >= row["min"]\!\tikz[remember picture] \node [] (statment){};!
                        and fva_result <= row["max"]
        log.assertion( !\tikz[remember picture] \node [] (d){};!
            stmt,
            "Reaction {} inside range".format(rxn.id),
            "Reaction {} outside range".format(rxn.id)
        )

\end{lstlisting}
\begin{tikzpicture}[remember picture, overlay,
    every edge/.append style = { ->, thick, >=stealth,
                                  DimGray, dashed, line width = 1pt },
    every node/.append style = { align = center, minimum height = 10pt,
                                 font = \bfseries, fill= green!20},
                  text width = 2.5cm ]
  \node [above left = 0.5cm and -1 cm of a,text width = 3cm] (A) {Model selection decorator};
  \node [right = 4cm of A, text width = 2cm] (B) {Strain design selection};

  \node[right=3.5cm of load_external_data] (Ext) {Loading external data};

  \node [right = 2.6cm of c] (C) {Required parameters};
  \node [right = 7.2cm of d]  (D) {Formal test definition with success and failure messages};
  \draw (A.south) + (0, 0) coordinate(x1) edge (x1|-b.north);
  \draw (B.south) + (0, 0) coordinate(x2) edge (x2|-b.north);
  \draw (Ext.west) edge (load_external_data.east) ;
  \draw (C.west) edge (c.east) ;
  \draw (D.west) edge (d.east) ;
\end{tikzpicture} 
\end{document}

### 2019-02-01 article/buildscript.sh

@@ -1,0 +1,20 @@
+#!/bin/bash
+
+pdflatex pathway.tex
+
+pdftops -eps pathway.pdf
+
+pdflatex cycle.tex
+pdflatex design.tex
+pdflatex test_case.tex
+pdflatex inheritence.tex
+
+pdftops -eps cycle.pdf
+pdftops -eps design.pdf
+pdftops -eps test_case.pdf
+pdftops -eps inheritence.pdf
+
+pdflatex article.tex
+bibtex article.aux
+pdflatex article.tex
+

### 2019-10-14 article/article.tex

@@
 Constraints based models of this form can include thousands of reactions and metabolites, with many crucial pathways that only become activated in specific simulation settings.
-However, despite their widespread use, power and the availability of tools to aid with the construction and analysis of large scale models, little methodology is suggested for the continued management of curated large scale models.
-For example, when genome annotations are updated or new understanding regarding behaviour of is discovered, models often need to be altered to reflect this.
-This is quickly becoming an issue for industrial systems and synthetic biotechnology applications, which require good quality reusable models integral to the design, build and test cycle.
+However, despite their widespread use, power and the availability of tools to aid with the construction and analysis of large scale models, little methodology is suggested for their continued management.
+For example, when genome annotations are updated or new understanding regarding behaviour is discovered, models often need to be altered to reflect this.
+This is quickly becoming an issue for industrial systems and synthetic biotechnology applications, which require good quality reusable models integral to the design, build, test and learn cycle.
 \\
@@
 Crucially, different conditions are configurable allowing users to examine how different designs or curation impact a wide range of system behaviours, minimising error between model versions.\\
-\textbf{Availability:} The software framework described within this paper is open source and freely available from \url{http://github.com/SBRCNottingham/gsmodutils} \\
+\textbf{Availability:} The software framework described within this paper is open source and freely available from http://github.com/SBRCNottingham/gsmodutils \\
+\textbf{Supplementary information:} Supplementary data are available online.
 \end{abstract}
@@
 Such large scale reconstructions are often referred to as genome scale metabolic models (GSMMs), as the processes is significantly aided through the advent of relatively inexpensive genome sequencing \cite{o2015using, land2015insights}.
-Indeed, owing to their ability to model complex aspects of metabolism, GSMMs have been widely adopted as a standard to elucidate and optimise of industrial biotechnology processes \cite{kim2017current}.
+Indeed, owing to their ability to model complex aspects of metabolism, GSMMs have been widely adopted as a standard to elucidate and optimise industrial biotechnology processes \cite{kim2017current}.
 
-The reconstruction of GSMMs is an arduous process that follows a complex protocol to ensure model validity \cite{thiele2010protocol}.
-Whilst many popular automated methods exists to construct GSMMs from reference genomes \cite{henry2010high, poolman2006scrumpy}, there is a still a significant amount of manual curation.
+The reconstruction of GSMMs is a time consuming process of manual curation that follows a complex protocol to ensure model validity \cite{thiele2010protocol}.
+Whilst many popular automated methods exist to construct GSMMs from reference genomes \cite{henry2010high, poolman2006scrumpy}, there is still a significant amount of manual curation.
 However, treating the creation of models as an isolated ``one-off'' event ignores the significant amount of curation that is required for applications such as biotechnology.
@@
 The BiGG models database \cite{king2015bigg}, for example, exists to provide a standardised repository of validated models that can be shared and reused.
-However, little focus is placed upon the collaborative design aspect of such models with few mechanisms existing to capture the differences between two model versions, \textit{model deltas}.
+Perhaps one of the best examples of a well curated model developed in an iterative manner is yeast-GEM, a model for \textit{Saccharomyces cerevisiae} under continuous development \cite{benjamin_sanchez_2018_1994089}.
 Similarly, the MetaNetX \cite{moretti2016metanetx} system exists to provide a standardised namespace and toolchain for GSM analysis.
-However, such tools often make it difficult to understand the design decisions made by the initial model authors.
+However, in many cases too little focus is placed upon the collaborative design aspect of such models with few mechanisms existing to capture the differences between two model versions, \textit{model deltas}.
 
-Furthermore, as with many areas of bioinformatic study the number available computational tools has become vast.
-This covers huge variety of software platforms including the COBRA toolbox for Matlab \cite{schellenberger2011quantitative}, ScrumPy and COBRApy in python \cite{poolman2006scrumpy, ebrahim2013cobrapy} with additional tools and libraries such as cameo \cite{cardoso2017cameo}, OptFlux in Java \cite{rocha2010optflux} and SurreyFBA \cite{gevorgyan2010surreyfba}.
+Better tools for developing automated reconstructions of genome scale models are always under development.
+For example, a recent development in the domain of genome scale models is the EMBL-GEMS model repository for automated reconstructions of bacterial species created from NCBI annotations by the CarveMe tool \cite{machado2018fast}.
+However, these tools will often add incorrect reactions, such as aerobic reactions in organisms that only survive in anaerobic environments \cite{norman2018agenome}.
+Furthermore, we feel a specific advantage of genome scale models is that they encode domain specific knowledge that allows contradictions in understanding to be uncovered.
+For example, an entirely automated process is unlikely to find the correct cofactors involved in reactions under specific conditions.
+
+Furthermore, as with many areas of bioinformatic study the number of available computational tools has become vast.
+This covers a huge variety of software platforms including the COBRA toolbox for MATLAB \cite{schellenberger2011quantitative}, ScrumPy and COBRApy in python \cite{poolman2006scrumpy, ebrahim2013cobrapy} with additional tools and libraries such as cameo \cite{cardoso2018cameo}, OptFlux in Java \cite{rocha2010optflux} and SurreyFBA \cite{gevorgyan2010surreyfba}.
 Whilst most of these tools are Open Source and follow standards, such as SBML \cite{finney2003systems}, it is often challenging to replicate the initial modelling efforts conducted by authors of papers.
 Consequently, we feel that software tools are urgently needed to address this issue.
-
-Similarly, The archetype design, build and test cycle of synthetic biology heavily relies on the usage of bioinformatics software and modelling to improve the production of natural products \cite{carbonell2016bioinformatics}.
+Similarly, the archetype design, build, test and learn cycle of synthetic biology heavily relies on bioinformatics software and modelling to improve the production of natural products \cite{carbonell2016bioinformatics}.
 In order to speed up the use of bioinformatics tools to produce high value platform chemicals, genome scale models are often used to discover methods for process optimisation \textit{in silico}.
@@
 
-%Many of the results presented in original sources are difficult to compare due to a lack of standardisation of reference identifiers for reaction names.
-
 Unfortunately, many of the results reported in \cite{lee2008genome, senger2008genome, mcanulty2012genome, dash2014capturing, yoo2015quantitative} are difficult to compare or reproduce as the result of a number of issues.
-Often, the models lack a standardisation of identifiers for reaction names, which makes direct comparison of model structure as well as differences between reactions a challenge.
+Often, model authors do not use a standardised set of identifiers for reaction names (such as the MetaNetX namespace \cite{moretti2016metanetx}), which makes direct comparison of model structure as well as differences between reactions a challenge.
 Where models are shared, it is often in non-standard spreadsheet formats, rather than SBML models.
@@
 
-In this paper, we present a software framework geared toward \textit{test-driven} genome scale model development, a concept that is taken directly from good software development practices \cite{martin2002agile}.
+In this paper, we present a software framework geared towards \textit{test-driven} genome scale model development, a concept that is taken directly from good software development practices \cite{martin2002agile}.
 By this we mean the notion that, as a model is curated to represent biological phenomena, much of the validation can be turned into specific test cases that can be repeated between model versions.
 We provide an example test case for \textit{Clostridium autoethanogenum}, an organism that has had considerable focus in terms of genome scale models and how a working methodology using the software presented here can reduce repetition of work and improve the reproducibility of results.
-This article aims to summarise the main objectives of the \textit{gsmodutils} software and we refer the reader to the software user guide for a more detailed exploration features.
+This article aims to summarise the main objectives of the \textit{gsmodutils} software and we refer the reader to the software user guide for a more detailed exploration of features.
+
+\begin{figure}[ht]
+ \includegraphics[width=0.9\textwidth]{cycle.eps}
+ \caption{Iterative cycle for systems and synthetic biology development, prevalent in industrial biotechnology applications.
+This approach captures an iterative mode of development, where models are used to inform wet lab decision making and the information is fed back into future modelling decisions.
+By integrating test-driven model development (top section) the objective is to simultaneously capture research questions, model validation criteria and minimise the impact of changes on previously completed models.
+ }
+ \label{fig:cycle}
+ \end{figure}
 
@@
 
-The design phase of industrial synthetic biotechnology, due to historical and regulatory convention, often follows a waterfall model for development of genome scale models.
-This is borne out of a number significant constraints in the development of genome scale models.
-The waterfall approach for modelling is characterised in Figure \ref{fig:methodology_approaches} where requirements for a model are captured at an early stage and are hard to refine during the production process.
+Recent efforts in systems and synthetic biology have been based around a form of iterative, design, build, test and learn cycle \cite{carbonell2016bioinformatics} (see Figure \ref{fig:cycle}).
+In terms of computational models, this iterative strategy requires adapting and updating models to integrate new biological knowledge \cite{reed2006towards}.
+However, the conventional processes of scientific literature often coalesces to a point at which models are published.
+In reality, an iterative process means that it is essential that digital experiments can be repeated in a reproducible manner \cite{cooper2015call}.
+Future changes to models, borne out of a need to meet new challenges and integrate new knowledge, should reflect this.
 
-In contrast, an \textit{agile} methodology for the development of models places the focus on adapting work to new requirements \cite{martin2002agile}.
+An \textit{agile} methodology for the development of models places the focus on adapting work to new requirements \cite{martin2002agile}.
 Such an approach best fits genome scale models because they are rarely created to investigate individual processes and, instead, capture the complexity of large systems.
@@
 Such annotations rely heavily on automated matches to related species, with the characterisations of individual genes or changes in cofactors and substrates for specific reactions often being left to a few of critical interest \cite{seemann2014prokka}.
-This modelling formation costs an in depth understanding of dynamic behaviour. 
+This modelling formation costs an in-depth understanding of dynamic behaviour. 
 However, capturing steady-state phenomena still provides a good understanding of system properties \cite{o2015using}.
@@
 
-\begin{figure}[ht]
-\includegraphics[width=\textwidth]{waterfall.eps}
-\caption{Conventional waterfall methodology for model development.
-Each process is considered to be an isolated aspect of model development.}
-\label{fig:methodology_approaches}
-\end{figure}
-
 Therefore, models undergo significant manual annotation and curation; a process which has a high chance of error.
-In this work, we advocate a test-driven approach to model development highlighted in Figure \ref{fig:test_driven}.
+In this work, we advocate a test-driven approach to model development highlighted in Figure \ref{fig:cycle} (top).
 Here, the model is changed to achieve research goals that are dynamic in response to the changes of a project.
-In order to meet this objective, validation criteria for a model such as growth conditions or the impact of gene knock outs, should be formally set.
+In order to meet this objective, validation criteria for a model such as growth conditions or the impact of gene knock-outs, should be formally set.
 When a model is changed, all such validation criteria should be retested to ensure that models do not regress to previous states.
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width=0.8\textwidth]{testdriven.eps}
-\caption{Workflow for model development proposed for \textit{gsmodutils}.
-The objective of this approach is to simultaneously capture research questions, model validation criteria and minimise the impact of changes on previously completed models.
-}
-\label{fig:test_driven}
-\end{figure}
 
@@
 \subsection{Outline and features}
-Test-driven development is driven by the idea of clearly defined test cases written before significant changes are made to any underlying architecture.
+Test-driven development is motivated by the idea of clearly defined test cases written before significant changes are made to any underlying architecture.
 In the case of genome scale models, errors can easily occur as a product of human curation designed to better represent newly discovered aspects of metabolism.
@@
 The use of flat files enables easy integration with version control software such as git and mercurial.
-In addition, projects are easy to export using portable standardised docker images \cite{merkel2014docker}, the idea being to allow users to share models as quickly and easily as possible without concern for custom system configurations.
+In addition, projects are easy to export using portable standardised docker images \cite{merkel2014docker}, the idea being to allow users to share models as quickly and easily as possible without concern for custom system configurations (see software documentation for more details).
 
 \begin{figure}[ht]
-
-\lstset{language=python}   
-\begin{lstlisting}
-def test_func(model, project, log):
-    status = model.solver.optimize()
-    log.assertion(status == 'optimal',
-                  ``Works'', # Success message
-		  ``Does not work'', # Failure message
-                  ``Test description'')
-\end{lstlisting}
-\caption{An example \textit{gsmodutils} test case written in python}
+\includegraphics[trim={1cm 0 2.5cm 0},clip,width=0.9\textwidth]{test_case.eps}
+\caption{An example \textit{gsmodutils} test case written in python. 
+In this test, flux variability analysis is used to compare a model against $^{13}$C carbon flux tracking data.
+The test also demonstrates how designs can be integrated into a test workflow by specifying the identifier in the `ModelTestSelector' function decorator.
+}
 \label{fig:test_case} 
@@
 \subsection{Strain designs}
+
+\begin{figure*}[ht]
+  \includegraphics[width=\textwidth]{inheritence.eps}
+  \caption{Examples of \textit{gsmodutils} design inheritance.
+  Each design stores the delta between the wild-type base model, any parents and the changes to constraints the design contains.
+  In the example presented above, a heterologous production pathway is combined with a reusable set of knock-outs.
+  Rather than keeping redundant copies of models, designs make projects easier to maintain and understand by only storing annotated differences between models.
+  Designs can then be loaded in a hierarchical manner.
+  In practice, ideally, these designs should relate to experimentally evaluated constructs and strains.
+  }
+    \label{fig:strain_hered}
+\end{figure*}
+
 A core aspect behind the implementation of \textit{gsmodutils} is the concept of a \textit{design}, this encompasses a simple set of changes to a ``wild-type'' model that are required for analysis.
-For example a mutant strain that has undergone several gene deletions is not sufficiently different from the original wild-type to merit having its own genome scale model.
 However, it is often the case that such deletions are of scientific or industrial interest and, as such, the strain will be used in future work.
 Consequently, such designs are hereditary in nature.
-By taking the \textit{model delta} between the constraints applied to an initial model and subsequent modifications, \textit{gsmodutils} allows users to easily reuse and export models with this design.
-As designs inherit from a base model, future curation to a wild type model will automatically be included in the designs.
-Similarly, designs are self contained and will not interfere with one another allowing a form of project management.
+By taking the difference between the constraints applied to an initial model and subsequent modifications, \textit{gsmodutils} allows users to easily reuse and export models with this \textit{design delta}.
+
+Formally, we consider a \textit{design delta} to be the set of differences between any reactions, metabolites and genes stored within the COBRApy object.
+This should not be considered the same as a \textit{diff} in version control systems such as git. 
+Instead, designs of this nature are stored as JSON objects within the \textit{gsmodutils} project and can be tracked by version control systems.
+
+As designs inherit from a base model, future curation to a wild-type base model is automatically included in the resulting models.
+Similarly, designs are self contained and will not interfere with one another allowing project management and annotation as to the function of each design.
 Figure \ref{fig:strain_hered} shows how this could work in a practical situation.
+Here we consider how functional gene knock-outs can be combined with heterologous genes to create production strains.
+As designs can be inherited common knock-outs or changes to designs can be combined.
 
@@
 An example of this is shown in Figure \ref{fig:programmatic_design}.
-This example converts an existing model to one based on a Mixed Integer Linear Program (MILP), as the reaction names do not need to be specified, should reactions be added or removed the process will still apply.
+This example converts an existing model to one based on a Mixed Integer Linear Program (MILP) and sets the objective to minimise the number of enzymatic reactions used with a fixed biomass constraint.
+This relates to minimising the number of active genes within a system.
+As the reaction names do not need to be specified, should reactions be altered within the base model the design will remain functionally the same.
 Alternative examples could include reductions of models through methods such as elementary flux modes or minimum cut sets, which can change dramatically with only small changes to stoichiometry.
-Furthermore, integrity of all strain designs is automatically tested as part of the default \textit{gsmodutils} testing framework.
-
-\begin{figure}[ht]
-\lstset{language=python}   
-\begin{lstlisting}
-def gsmdesign_example_milp(model, project):
-    for reaction in model.reactions:
-	reaction.variable_kind = "integer"
-    return model
-\end{lstlisting}
-\caption{An example \textit{gsmodutils} programmatic design written in python. 
-This design converts reactions to integer type, allowing an MILP formation.}
-\label{fig:programmatic_design} 
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\textwidth]{inheritence2.eps}
-  \caption{Examples of \textit{gsmodutils} design inheritance.
-  Each design stores the delta between the wild-type base model, any parents and the changes to constraints the design contains.
-  In the example presented above a heterologous production pathway is combined with a reusable set of knock-outs.
-  In practice, ideally, these designs should relate to real constructs and strains evaluated by experiment.
-  }
-    \label{fig:strain_hered}
-\end{figure}
+Furthermore, functionality of all strain designs is automatically included in tests as part of the default \textit{gsmodutils} testing framework.
 
 \subsection{Development workflow}
-In this section we propose a method for the development of genome scale models that integrates \textit{gsmodutils} with version control systems.
+In this section, we propose a method for the development of genome scale models that 
+integrates \textit{gsmodutils} with version control systems.
+The basic workflow is that the user writes a formal test case for some modelling goal, perhaps driven by captured experimental data, that fits a specific form of validation criteria.
+We note that, in principle, test cases should be written before changes to a model are made.
+
 Figure \ref{fig:test_case} highlights the notion of test cases, taken from test-driven development.
-The outlined workflow is that the user writes a formal test case for some modelling goal, perhaps driven by captured experimental data, that fits a specific form of validation criteria.
-We note that, in principle, this test case should be written before changes to a model.
+In this example, a reusable test is written that incorperates data from $^{13}$C metabolic flux tracking.
+Flux variability analysis is then used to compare the expected flux ranges of a model against the experimental data.
+The test assertions will pass or fail based upon the models flux values when compared to the experimentally observed data.
 
 \section{Case study usage Clostridium autoethanogenum}
-\textit{Clostridium autoethanogenum} is a bacterial species used for the production of commodity chemicals at industrial scale \cite{abrini1994clostridium,norman2018progress,LT}.
+
+\textit{Clostridium autoethanogenum} is a bacterial species used for the production of commodity chemicals at industrial scale \cite{abrini1994clostridium,norman2018progress}.
 A new GSMM of \textit{C.
-autoethanogenum}, `MetaCLAU', has been analysed to improve this bioprocess (\cite{norman2018progress}, submitted for peer review).
+autoethanogenum}, `MetaCLAU', has been analysed to improve this bioprocess \cite{norman2018agenome} (submitted for peer review).
 In this section we describe how \textit{gsmodutils} has been utilised to ensure that future versions of MetaCLAU will remain functionally relevant from the perspective of industrial biotechnology.
@@
 \textit{C.
-autoethanogenum} is a strictly anaerobic, acetogenic bacterium which naturally produces ethanol and trace amounts of \textit{2,3-butanediol} (\textit{2,3 BD}) from carbon monoxide and water \cite{abrini1994clostridium,schuchmann2014autotrophy,norman2018progress}.
-Since carbon monoxide is readily available in the form of industrial waste gas, and \textit{2,3 BD} has a global market value of \$43 billion \cite{kopke201123butanediol}, the optimisation of yields of \textit{2,3 BD} from carbon monoxide is highly desirable in the context of industry \cite{norman2018progress}.
+autoethanogenum} is a strictly anaerobic, acetogenic bacterium which naturally produces ethanol and trace amounts of 2,3-butanediol (2,3-BD) from carbon monoxide and water \cite{abrini1994clostridium,schuchmann2014autotrophy,norman2018progress}.
+Since carbon monoxide is readily available in the form of industrial waste gas, and 2,3-BD has a global market value of \$43 billion \cite{kopke201123butanediol}, the optimisation of yields of 2,3-BD from carbon monoxide is highly desirable in the context of industry \cite{norman2018progress}.
 MetaCLAU was built using Pathway Tools \cite{karp2002pathway} and ScrumPy \cite{poolman2006scrumpy}, and is based on a manually annotated genome sequence of \textit{C.
@@
 The resulting model consisted of 758 reactions, 773 metabolites and 518 genes.
-For full details of the model, see (\cite{norman2018progress}.
-submitted for peer review).
+For full details of the model, see \cite{norman2018agenome} (submitted for peer review).
 The model has been integrated with the \textit{gsmodutils} modelling framework as a test-driven project. 
@@
 \subsection{Evaluation of model validation criteria}
+In this section we outline specific test criteria that have been applied for the \textit{C. autoethanogenum} model discussed in this study.
+All of the examples discussed here are available in more detail in the supplementary repository \textbf{Supplementary Material File S1}.
+
 \textbf{Energetic consistency:} 
 An important limitation of FBA is that optimal solutions may be thermodynamically infeasible if appropriate constraints are not applied \cite{Fell1197rohtua}.
-In order to identify these unwanted flux distributions and constrain the model such that they are not in the feasible solution space, a diagnostic FBA is applied with the following constraints:
-1) All transport reactions are constrained to allow no uptake (lower bound = 0), and 2) the ATPase reaction is given a fixed flux of one.
+In order to identify these unwanted flux distributions and to constrain the model such that they are not in the feasible solution space, a diagnostic FBA is applied with the following constraints:
+1) All transport reactions are constrained to allow no uptake, and 2) the ATPase reaction is given a fixed flux of one.
 If a (non-zero) solution to this problem exists, it must contain a thermodynamic inconsistency, which can be dealt with by manual inspection of the solution and modification of one or more of the involved reactions \cite{Fell1197rohtua}.
 
-\textbf{Minimisation of flux based tests:}
+\textbf{Flux-minimisation tests:}
 One conventional approach in FBA is to set an optimisation criterion of minimising flux across enzymatic reactions with a fixed biomass constraint \cite{holzhutter2006thegeneralised}. 
 The solution to this FBA problem represents minimal protein investment \cite{holzhutter2006thegeneralised}.
-Since execution of flux minimisation in COBRApy requires a model in which reversible reactions are split into two irreversible reactions, a \textit{gsmodutils} design was created in which all reversible reactions are split in a programmatic manner.
-In the case of MetaCLAU, the flux minimal solution includes both ethanol and acetate production, which represents good qualitative agreement with experimental data (\cite{norman2018progress}, under review).
-Since any changes to this predicted phenotype must be investigated, the flux minimisation analysis has been formulated as a \textit{gsmodutils} test.
+Execution of flux minimisation in COBRApy requires a model in which reversible reactions are split into two irreversible reactions, representing forward and backward reactions.
+A \textit{gsmodutils} strain design was created in which all reversible reactions are split using a programmatic python based design.
+
+In the case of MetaCLAU, the minimal-flux solution includes both ethanol and acetate production, which represents good qualitative agreement with experimental data (\cite{norman2018progress}, under review).
+Since any changes to this predicted phenotype must be investigated, the flux minimisation analysis has been formulated as a \textit{gsmodutils} test which utilises the above mentioned design.
 
@@
 autoethanogenum} under conditions where the organism can and cannot produce molecular hydrogen (with carbon monoxide as sole carbon and energy source).
-The hypothesis tested in (\cite{norman2018progress}, under review), was that in the case where hydrogen production infeasible, alternative electron sinks like lactate and \textit{2,3 BD} would be produced.
-As in the previous case, the model-predicted behaviour showing both lactate and \textit{2,3 BD} was deemed an important result, which model curators should be notified of if lost during model development.
+The hypothesis tested in (\cite{norman2018progress}, under review), was that in the case where hydrogen production is infeasible, alternative electron sinks like lactate and 2,3-BD would be produced.
+As in the previous case, the model-predicted behaviour showing both lactate and 2,3-BD was deemed an important result, which model curators should be notified of if lost during model development.
 Thus the analysis was built into a \textit{gsmodutils} test.
 
-\textbf{Lethal knockout mutants:} 
+\textbf{Lethal knock-out mutants:} 
 The prediction of lethal single-gene KO mutants through FBA of a GSMM is useful in two ways: 
@@
 and 2) with the advent of high-throughput TraDIS gene-essentiality data sets \cite{langridge2009simultaneous}, GSMMs can be validated by their ability to predict essential genes.
-Furthermore, any change in the set of essential genes (particularly an increase in the number of essentials) represents important information for metabolic engineering.
+Furthermore, any change in the set of essential genes (particularly an increase in their number) represents important information for metabolic engineering.
 For these reasons, a test has been built into the MetaCLAU project which enables the computation of the set of essential genes and their comparison with TraDIS data sets.
 
-\section{Related work}
-\textbf{Memote:} \cite{lieven2017memote} is tool set that has very similar ambitions to \textit{gsmodutils}.
+\begin{figure}[ht]
+\includegraphics[trim={1cm 0 4cm 0},clip,width=0.9\textwidth]{design.eps}
+\caption{An example \textit{gsmodutils} programmatic design written in python. 
+This design converts reactions to integer type, allowing an MILP formation.
+The above example seeks to utilise the MILP problem in order to minimise the number of reactions to produce the required biomass components.
+Loading a model of this form dynamically, as opposed to storing it as an SBML model, allows any underlying reactions to be changed.
+Designs of this form can also easily be exported to model files via the command line utility.
+}
+\label{fig:programmatic_design} 
+\end{figure}
+
+\vspace*{-12pt}
+\section{Related software}
+
+The reproducibility of computational based research has achieved more and more attention within the last decade \cite{peng2011reproducible, sandve2013ten, cooper2015call}.
+Consequently, there has been a proliferation of tools to support researchers in this endeavour.
+In this section we briefly review a number of tools that exist, both for genome scale models and from the wider mathematical and computational biology community.
+
+\textbf{PSAMM:} PSAMM \cite{steffensen2016psamm} is a tool that has similar goals to \textit{gsmodutils} in aiding the portability of genome scale models.
+PSAMM uses a custom YAML format which allows the annotation of models in a simpler manner than the conventional SBML standard.
+This, simultaneously allows model settings to be easily curated in a human friendly manner as well as allowing changes to be tracked in version control software such as git.
+This format relates, directly, to a \textit{gsmodutils} design object, which captures the difference between cobra objects.
+Both approaches allow export to standardised SBML, MATLAB and JSON models for usage in other tools.
+A core difference between the two frameworks, however, is that PSAMM includes much more functionality for working with genome scale models including gap filling and even its own interface to linear programming solvers.
+In contrast, \textit{gsmodutils} is designed to sit on top of the already existing COBRApy stack, with tools such as cameo \cite{cardoso2018cameo} providing additional functionality for more complex analysis.
+To this end, \textit{gsmodutils} also has a full python API allowing models and designs to be loaded from within external scripts or jupyter notebooks.
+
+\textbf{Memote:} \cite{lieven2018memote} is an excellent tool with similar ambitions to \textit{gsmodutils} for making reusable genome scale models.
 It features a fully specified set of tests, including custom test cases and has strong version control integration with git.
-A core difference between these projects is that \textit{gsmodutils} has a stronger focus on reducing the redundancy in model storage through the use of \textit{design deltas}, as described above.
+The core functionality of memote is to provide a standardised, community driven set of tests that check model consistency as well as annotations. 
+Along with user defined tests for experimental data, these tests allow continuous integration as models are updated within a git repository.
+One of memotes strongest aspects is providing human readable reports between model versions, this allows one to easily track the changes between annotations in reconstructions.
+
+A major difference between these projects is that \textit{gsmodutils} has a stronger focus on reducing the redundancy in model storage through the use of \textit{design deltas}, as described above.
 Similarly, a core goal of \textit{gsmodutils} is to allow easy import and export outside of the framework for compatibility with other modelling suites.
-It should be noted that, as memote is written in python, utilises COBRApy and, at the time of writing, it is fully compatible with \textit{gsmodutils}.
-\\
-\textbf{Cameo:} \cite{cardoso2017cameo} is a set of strain design utilities, such as genetic algorithms for finding optimal gene knock-out
-strategies, as well as other flux balance analysis utilities. Written in python and part of the COBRApy suite of applications, \textit{gsmodutils} works well with cameo.
-Indeed, making a set of design changes can easily be integrated into \textit{gsmodutils} through the python API.
-\\
-\textbf{Model repositories:} models are frequently shared, at the time of publication through services such as BiGG \cite{king2015bigg} and Biomodels \cite{chelliah2013biomodels}. 
-Whilst these repositories encourage the reuse of models and the reproducibility of \textit{in silico} predictions they are not designed to improve collaboration.
+It should be noted that, as memote is written in python, utilises COBRApy and, at the time of writing, is fully compatible with \textit{gsmodutils}.
+
+\textbf{Model repositories:} Models are frequently shared, at the time of publication through services such as BiGG \cite{king2015bigg} and BioModels \cite{chelliah2013biomodels}. 
+Whilst these repositories encourage the reuse of models and the reproducibility of \textit{in silico} predictions, they are not designed to improve collaboration.
 The software presented here is designed with the notion that genome scale models are never finished, \textit{per se}, but under continuous development.
 The cornerstone of this is the use of test cases, which formalise modelling validation criteria.
+Model repositories such as EMBL-GEMs, based on automated reconstructions generated by CarveMe \cite{machado2018fast}, could greatly benefit from an ever evolving set of tests that better capture biological understanding.
+Indeed, adding more features to control the future management of the BioModels repository has achieved recent attention with tools such as JUMMP \footnote{JUst a Model Management Platform (JUMMP) is available at {https://bitbucket.org/jummp/jummp/}, Accessed:2018-12-13} that aim to add version control to the management of repositories.
+
+Tools such as BiVes and BudHat \cite{scharm2018algorithm} also exist and apply more generally than \textit{gsmodutils} to capture the changes to models between versions.
+In a similar manner to \textit{gsmodutils}, this utility lends itself to version control between model versions by capturing changes to parameters that impact a models performance.
+As with \textit{gsmodutils}, this type of approach ignores irrelevant properties such as changes to the white space within XML files captured by a conventional unix diff.
 
@@
 In order to facilitate the sharing and dissemination of high quality computational research, good standards and software are required \cite{jimenez2017four}.
-Naturally a great deal of effort has gone in to producing high quality systems and synthetic biology standards \cite{hucka2003systems, cox2018synthetic}.
-Furthermore, when research projects end it all to is common for important large models to be published and become relics lost within the literature, forgotten to all but the most dedicated of individuals.
-As GSMMs grow in terms of the metabolism the contain as well as the biological problems they are used to solve, problems with annotation and curation naturally accumulate as a product of human error.
+Naturally a great deal of effort has gone into producing high quality systems and synthetic biology standards \cite{hucka2003systems, cox2018synthetic}.
+Furthermore, when research projects end it is common for important, large models to be published and become relics lost within the literature, forgotten to all but the most dedicated of individuals.
+As GSMMs grow in terms of the information about metabolism they contain as well as the biological problems they are used to solve, problems with annotation and curation naturally accumulate as a product of human error.
 Software that facilitates actively improving how researchers develop and apply models to new phenomena is required.
@@
 Furthermore, the objective of the project is to cultivate collaboration by simplifying the process of distributing large models to different users.
-As a consequence, methods for sharing reusable docker containers will be investigated.
 

### 2019-10-14 article/cycle.tex

\documentclass[tikz,border=3pt]{standalone}
% \usetikzlibrary{arrows.meta}
\usetikzlibrary{decorations.text}
\definecolor{mygray}{RGB}{143,188,143}
\definecolor{boxgray}{RGB}{200,200,200}

\definecolor{testblue}{RGB}{200,200,200}
\definecolor{designyellow}{RGB}{200,200,200}

\definecolor{darkgreen}{RGB}{0,100,0}
\newcommand*{\mytextstyle}{\sffamily\Large\bfseries\color{black!85}}

\newcommand*{\boxtextstyle}{\sffamily\small\bfseries\color{black!85}}
\newcommand{\arcarrow}[3]{%
   % inner radius, middle radius, outer radius, start angle,
   % end angle, tip protusion angle, options, text
   \pgfmathsetmacro{\rin}{1.7}
   \pgfmathsetmacro{\rmid}{2.2}
   \pgfmathsetmacro{\rout}{2.7}
   \pgfmathsetmacro{\astart}{#1}
   \pgfmathsetmacro{\aend}{#2}
   \pgfmathsetmacro{\atip}{5}
   \fill[mygray, very thick] (\astart+\atip:\rin)
                         arc (\astart+\atip:\aend:\rin)
      -- (\aend-\atip:\rmid)
      -- (\aend:\rout)   arc (\aend:\astart+\atip:\rout)
      -- (\astart:\rmid) -- cycle;
   \path[
      decoration = {
         text along path,
         text = {|\mytextstyle|#3},
         text align = {align = center},
         raise = -1.0ex
      },
      decorate
   ](\astart+\atip:\rmid) arc (\astart+\atip:\aend+\atip:\rmid);
}
\begin{document}
\begin{tikzpicture} %[>={Triangle[width=5mm,length=5mm]},->]

     \fill[even odd rule,green!8, draw=black] (-4.8, 7.1) rectangle (4.8,1.5);

    \node[
        text width = 5cm,
        font=\boxtextstyle,
        align=center] at (0, 5) {Iterative model \\ development};

    \fill[even odd rule,green!20] (0, 6.5) ellipse (1.5cm  and 0.5cm);
    \node[
        text width = 3cm,
        font=\boxtextstyle,
        align=center] (test) at (0, 6.5) {write test cases};

    \fill[even odd rule,blue!20] (3, 4.5) ellipse (1.5cm and 0.5cm);
    \node[
        text width = 3cm,
        font=\boxtextstyle,
        align=center
        ] (model) at (3, 4.5){update model to pass tests};

    \fill[even odd rule,red!20] (-3, 4.5) ellipse (1.5cm  and 0.5cm);
     \node[text width = 3cm, font=\boxtextstyle, align=center] (data) at (-3, 4.5)
     {design tests around new data};

    \path [->,very thick]  (model) edge[bend left] (data);
    \path [->,very thick] (data.north) edge[bend left] (test.west);
    \path [->,very thick] (test.east) edge[bend left] (model.north);

   \fill[even odd rule,darkgreen] circle (1.6);
   \node at (0,0) [
      font  = \mytextstyle,
      color = white,
      align = center
   ]{
        IB \\
      Development \\
      Cycle
   };
   \arcarrow{ 85}{  3}{Design}
   \arcarrow{270}{357}{Build}
   \arcarrow{182}{269}{Test}
   \arcarrow{176}{ 96}{Learn}

    %design
    \node[fill=blue!20, text width = 3cm,draw=black]  at (3,2.5)[
      font  = \boxtextstyle,
      align = left](design)
      {Use model to \\ predict behaviour};

%     \path [dashed] (design.north) edge (model.south);

    \node[fill=boxgray, text width = 3cm, draw=black] at (3,-2.5)[
      font  = \boxtextstyle,
      align = left]{Construct \\ 
        biological systems};

    \node[fill=boxgray, text width = 3cm, draw=black] at (-3,-2.5)[
      font  = \boxtextstyle,
      align = left]{Perform wet lab \\ 
      experiments};

    % learn

    \node[fill=red!20, text width = 3cm, draw=black]  at (-3, 2.5) [
      font  = \boxtextstyle,
      align = left] (learn) {Integrate new data\\ in to model};

%     \path [dashed] (learn.north) edge (data.south);

\end{tikzpicture}
\end{document}

### 2019-10-14 article/design.tex

--- 2019-02-01 article/test_case.tex
@@
 \begin{document}
-\begin{lstlisting}[language = python, numbers = none, escapechar = !,
+\begin{lstlisting}[language = python, numbers = none, escapechar = $,
     basicstyle = \ttfamily\bfseries, linewidth = \linewidth] 
-        !
- \tikz[remember picture] \node [] (a) {}; !
-@ModelTestSelector(designs=["calvin_cycle"]!\tikz[remember picture]\node[](b){};!)
-def test_fluxomic_data(model, log, project)!\tikz[remember picture] \node [] (c){};!
-    """ 
-    Example test comparing 13C fluxomic data to
-    model predicted values.
-    """
-    path = project.project_path + "data/13C_flux_dt.csv"
-    flux_data = pandas.read_csv(path) !\tikz[remember picture] \node [] (load_external_data){};!
+        $
+ \tikz[remember picture] \node [] (a) {}; $
+ @annotate(parent="calvin_cycle"$\tikz[remember picture]\node[](b){};$)
+ def gsmdesign_minimum_reactions_milp(model, project):$\tikz[remember picture] \node [] (c){};$
+    """ Docstring is used for description """
+    for reaction in model.reactions:
+        if reaction.id != "Biomass" 
+                and len(reaction.genes):
+            reaction.variable_kind = "integer"
 
-    # Call to external cobrapy flux variability function
-    fva_result = fva(model) !\tikz[remember picture] \node [] (cobra_fva){};!
+    model.reactions.Biomass.bounds = (0.2, 0.2)
+    model.objective_direction = "min"
+    model.solver = "cplex" $\tikz[remember picture] \node [] (d){};$
 
-    for row in flux_data.iterrows():
-        # Get reactions found in the data file
-        rxn = model.reactions.get_by_id(row["rxn_id"])
-        # Compare against fva
-        stmt = fva_result[rxn.id].minimum >= row["min"]\!\tikz[remember picture] \node [] (statment){};!
-                        and fva_result <= row["max"]
-        log.assertion( !\tikz[remember picture] \node [] (d){};!
-            stmt,
-            "Reaction {} inside range".format(rxn.id),
-            "Reaction {} outside range".format(rxn.id)
-        )
+    return model
 
@@
                   text width = 2.5cm ]
-  \node [above left = 0.5cm and -1 cm of a,text width = 3cm] (A) {Model selection decorator};
-  \node [right = 4cm of A, text width = 2cm] (B) {Strain design selection};
+  \node [above left = 0.5cm and -1 cm of a,text width = 3cm] (A) {Annotation decorator};
+  \node [right = 1.5cm of A, text width = 3cm] (B) {Parent design selection};
 
-  \node[right=3.5cm of load_external_data] (Ext) {Loading external data};
-
-  \node [right = 2.6cm of c] (C) {Required parameters};
-  \node [right = 7.2cm of d]  (D) {Formal test definition with success and failure messages};
+  \node [right = 2.0cm of B] (C) {Required parameters};
+  \node [right = 4.cm of d]  (D) {Setting custom solver interface};
   \draw (A.south) + (0, 0) coordinate(x1) edge (x1|-b.north);
   \draw (B.south) + (0, 0) coordinate(x2) edge (x2|-b.north);
-  \draw (Ext.west) edge (load_external_data.east) ;
-  \draw (C.west) edge (c.east) ;
+  \draw (C.south) + (0, 0) coordinate(x2) edge (x2|-b.north);
   \draw (D.west) edge (d.east) ;
